### 2024-03-21 scripts/data_collation.Rmd
writing script to clean and mmerge data frames

---
title: "data_collation"
author: "daniellembecker"
date: "2024-03-21"
output: html_document
---

# Use this script to clean and merge unique data types (e.g., swath, fish, UPC) from multiple years

# Load libraries 
```{r}
library(dplyr)
library(tidyverse)
```

# Load all data files from multiple years and seperate them into specific categories: UPC, swath, Morph, Site, QUAD, kelp, fish, and COVER
```{r}

# Set your directory path
directory <- "data/kelp_timeseries_data/"

# List files in the directory
files <- list.files(directory)

# Define categories
categories <- c("UPC", "swath", "Morph", "Site", "QUAD", "kelp", "fish", "COVER")

# Initialize an empty list to store data frames for each category
category_data_frames <- lapply(categories, function(cat) list())

# Read files into R and store as data frames
read_files <- function(files, category) {
  data_frames <- list()
  for (file in files) {
    # Read each file
    filepath <- file.path(directory, file)
    # Assuming your files are CSV files, adjust accordingly if they're different
    data <- tryCatch({
      read.csv(filepath, stringsAsFactors = FALSE, fileEncoding = "UTF-8")
    }, warning = function(w) {
      # Print the warning message
      print(paste("Warning while reading file:", filepath))
      print(w)
      # Return NULL if there's a warning
      return(NULL)
    }, error = function(e) {
      # Print the error message
      print(paste("Error reading file:", filepath))
      print(e)
      # Return NULL if there's an error
      return(NULL)
    })

    if (!is.null(data)) {
      # Extract subject from the file name
      subject <- gsub(".csv", "", file)  # Remove the file extension
      subject <- gsub(paste0(category, "_"), "", subject, fixed = TRUE)  # Remove category prefix

      # Store each data frame with the subject name as the key
      if (!subject %in% names(data_frames)) {
        data_frames[[subject]] <- data
      } else {
        # If subject already exists, combine the data frames
        data_frames[[subject]] <- rbind(data_frames[[subject]], data)
      }
    }
  }
  return(data_frames)
}

# Read files for each category into R and store as data frames
for (category in categories) {
  category_files <- files[grep(category, files)]
  category_data_frames[[category]] <- read_files(category_files, category)
}

# Now you have a list of lists of data frames for each category, combined by subjects in the title

# Create individual data frames for each category
UPC_df <- combined_data_frames[["UPC"]]
swath_df <- combined_data_frames[["swath"]]
Morph_df <- combined_data_frames[["Morph"]]
Site_df <- combined_data_frames[["Site"]]
QUAD_df <- combined_data_frames[["QUAD"]]
kelp_df <- combined_data_frames[["kelp"]]
fish_df <- combined_data_frames[["fish"]]
COVER_df <- combined_data_frames[["COVER"]]

```

### 2024-03-21 scripts/data_collation.Rmd
commbined data frames by category, some questions for cleaning up dataa for austin

@@
 ```{r}
+library(plyr)
 library(dplyr)
@@
 
-# Define categories
-categories <- c("UPC", "swath", "Morph", "Site", "QUAD", "kelp", "fish", "COVER")
+# Initialize an empty list to store data frames
+data_frames <- list()
 
-# Initialize an empty list to store data frames for each category
-category_data_frames <- lapply(categories, function(cat) list())
-
-# Read files into R and store as data frames
-read_files <- function(files, category) {
-  data_frames <- list()
+# Loop through each file
 for (file in files) {
-    # Read each file
-    filepath <- file.path(directory, file)
-    # Assuming your files are CSV files, adjust accordingly if they're different
-    data <- tryCatch({
-      read.csv(filepath, stringsAsFactors = FALSE, fileEncoding = "UTF-8")
-    }, warning = function(w) {
-      # Print the warning message
-      print(paste("Warning while reading file:", filepath))
-      print(w)
-      # Return NULL if there's a warning
-      return(NULL)
-    }, error = function(e) {
-      # Print the error message
-      print(paste("Error reading file:", filepath))
-      print(e)
-      # Return NULL if there's an error
-      return(NULL)
-    })
-
-    if (!is.null(data)) {
-      # Extract subject from the file name
-      subject <- gsub(".csv", "", file)  # Remove the file extension
-      subject <- gsub(paste0(category, "_"), "", subject, fixed = TRUE)  # Remove category prefix
-
-      # Store each data frame with the subject name as the key
-      if (!subject %in% names(data_frames)) {
-        data_frames[[subject]] <- data
-      } else {
-        # If subject already exists, combine the data frames
-        data_frames[[subject]] <- rbind(data_frames[[subject]], data)
-      }
-    }
-  }
-  return(data_frames)
+    # Read the file into a data frame
+    # Adjust the read function according to your file type (e.g., read.csv, read.table)
+    # Specify other parameters as needed (e.g., header = TRUE)
+    file_path <- file.path(directory, file)
+    df_name <- gsub("\\..*", "", file)  # Extract file name without extension
+    assign(df_name, read.csv(file_path, stringsAsFactors = FALSE))
 }
 
-# Read files for each category into R and store as data frames
-for (category in categories) {
-  category_files <- files[grep(category, files)]
-  category_data_frames[[category]] <- read_files(category_files, category)
-}
+# Now, each file is a separate data frame with its own variable name
 
-# Now you have a list of lists of data frames for each category, combined by subjects in the title
-
-# Create individual data frames for each category
-UPC_df <- combined_data_frames[["UPC"]]
-swath_df <- combined_data_frames[["swath"]]
-Morph_df <- combined_data_frames[["Morph"]]
-Site_df <- combined_data_frames[["Site"]]
-QUAD_df <- combined_data_frames[["QUAD"]]
-kelp_df <- combined_data_frames[["kelp"]]
-fish_df <- combined_data_frames[["fish"]]
-COVER_df <- combined_data_frames[["COVER"]]
+# COVER data frame is its own category, is this supposed to combine with another? it was labeled, Kelp_COVER
 
 ```
+
+# Combine fish dataframes
+```{r}
+
+# use rbind.fill() method in R to combine data frames with different columns
+fish_data <- rbind.fill(FtWeatherill_fish_2016, FtWeatherill_fish_2017, fish_2018_Hump_Lab_size_classes, fish, fish_Aggregated)
+
+```
+
+# Combine UPC dataframes
+```{r}
+
+# use rbind.fill() method in R to combine data frames with different columns
+UPC_data <- rbind.fill(UPC, UPC_2017, UPC_2018, FtWeatherill_UPC_2016, FtWeatherill_UPC_2016_CT)
+
+```
+
+# Combine swath dataframes
+```{r}
+
+# use rbind.fill() method in R to combine data frames with different columns
+swath_data <- rbind.fill(swath, swath_2018, swath_count, FtWeatherill_swath_2016)
+
+# the one swath data.frame has 115 variables because each species is listed instead of its code, fo you want me to replace with the code and make it wide to long format?
+
+```
+
+# Combine Kelp_Morph dataframes
+```{r}
+
+# use rbind.fill() method in R to combine data frames with different columns
+morph_data <- rbind.fill(Subsurface_Kelp_Morph_2017, Subsurface_Kelp_Morph_2018)
+
+```
+
+# Combine site.info dataframes
+```{r}
+
+# use rbind.fill() method in R to combine data frames with different columns
+site.info.dat <- rbind.fill(Site_info_2017, Site_info_2018, Site_info_template_v1)
+
+```
+
+# Combine site.info dataframes
+```{r}
+
+# use rbind.fill() method in R to combine data frames with different columns
+site.info.dat <- rbind.fill(Site_info_2017, Site_info_2018, Site_info_template_v1)
+
+```
+
+# Combine QUAD dataframes
+```{r}
+
+# use rbind.fill() method in R to combine data frames with different columns
+QUAD_data <- rbind.fill(QUAD, QUAD_2018, FtWeatherill_QUAD_2017, FtWeatherill_QUAD_2016_TALBENHORIN)
+
+# the one QUAD data.frame has 130 variables because each species is listed instead of its code, fo you want me to replace with the code and make it wide to long format?
+
+```
+
+# Combine kelp dataframes
+```{r}
+
+# use rbind.fill() method in R to combine data frames with different columns
+kelp_data <- rbind.fill(kelp, FtWeatherill_kelp_2016)
+
+#delete repetative data collector columns
+kelp_data <- kelp_data[ -c(13:16) ]
+
+```

### 2024-03-29 scripts/drop.long.line.Rmd
made boat type data plots

---
title: "drop.long.line"
author: "daniellembecker"
date: "2024-03-29"
output: html_document
---

# Load libraries 
```{r}
library(plyr)
library(dplyr)
library(tidyverse)
library(ggplot2)
```

```{r}

# Load dataframe
boat_data <- read.csv("data/dropline.longline.csv")

# Define the order of size categories
boat_data$size_cat <- factor(boat_data$size.cat, levels = c('Nano', 'Small', 'Medium', 'Large'))

# make stacked bar plot
stacked <- ggplot(boat_data, aes(x = size_cat)) +
  geom_bar(aes(y = dropline), fill = 'blue', width = 0.35, stat = "identity") +
  geom_bar(aes(y = longline), fill = 'lightblue', width = 0.35, stat = "identity") +
  labs(x = 'Size Category', y = 'Number of Boats', title = 'Number of Boats by Size Category') +
  theme_classic() +
  scale_y_continuous(expand = c(0, 0));stacked   # Remove space between bars and x-axis

# Save the plot
ggsave("output/stacked_plot.png", plot = stacked, width = 8, height = 6, dpi = 300)

#make not stacked bar plot

# Reshape the data from wide to long format
library(reshape2)
boat_data_long <- melt(boat_data, id.vars = "size_cat", measure.vars = c("dropline", "longline"))
boat_data_long$variable <- factor(boat_data_long$variable, levels = c("dropline", "longline"))

# Plot the data
bar.plot <- ggplot(boat_data_long, aes(x = size_cat, y = value, fill = variable)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.4), width = 0.35) +
  labs(x = 'Size Category', y = 'Number of Boats', title = 'Number of Boats by Size Category',
       fill = "Boat Type") +
  scale_fill_manual(values = c("blue", "lightblue"), labels = c("Dropline", "Longline")) +
  theme_classic() +
  scale_y_continuous(expand = c(0, 0)); bar.plot

# Save the plot
ggsave("output/unstacked.plot.png", plot = bar.plot, width = 8, height = 6, dpi = 300)

```

### 2024-04-12 scripts/data_collation.Rmd
commbined data frames by category, some questions for cleaning up dataa for austin
cleaning up data

@@
 
-# COVER data frame is its own category, is this supposed to combine with another? it was labeled, Kelp_COVER
+```
+
+# Combine kelp dataframes
+```{r}
+
+# use rbind.fill() method in R to combine data frames with different columns
+kelp_data <- rbind.fill(kelp, FtWeatherill_kelp_2016)
+
+#delete repetative data collector columns
+kelp_data <- kelp_data[ -c(11:17) ]
 
@@
 
+# remove columns with observer and unneeded data
+fish_data <- fish_data[ -c(10:13) ]
+
+# fish dataframe has size in wide format, discuss I can change this to long is you want but wanted to make sure size classes matched
+
 ```
@@
 
+# Rename ft. wetherill to fort wetherill
+UPC_data$SITE[UPC_data$SITE == "Ft. Weatherill"] <- "Fort Wetherill"
+
+# remove columns with observer and unneeded data
+UPC_data <- UPC_data[ -c(21:24) ]
 ```
@@
 swath_data <- rbind.fill(swath, swath_2018, swath_count, FtWeatherill_swath_2016)
+
+# Rename ft. wetherill to fort wetherill
+swath_data$SITE[swath_data$SITE == "Ft.Weatherill"] <- "Fort Wetherill"
 
@@
 
-```
-
-# Combine site.info dataframes
-```{r}
-
-# use rbind.fill() method in R to combine data frames with different columns
-site.info.dat <- rbind.fill(Site_info_2017, Site_info_2018, Site_info_template_v1)
-
+# remove columns with observer and unneeded data
+morph_data <- morph_data[ -c(11:14) ]
 ```
@@
 
+# Rename ft. wetherill to fort wetherill
+QUAD_data$SITE[QUAD_data$SITE == "Ft Wetherill"] <- "Fort Wetherill"
+
 # the one QUAD data.frame has 130 variables because each species is listed instead of its code, fo you want me to replace with the code and make it wide to long format?
@@
 ```
-
-# Combine kelp dataframes
-```{r}
-
-# use rbind.fill() method in R to combine data frames with different columns
-kelp_data <- rbind.fill(kelp, FtWeatherill_kelp_2016)
-
-#delete repetative data collector columns
-kelp_data <- kelp_data[ -c(13:16) ]
-
-```

### 2024-04-12 scripts/data_collation.Rmd
cleaning up data

@@
 ```{r}
+# Kelp data is cleaned and ready
 
@@
 
+# Combine Kelp_Morph dataframes
+```{r}
+
+# Morph data is cleaned and ready
+
+# use rbind.fill() method in R to combine data frames with different columns
+morph_data <- rbind.fill(Subsurface_Kelp_Morph_2017, Subsurface_Kelp_Morph_2018)
+
+# remove columns with observer and unneeded data
+morph_data <- morph_data[ -c(11:14) ]
+```
+
 # Combine fish dataframes
@@
 # use rbind.fill() method in R to combine data frames with different columns
-fish_data <- rbind.fill(FtWeatherill_fish_2016, FtWeatherill_fish_2017, fish_2018_Hump_Lab_size_classes, fish, fish_Aggregated)
+fish_data <- rbind.fill(FtWeatherill_fish_2016, FtWeatherill_fish_2017, fish_2018_Hump_Lab_size_classes)
 
@@
 
-# fish dataframe has size in wide format, discuss I can change this to long is you want but wanted to make sure size classes matched
+# look at fish_Aggregated later, this is kelp density possibly?
 
@@
 
-# Combine Kelp_Morph dataframes
-```{r}
-
-# use rbind.fill() method in R to combine data frames with different columns
-morph_data <- rbind.fill(Subsurface_Kelp_Morph_2017, Subsurface_Kelp_Morph_2018)
-
-# remove columns with observer and unneeded data
-morph_data <- morph_data[ -c(11:14) ]
-```
-
 # Combine site.info dataframes
@@
 
+# To do: species codes transform
+
 # use rbind.fill() method in R to combine data frames with different columns
@@
 
-# the one QUAD data.frame has 130 variables because each species is listed instead of its code, fo you want me to replace with the code and make it wide to long format?
-
 ```

### 2024-04-12 scripts/data_collation.Rmd
updated fish, mmorph, and kelp data

@@
 library(tidyverse)
+library(lubridate)
 ```
@@
 
+# Use lubridate to format date to be consistent
+kelp_data$DATE <- mdy(kelp_data$DATE)
+
+# Rename King's Beach to Kings Beach
+kelp_data$SITE[kelp_data$SITE == "King's Beach"] <- "Kings Beach"
+
+# output updated dataframe
+write.csv(kelp_data, "output/cleaned_data/kelp.data.clean.csv", row.names = FALSE)
 ```
@@
 morph_data <- morph_data[ -c(11:14) ]
+
+# Make all columns capitalized
+colnames(morph_data) <- toupper(colnames(morph_data))
+
+# Use lubridate to format date to be consistent
+morph_data$DATE <- mdy(morph_data$DATE)
+
+# output updated dataframe
+write.csv(morph_data, "output/cleaned_data/morph.data.clean.csv", row.names = FALSE)
 ```
@@
 
+# Fish data is cleaned and ready
+
 # use rbind.fill() method in R to combine data frames with different columns
@@
 # remove columns with observer and unneeded data
-fish_data <- fish_data[ -c(10:13) ]
+fish_data <- fish_data[ -c(10:14) ]
 
-# look at fish_Aggregated later, this is kelp density possibly?
+# Use the actual size data to bin into the size classes to combine with other dataframe
+fish$SIZE <- cut(fish$actual_size, breaks = c(0, 10, 20, 30, 40, 50, 70, 100, Inf),
+               labels = c("< 10", "10_20", "20_30", "30_40", "40_50", "50_70", "70_100", "> 100"),include.lowest = TRUE, right = FALSE)
+
+# combine updated dataframes
+fish_data <- left_join(fish, fish_data)
+
+# remove columns with VIS
+fish_data <- fish_data[ -c(14) ]
+
+# Use lubridate to format date to be consistent
+fish_data$DATE <- mdy(fish_data$DATE)
+
+# output updated dataframe
+write.csv(fish_data, "output/cleaned_data/fish.data.clean.csv", row.names = FALSE)
+
+# look at fish_Aggregated later - Austin, this is kelp density possibly?
 

### 2024-04-12 scripts/data_collation.Rmd
updated site info data

@@
 kelp_data$SITE[kelp_data$SITE == "King's Beach"] <- "Kings Beach"
+
+# Rename ft. weatherill to fort wetherill
+kelp_data$SITE[kelp_data$SITE == "Ft Weatherill"] <- "Fort Wetherill"
 
@@
 
+# Rename fort weatherill
+UPC_data$SITE[UPC_data$SITE == "Fort Weatherill"] <- "Fort Wetherill"
+
 # remove columns with observer and unneeded data
 UPC_data <- UPC_data[ -c(21:24) ]
+
+# Make all columns capitalized
+colnames(UPC_data) <- toupper(colnames(UPC_data))
+
+# Use lubridate to format date to be consistent
+UPC_data$DATE <- mdy(UPC_data$DATE)
+
+# output updated dataframe
+write.csv(UPC_data, "output/cleaned_data/UPC.data.clean.csv", row.names = FALSE)
+
+#UPC dataframe does not have month and day so that is why it isn't connecting fully
 ```
@@
 
-# use rbind.fill() method in R to combine data frames with different columns
+# Make all columns capitalized
+colnames(Site_info_2018) <- toupper(colnames(Site_info_2018))
+
+# combine site collection data
 site.info.dat <- rbind.fill(Site_info_2017, Site_info_2018, Site_info_template_v1)
+
+# Austin - the site info for 2018 has a start and stop latitude so don't know what to do here
 

### 2024-04-12 scripts/data_collation.Rmd
finished cleanign data

@@
 
+# UPC data is cleaned and ready
+
 # use rbind.fill() method in R to combine data frames with different columns
@@
 
-#UPC dataframe does not have month and day so that is why it isn't connecting fully
+#UPC data frame does not have month and day so that is why it isn't connecting fully
+```
+
+# Combine site.info dataframes
+```{r}
+
+# site information data is cleaned and ready
+
+# Make all columns capitalized
+colnames(Site_info_2018) <- toupper(colnames(Site_info_2018))
+
+# combine site collection data
+site.info.dat <- rbind.fill(Site_info_2017, Site_info_2018, Site_info_template_v1)
+
+# remove columns with observer and unneeded data
+site.info.dat <- site.info.dat[ -c(12:15) ]
+
+# output updated dataframe
+write.csv(site.info.dat, "output/cleaned_data/site.info.data.clean.csv", row.names = FALSE)
+
+# Austin - the site info for 2018 has a start and stop latitude so don't know what to do here
+
 ```
@@
 
-# the one swath data.frame has 115 variables because each species is listed instead of its code, fo you want me to replace with the code and make it wide to long format?
+# swath dataframe has species names as columns in wide format, need to change names to species codes to combine with other dataframes
 
-```
+# load species names and species code dataframe
+species.metadata <- read.csv("data/species_code_metadata.csv")
 
-# Combine site.info dataframes
-```{r}
+# Convert species names to codes
+for (col in names(swath)) {
+  swath[, col] <- species.metadata$code[match(swath[, col], species.metadata$species.substrate)]
+}
 
-# Make all columns capitalized
-colnames(Site_info_2018) <- toupper(colnames(Site_info_2018))
-
-# combine site collection data
-site.info.dat <- rbind.fill(Site_info_2017, Site_info_2018, Site_info_template_v1)
-
-# Austin - the site info for 2018 has a start and stop latitude so don't know what to do here
+#some of the species in swath dataframe are common names and species metadata in species names
 

### 2024-05-03 scripts/data_collation.Rmd
updated metadata sheet

@@
 # Convert species names to codes
-for (col in names(swath)) {
+updated.swath <- for (col in names(swath)) {
   swath[, col] <- species.metadata$code[match(swath[, col], species.metadata$species.substrate)]

### 2024-05-03 scripts/data_collation.Rmd
almost finished organizing swaath data

--- 2024-04-12 scripts/data_collation.Rmd
@@
 
+# swath dataframe has species names as columns in wide format, need to change names to species codes to combine with other dataframes
+
+# Load species names and species code data frame
+species.metadata <- read.csv("data/species_code_metadata.csv")
+
+# Function to rename columns using metadata
+# Get the unique column names from the swath dataframe
+column_names <- colnames(swath)
+
+# Create a lookup table with name-to-code mappings
+name_to_code <- setNames(species.metadata$code, species.metadata$name)
+
+# Rename the columns using the lookup table
+  new_column_names <- sapply(column_names, function(name) {
+    if (name %in% names(name_to_code)) {
+      return(name_to_code[name])
+    } else {
+      return(name)
+    }
+  })
+
+# Rename the columns
+colnames(swath) <- new_column_names
+
+# Convert data in numeric columns to numeric
+swath.con <- swath %>%
+  mutate(across(AGH:BLOA, as.numeric))
+
+# Melt only columns AGH:WHEL into long format
+swath_long <- swath.con %>%
+  pivot_longer(cols = AGH:BLOA, names_to = "SP_CODE", values_to = "value")
+
+# Need to now move values to binned swatch categories as columns
+# Pivot the dataframe into wide format
+swath_binned <- swath_long %>%
+  pivot_wider(names_from = swath, values_from = value)
+
+# make columns lowercase
+# Get the current column names
+current_names <- colnames(swath_binned)
+
+# Convert the column names to lowercase while preserving the first letter
+new_names <- sapply(current_names[1:6], function(x) {
+  paste0(toupper(substring(x, 1, 1)), tolower(substring(x, 2)))
+})
+
+# Assign the new column names to the first six columns of the dataframe
+colnames(swath_binned)[1:6] <- new_names
+
+# rename swatch category columns in swath count and swath 208 dataframe
+swath_2018 <- swath_2018 %>%
+  rename(
+    "0-20 IN" = "X0.20.IN",
+    "20-40 IN" = "X20.40.IN",
+    "40-20 OFF" = "X40.20.OFF",
+    "20-0 OFF" = "X20.0.OFF"
+  )
+
+swath_count <- swath_count %>%
+  rename(
+    "0-20 IN" = "X0.20.IN",
+    "20-40 IN" = "X20.40.IN",
+    "40-20 OFF" = "X40.20.OFF",
+    "20-0 OFF" = "X20.0.OFF"
+  )
+
+# Remove four columns in a specific order swath_2018
+swath_2018 <- swath_2018 %>%
+  select(-12:-15)
+
+# Remove four columns in a specific order swath_count
+swath_count <- swath_count %>%
+  select(-12:-15)
+
+# Remove four columns in a specific order FtWeatherill_swath_2016
+FtWeatherill_swath_2016 <- FtWeatherill_swath_2016 %>%
+  select(-9:-14)
+
+# make columns lowercase
+# Get the current column names
+current_names <- colnames(FtWeatherill_swath_2016)
+
+# Convert the column names to lowercase while preserving the first letter
+new_names <- sapply(current_names[1:6], function(x) {
+  paste0(toupper(substring(x, 1, 1)), tolower(substring(x, 2)))
+})
+
+# Assign the new column names to the first six columns of the dataframe
+colnames(FtWeatherill_swath_2016)[1:6] <- new_names
+
 # use rbind.fill() method in R to combine data frames with different columns
-swath_data <- rbind.fill(swath, swath_2018, swath_count, FtWeatherill_swath_2016)
+swath_data <- rbind.fill(swath_binned, swath_2018, swath_count, FtWeatherill_swath_2016)
 
@@
 
-# swath dataframe has species names as columns in wide format, need to change names to species codes to combine with other dataframes
-
-# load species names and species code dataframe
-species.metadata <- read.csv("data/species_code_metadata.csv")
-
-# Convert species names to codes
-for (col in names(swath)) {
-  swath[, col] <- species.metadata$code[match(swath[, col], species.metadata$species.substrate)]
-}
-
-#some of the species in swath dataframe are common names and species metadata in species names
+#still need to edit ft.wetherill dataframe
 

### 2024-05-17 scripts/data_collation.Rmd
updated UPC data

@@
 # remove columns with observer and unneeded data
-UPC_data <- UPC_data[ -c(21:24) ]
+UPC_data <- UPC_data[ -c(20:24) ]
 
@@
 
+# make columns lowercase
+# Get the current column names
+current_names <- colnames(UPC_data)
+
+# Convert the column names to lowercase while preserving the first letter
+new_names <- sapply(current_names[1:13], function(x) {
+  paste0(toupper(substring(x, 1, 1)), tolower(substring(x, 2)))
+})
+
+# Assign the new column names to the first six columns of the dataframe
+colnames(UPC_data)[1:13] <- new_names
+
 # output updated dataframe
@@
 
-#UPC data frame does not have month and day so that is why it isn't connecting fully
 ```
@@
 
+FtWeatherill_swath_2016 <- FtWeatherill_swath_2016 %>%
+  rename(
+    "0-20 IN" = "X0.20.IN",
+    "20-40 IN" = "X20.40.IN",
+    "40-20 OFF" = "X40.20.OFF",
+    "20-0 OFF" = "X20.0.OFF")
+
 # Remove four columns in a specific order swath_2018
@@
 
-# Remove four columns in a specific order FtWeatherill_swath_2016
-FtWeatherill_swath_2016 <- FtWeatherill_swath_2016 %>%
-  select(-9:-14)
-
 # make columns lowercase
 # Get the current column names
-current_names <- colnames(FtWeatherill_swath_2016)
+#current_names <- colnames(FtWeatherill_swath_2016)
 
 # Convert the column names to lowercase while preserving the first letter
-new_names <- sapply(current_names[1:6], function(x) {
-  paste0(toupper(substring(x, 1, 1)), tolower(substring(x, 2)))
-})
+#new_names <- sapply(current_names[1:6], function(x) {
+#  paste0(toupper(substring(x, 1, 1)), tolower(substring(x, 2)))
+#})
 
 # Assign the new column names to the first six columns of the dataframe
-colnames(FtWeatherill_swath_2016)[1:6] <- new_names
+#colnames(FtWeatherill_swath_2016)[1:6] <- new_names
 
@@
 # Rename ft. wetherill to fort wetherill
-swath_data$SITE[swath_data$SITE == "Ft.Weatherill"] <- "Fort Wetherill"
-
-#still need to edit ft.wetherill dataframe
+swath_data$Site[swath_data$Site == "Ft.Weatherill"] <- "Fort Wetherill"
 

### 2024-05-17 scripts/data_collation.Rmd
added updated swath data

@@
 morph_data$DATE <- mdy(morph_data$DATE)
+
+# rename moprph category columns in morph and kelp data
+morph_data <- morph_data %>%
+  rename(
+    "BLADE_WIDTH_CM" = "WIDTH_CM")
+
+kelp_data <- kelp_data %>%
+  rename(
+    "BLADE_WIDTH_CM" = "WIDTH_CM",
+    "BLADE_LENGTH_CM" = "LENGTH_CM")
+
+#MAKE STIPE LENGTH NUMERICAL INSTEAD OF CHARACTER in morph data
+morph_data$STIPE_LENGTH_CM <- as.numeric(morph_data$STIPE_LENGTH_CM)
+
+#combine kelp and morph daat
+morph_data <- left_join(morph_data, kelp_data)
 
@@
 
+# output updated dataframe
+write.csv(swath_data, "output/cleaned_data/swath.data.clean.csv", row.names = FALSE)
+
 ```

### 2024-05-23 scripts/data_collation.Rmd
updaated site info with master

@@
 ```{r}
-
-# site information data is cleaned and ready
-
-# Make all columns capitalized
-colnames(Site_info_2018) <- toupper(colnames(Site_info_2018))
-
+#site info cleaned metadta
 # combine site collection data
-site.info.dat <- rbind.fill(Site_info_2017, Site_info_2018, Site_info_template_v1)
-
-# remove columns with observer and unneeded data
-site.info.dat <- site.info.dat[ -c(12:15) ]
+site.info.dat <- kelp_siteinfo_MASTER
 
@@
 write.csv(site.info.dat, "output/cleaned_data/site.info.data.clean.csv", row.names = FALSE)
-
-# Austin - the site info for 2018 has a start and stop latitude so don't know what to do here
 

### 2024-05-23 scripts/data_collation.Rmd
redid swath, cleaned quaad, and species meta

@@
 
-# output updated dataframe
-write.csv(kelp_data, "output/cleaned_data/kelp.data.clean.csv", row.names = FALSE)
 ```
@@
 #combine kelp and morph daat
-morph_data <- left_join(morph_data, kelp_data)
+morph_data <- rbind.fill(morph_data, kelp_data)
 
@@
 
+# Kelp COVER data 
+```{r}
+# kelp cover dataframe has species names as columns in wide format, need to change names to species codes to combine with other dataframes
+# Load species names and species code data frame
+species.metadata <- read.csv("data/species_code_metadata.csv")
+
+# Remove the first column from Kelp_COVER if it's unnecessary
+Kelp_COVER <- Kelp_COVER[ , -1 ]
+
+# Function to rename columns using metadata
+# Get the unique column names from the Kelp_COVER dataframe
+column_names.kelp <- colnames(Kelp_COVER)
+
+# Create a lookup table with name-to-code mappings
+name_to_code.kelp <- setNames(species.metadata$code, species.metadata$name)
+
+# Rename the columns using the lookup table
+new_column_names.kelp <- sapply(column_names.kelp, function(name) {
+  if (name %in% names(name_to_code.kelp)) {
+    return(name_to_code.kelp[name])
+  } else {
+    return(name)
+  }
+})
+
+# Rename the columns in Kelp_COVER
+colnames(Kelp_COVER) <- new_column_names.kelp
+
+# Melt only columns AGH:WHEL into long format
+Kelp_COVER_long <- Kelp_COVER %>%
+  pivot_longer(cols = AGH:CLT, names_to = "SP_CODE", values_to = "value")
+
+# output updated dataframe
+write.csv(Kelp_COVER_long, "output/cleaned_data/kelp.cover.data.clean.csv", row.names = FALSE)
+
+```
+
 # Combine fish dataframes
 ```{r}
-
-# Fish data is cleaned and ready
 
@@
 # combine updated dataframes
-fish_data <- left_join(fish, fish_data)
+fish_data <- rbind.fill(fish, fish_data)
 
@@
 write.csv(fish_data, "output/cleaned_data/fish.data.clean.csv", row.names = FALSE)
-
-# look at fish_Aggregated later - Austin, this is kelp density possibly?
 
@@
 
-# make columns lowercase
-# Get the current column names
-current_names <- colnames(UPC_data)
-
-# Convert the column names to lowercase while preserving the first letter
-new_names <- sapply(current_names[1:13], function(x) {
-  paste0(toupper(substring(x, 1, 1)), tolower(substring(x, 2)))
-})
-
-# Assign the new column names to the first six columns of the dataframe
-colnames(UPC_data)[1:13] <- new_names
-
 # output updated dataframe
@@
 
-# swath dataframe has species names as columns in wide format, need to change names to species codes to combine with other dataframes
-
-# Load species names and species code data frame
-species.metadata <- read.csv("data/species_code_metadata.csv")
-
-# Function to rename columns using metadata
-# Get the unique column names from the swath dataframe
-column_names <- colnames(swath)
-
-# Create a lookup table with name-to-code mappings
-name_to_code <- setNames(species.metadata$code, species.metadata$name)
-
-# Rename the columns using the lookup table
-  new_column_names <- sapply(column_names, function(name) {
-    if (name %in% names(name_to_code)) {
-      return(name_to_code[name])
-    } else {
-      return(name)
-    }
-  })
-
-# Rename the columns
-colnames(swath) <- new_column_names
-
-# Convert data in numeric columns to numeric
-swath.con <- swath %>%
-  mutate(across(AGH:BLOA, as.numeric))
-
-# Melt only columns AGH:WHEL into long format
-swath_long <- swath.con %>%
-  pivot_longer(cols = AGH:BLOA, names_to = "SP_CODE", values_to = "value")
-
-# Need to now move values to binned swatch categories as columns
-# Pivot the dataframe into wide format
-swath_binned <- swath_long %>%
-  pivot_wider(names_from = swath, values_from = value)
-
-# make columns lowercase
-# Get the current column names
-current_names <- colnames(swath_binned)
-
-# Convert the column names to lowercase while preserving the first letter
-new_names <- sapply(current_names[1:6], function(x) {
-  paste0(toupper(substring(x, 1, 1)), tolower(substring(x, 2)))
-})
-
-# Assign the new column names to the first six columns of the dataframe
-colnames(swath_binned)[1:6] <- new_names
-
 # rename swatch category columns in swath count and swath 208 dataframe
@@
 
-swath_count <- swath_count %>%
+# remove columns with observer and unneeded data
+swath_2018 <- swath_2018[ -c(12:15) ]
+
+# Make all columns capitalized
+colnames(swath_2018) <- toupper(colnames(swath_2018))
+
+#rename in and out size classes
+Swath_count <- Swath_count %>%
   rename(
@@
 
-FtWeatherill_swath_2016 <- FtWeatherill_swath_2016 %>%
-  rename(
-    "0-20 IN" = "X0.20.IN",
-    "20-40 IN" = "X20.40.IN",
-    "40-20 OFF" = "X40.20.OFF",
-    "20-0 OFF" = "X20.0.OFF")
+# remove columns with observer and unneeded data
+Swath_count <- Swath_count[ -c(12:15) ]
 
-# Remove four columns in a specific order swath_2018
-swath_2018 <- swath_2018 %>%
-  select(-12:-15)
+# Make all columns capitalized
+colnames(Swath_count) <- toupper(colnames(Swath_count))
 
-# Remove four columns in a specific order swath_count
-swath_count <- swath_count %>%
-  select(-12:-15)
+# remove columns with observer and unneeded data
+FtWeatherill_Swath_2016 <- FtWeatherill_Swath_2016[ -c(9:14) ]
 
-# make columns lowercase
-# Get the current column names
-#current_names <- colnames(FtWeatherill_swath_2016)
+# Renaming values in the 'SIDE' column
+FtWeatherill_Swath_2016 <- FtWeatherill_Swath_2016 %>%
+  mutate(SIDE = case_when(
+    SIDE == "IN 20" ~ "0-20 IN",
+    SIDE == "OFF 40" ~ "40-20 OFF",
+    SIDE == "IN 40" ~ "20-40 IN",
+    SIDE == "OFF 20" ~ "20-0 OFF",
+    TRUE ~ SIDE))  # Keep other values unchanged
 
-# Convert the column names to lowercase while preserving the first letter
-#new_names <- sapply(current_names[1:6], function(x) {
-#  paste0(toupper(substring(x, 1, 1)), tolower(substring(x, 2)))
-#})
-
-# Assign the new column names to the first six columns of the dataframe
-#colnames(FtWeatherill_swath_2016)[1:6] <- new_names
+# Pivoting the data from long to wide format
+FtWeatherill_Swath_2016_wide <- FtWeatherill_Swath_2016 %>%
+  pivot_wider(
+    names_from = SIDE,
+    values_from = COUNT)
 
 # use rbind.fill() method in R to combine data frames with different columns
-swath_data <- rbind.fill(swath_binned, swath_2018, swath_count, FtWeatherill_swath_2016)
+swath_data <- rbind.fill(swath_2018, Swath_count, FtWeatherill_Swath_2016_wide)
 
 # Rename ft. wetherill to fort wetherill
-swath_data$Site[swath_data$Site == "Ft.Weatherill"] <- "Fort Wetherill"
+swath_data$SITE[swath_data$SITE == "Ft.Weatherill"] <- "Fort Wetherill"
+
+# Rename ft. wetherill to fort wetherill
+swath_data$SITE[swath_data$SITE == "Ft Wetherill"] <- "Fort Wetherill"
 
@@
 
-# To do: species codes transform
+#make QUAD data long
+# Function to rename columns using metadata
+# Get the unique column names from the QUAD ataframe
+column_names.quad <- colnames(QUAD)
+
+# Create a lookup table with name-to-code mappings
+name_to_code.quad <- setNames(species.metadata$code, species.metadata$name)
+
+# Rename the columns using the lookup table
+new_column_names.quad <- sapply(column_names.quad, function(name) {
+  if (name %in% names(name_to_code.quad)) {
+    return(name_to_code.quad[name])
+  } else {
+    return(name)
+  }
+})
+
+# Rename the columns in Kelp_COVER
+colnames(QUAD) <- new_column_names.quad
+
+# Melt only columns AGH:WHEL into long format
+QUAD_long <- QUAD %>%
+  pivot_longer(cols = AGH:Oar.Weed.Total, names_to = "SP_CODE", values_to = "COUNT")
 
 # use rbind.fill() method in R to combine data frames with different columns
-QUAD_data <- rbind.fill(QUAD, QUAD_2018, FtWeatherill_QUAD_2017, FtWeatherill_QUAD_2016_TALBENHORIN)
+QUAD_data <- rbind.fill(QUAD_long, QUAD_2018, FtWeatherill_QUAD_2017, FtWeatherill_QUAD_2016_TALBENHORIN)
+
+#remove 131-138 columns for observers not needed information
+QUAD_data <- QUAD_data[ -c(14:17) ]
 
@@
 
+# Rename Kings Beach to King's Beach
+QUAD_data$SITE[QUAD_data$SITE == "Kings Beach"] <- "King's Beach"
+
+# output updated dataframe
+write.csv(QUAD_data, "output/cleaned_data/QUAD.data.clean.csv", row.names = FALSE)
+
 ```

### 2024-05-23 scripts/data_collation.Rmd
added rugosity dataframe clean

@@
 
+# rename swatch category columns in kelp_swath 2023 and kelp swath 2022
+kelp_swath_2022 <- kelp_swath_2022 %>%
+  rename(
+    "0-20 IN" = "X0.20.IN",
+    "20-40 IN" = "X20.40.IN",
+    "40-20 OFF" = "X40.20.OFF",
+    "20-0 OFF" = "X20.0.OFF"
+  )
+
+# remove columns with observer and unneeded data
+kelp_swath_2022 <- kelp_swath_2022[ -c(15:22) ]
+
+# rename KB to King's Beach and FW to Fort Wetherill
+kelp_swath_2022$SITE[kelp_swath_2022$SITE == "FW"] <- "Fort Wetherill"
+kelp_swath_2022$SITE[kelp_swath_2022$SITE == "KB"] <- "King's Beach"
+
+# rename swatch category columns in kelp_swath 2023 and kelp swath 2022
+kelp_swath_2023 <- kelp_swath_2023 %>%
+  rename(
+    "0-20 IN" = "X0.20.IN",
+    "20-40 IN" = "X20.40.IN",
+    "40-20 OFF" = "X40.20.OFF",
+    "20-0 OFF" = "X20.0.OFF"
+  )
+
+# remove columns with observer and unneeded data
+kelp_swath_2023 <- kelp_swath_2023[ -c(12) ]
+
+# rename KB to King's Beach and FW to Fort Wetherill
+kelp_swath_2023$SITE[kelp_swath_2023$SITE == "FW"] <- "Fort Wetherill"
+kelp_swath_2023$SITE[kelp_swath_2023$SITE == "KB"] <- "King's Beach"
+
+# rename swatch category columns in 2019, 2020, and 2021 swath
+swath_2021 <- swath_2021_DEM %>%
+  rename(
+    "0-20 IN" = "X0.20.IN",
+    "20-40 IN" = "X20.40.IN",
+    "40-20 OFF" = "X40.20.OFF",
+    "20-0 OFF" = "X20.0.OFF"
+  )
+
+# rename KB to King's Beach and FW to Fort Wetherill
+swath_2021$SITE[swath_2021$SITE == "Fort Weatherill"] <- "Fort Wetherill"
+swath_2021$SITE[swath_2021$SITE == "Kings Beach"] <- "King's Beach"
+
+# rename swatch category columns in 2019, 2020, and 2021 swath
+swath_2020 <- swath_2020_DEM %>%
+  rename(
+    "0-20 IN" = "X0.20.IN",
+    "20-40 IN" = "X20.40.IN",
+    "40-20 OFF" = "X40.20.OFF",
+    "20-0 OFF" = "X20.0.OFF"
+  )
+
+# rename Fort weatherill to Fort Wetherill
+swath_2020$SITE[swath_2020$SITE == "Fort Weatherill"] <- "Fort Wetherill"
+
+# rename swatch category columns in 2019, 2020, and 2021 swath
+swath_2019 <- swath_2019_DEM %>%
+  rename(
+    "0-20 IN" = "X0.20.IN",
+    "20-40 IN" = "X20.40.IN",
+    "40-20 OFF" = "X40.20.OFF",
+    "20-0 OFF" = "X20.0.OFF"
+  )
+
 # use rbind.fill() method in R to combine data frames with different columns
-swath_data <- rbind.fill(swath_2018, Swath_count, FtWeatherill_Swath_2016_wide)
+swath_data <- rbind.fill(swath_2018, swath_2019, swath_2020, swath_2021, Swath_count, FtWeatherill_Swath_2016_wide, kelp_swath_2022, kelp_swath_2023)
 
@@
 ```
+
+# Combine rugosity dataframes
+```{r}
+
+#make QUAD data long
+# Function to rename columns using metadata
+# Get the unique column names from the QUAD ataframe
+column_names.quad <- colnames(QUAD)
+
+# Create a lookup table with name-to-code mappings
+name_to_code.quad <- setNames(species.metadata$code, species.metadata$name)
+
+# Rename the columns using the lookup table
+new_column_names.quad <- sapply(column_names.quad, function(name) {
+  if (name %in% names(name_to_code.quad)) {
+    return(name_to_code.quad[name])
+  } else {
+    return(name)
+  }
+})
+
+# Rename the columns in Kelp_COVER
+colnames(QUAD) <- new_column_names.quad
+
+# Melt only columns AGH:WHEL into long format
+QUAD_long <- QUAD %>%
+  pivot_longer(cols = AGH:Oar.Weed.Total, names_to = "SP_CODE", values_to = "COUNT")
+
+# use rbind.fill() method in R to combine data frames with different columns
+QUAD_data <- rbind.fill(QUAD_long, QUAD_2018, FtWeatherill_QUAD_2017, FtWeatherill_QUAD_2016_TALBENHORIN)
+
+#remove 131-138 columns for observers not needed information
+QUAD_data <- QUAD_data[ -c(14:17) ]
+
+# Rename ft. wetherill to fort wetherill
+QUAD_data$SITE[QUAD_data$SITE == "Ft Wetherill"] <- "Fort Wetherill"
+
+# Rename Kings Beach to King's Beach
+QUAD_data$SITE[QUAD_data$SITE == "Kings Beach"] <- "King's Beach"
+
+# output updated dataframe
+write.csv(QUAD_data, "output/cleaned_data/QUAD.data.clean.csv", row.names = FALSE)
+
+```

### 2024-05-23 scripts/data_cleaning.Rmd
updated fish data

--- 2024-05-23 scripts/data_collation.Rmd
@@
 ```{r}
-
-#make QUAD data long
-# Function to rename columns using metadata
-# Get the unique column names from the QUAD ataframe
-column_names.quad <- colnames(QUAD)
-
-# Create a lookup table with name-to-code mappings
-name_to_code.quad <- setNames(species.metadata$code, species.metadata$name)
-
-# Rename the columns using the lookup table
-new_column_names.quad <- sapply(column_names.quad, function(name) {
-  if (name %in% names(name_to_code.quad)) {
-    return(name_to_code.quad[name])
-  } else {
-    return(name)
-  }
-})
-
-# Rename the columns in Kelp_COVER
-colnames(QUAD) <- new_column_names.quad
-
-# Melt only columns AGH:WHEL into long format
-QUAD_long <- QUAD %>%
-  pivot_longer(cols = AGH:Oar.Weed.Total, names_to = "SP_CODE", values_to = "COUNT")
+# add a YEAR column for 2021 data
+# Add a new column named "year" with all values set to 2021
+rugosity_2021_DEM <- data.frame(YEAR = 2021, rugosity_2021_DEM)
 
 # use rbind.fill() method in R to combine data frames with different columns
-QUAD_data <- rbind.fill(QUAD_long, QUAD_2018, FtWeatherill_QUAD_2017, FtWeatherill_QUAD_2016_TALBENHORIN)
+rug_data <- rbind.fill(rugosity_2019_DEM, rugosity_2020_DEM, rugosity_2021_DEM)
 
-#remove 131-138 columns for observers not needed information
-QUAD_data <- QUAD_data[ -c(14:17) ]
-
-# Rename ft. wetherill to fort wetherill
-QUAD_data$SITE[QUAD_data$SITE == "Ft Wetherill"] <- "Fort Wetherill"
+#remove 22-25 columns for observers not needed information
+rug_data <- rug_data[ -c(22:25) ]
 
 # Rename Kings Beach to King's Beach
-QUAD_data$SITE[QUAD_data$SITE == "Kings Beach"] <- "King's Beach"
+rug_data$SITE[rug_data$SITE == "Kings Beach"] <- "King's Beach"
+rug_data$SITE[rug_data$SITE == "KB"] <- "King's Beach"
+
+# rename rugosity data columns
+rug_data <- rug_data %>%
+  rename(
+    "0M" = "X0M",
+    "4M" = "X4M",
+    "8M" = "X8M",
+    "12M" = "X12M",
+    "16M" = "X16M",
+    "20M" = "X20M",
+    "24M" = "X24M",
+    "28M" = "X28M",
+    "32M" = "X32M",
+    "36M" = "X36M",
+    "40M" = "X40M", 
+    "DEPTH.0M" = "DEPTH...0M..ft.",
+    "DEPTH.40M" = "DEPTH....40M.",
+    "VIS.M" = "VIS..M."
+    )
 
 # output updated dataframe
-write.csv(QUAD_data, "output/cleaned_data/QUAD.data.clean.csv", row.names = FALSE)
+write.csv(rug_data, "output/cleaned_data/rugosity.data.clean.csv", row.names = FALSE)
 

### 2024-05-23 scripts/data_collation.Rmd
started script for figures

@@
 Kelp_COVER_long <- Kelp_COVER %>%
-  pivot_longer(cols = AGH:CLT, names_to = "SP_CODE", values_to = "value")
+  pivot_longer(cols = AGH:CLT, names_to = "SP_CODE", values_to = "COVER")
 
@@
 # remove columns with observer and unneeded data
-fish_data <- fish_data[ -c(10:14) ]
+fish_data <- fish_data[ -c(10:13) ]
 
@@
 
-# remove columns with VIS
-fish_data <- fish_data[ -c(14) ]
-
 # Use lubridate to format date to be consistent
@@
 
+#combine with fish aggregated data
+fish_data_agg <- rbind.fill(fish_data, fish_Aggregated)
+
+# remove columns with X fromm numerical order of columns
+fish_data_agg <- fish_data_agg[ -c(16) ]
+
+# Rename ft. wetherill to fort wetherill
+fish_data_agg$SITE[fish_data_agg$SITE == "Kings Beach"] <- "King's Beach"
+
+# Rename fort weatherill
+fish_data_agg$SITE[fish_data_agg$SITE == "Fort Weatherill"] <- "Fort Wetherill"
+
+# Rename transcets to not have FW or KB before them
+fish_data_agg$TRANSECT <- gsub("FW\\s*", "", fish_data_agg$TRANSECT)
+fish_data_agg$TRANSECT <- gsub("KB\\s*", "", fish_data_agg$TRANSECT)
+
 # output updated dataframe
-write.csv(fish_data, "output/cleaned_data/fish.data.clean.csv", row.names = FALSE)
+write.csv(fish_data_agg, "output/cleaned_data/fish.data.clean.csv", row.names = FALSE)
 

### 2024-05-24 scripts/manuscript_figures.Rmd
making figure 2

---
title: "manuscript_figures"
author: "daniellembecker"
date: "2024-05-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Use this script to use cleaned data and make plots for RI kelp data manuscript

# Load libraries 
```{r}
library(plyr)
library(dplyr)
library(tidyverse)
library(lubridate)
library(ggplot2)
library(patchwork)
```

# load cleaned dataframes

```{r}
# Set your directory path
directory <- "output/cleaned_data/"

# List files in the directory
files <- list.files(directory)

# Initialize an empty list to store data frames
data_frames <- list()

# Loop through each file
for (file in files) {
    # Read the file into a data frame
    # Adjust the read function according to your file type (e.g., read.csv, read.table)
    # Specify other parameters as needed (e.g., header = TRUE)
    file_path <- file.path(directory, file)
    df_name <- gsub("\\..*", "", file)  # Extract file name without extension
    assign(df_name, read.csv(file_path, stringsAsFactors = FALSE))
}

# Now, each file is a separate data frame with its own variable name

```

# Figure 2: Fish biomass (g/m2) on y-axis and Kelp cover (%) on the x-axis. Point scatter plot with a linear regression and 95% CI bands. For all data points, you will have one value for fish biomass and a corresponding value for kelp cover. So it will be by: year x site x transect.

```{r}
# take kelp cover long and just extract data for sugar kelp 2019-2023
kelp.filt <- kelp %>%
        filter(SP_CODE == "SUGK")

# Convert the integer column to character
kelp.filt$CONTROL <- as.character(kelp.filt$CONTROL)
kelp.filt$TRANSECT <- as.character(kelp.filt$TRANSECT)

# Read the fish aggregated CSV file 2019-2023 and skip the first column
fish <- read.csv("data/kelp_timeseries_data/fish_Aggregated.csv") %>%
          select(-1)

# Remove "KB", "FW", and space, keep only numbers
fish.filt <- fish %>%
  mutate(TRANSECT = gsub("KB |FW ", "", TRANSECT))

# Convert the integer column to character
fish.filt$TRANSECT <- as.character(fish.filt$TRANSECT)

# Rename fort weatherill
fish.filt$SITE[fish.filt$SITE == "Kings Beach"] <- "King's Beach"

# Assuming you want to join by the 'transect' column
merged.dat <- left_join(fish.filt, kelp.filt)

```

### 2024-05-24 scripts/manuscript_figures.Rmd
making figure 2

@@
 library(patchwork)
+library(broom)
+
 ```
@@
 ```
+# Make Figure 2
+
+```{r}
+
+# Aggregate the data by year, site, and transect and compute mean values
+aggregated_data <- merged.dat %>%
+  group_by(year, site, transect) %>%
+  summarize(mean_fish_biomass = mean(fish_biomass, na.rm = TRUE),
+            mean_kelp_cover = mean(kelp_cover, na.rm = TRUE)) %>%
+  ungroup()
+
+# Perform linear regression
+lm_model <- lm(mean_fish_biomass ~ mean_kelp_cover, data = aggregated_data)
+lm_summary <- tidy(lm_model)
+
+# Create scatter plot
+scatter_plot <- ggplot(aggregated_data, aes(x = mean_kelp_cover, y = mean_fish_biomass)) +
+  geom_point() +
+  geom_smooth(method = "lm", se = TRUE, formula = y ~ x, color = "blue") +
+  geom_text(data = lm_summary, aes(label = paste("R² =", round(r.squared, 2), ", p =", signif(p.value, 2))),
+            x = Inf, y = Inf, hjust = 1, vjust = 1, size = 4, parse = TRUE) +
+  labs(x = "Kelp cover (%)", y = "Fish biomass (g/m2)",
+       title = "Fish Biomass vs Kelp Cover",
+       caption = "Linear regression with 95% CI bands") +
+  theme_minimal()
+
+# Display the plot
+print(scatter_plot)
+
+```

### 2024-05-24 scripts/manuscript_figures.Rmd
making figure 2

--- 2024-05-24 scripts/manuscript_figures.Rmd
@@
 ```
+# Make Figure 2
+
+```{r}
+
+# Aggregate the data by year, site, and transect and compute mean values
+aggregated_data <- merged.dat %>%
+  group_by(year, site, transect) %>%
+  summarize(mean_fish_biomass = mean(Fish.Biomass..g.m2., na.rm = TRUE),
+            mean_kelp_cover = mean(COVER, na.rm = TRUE)) %>%
+  ungroup()
+
+# Perform linear regression
+lm_model <- lm(mean_fish_biomass ~ mean_kelp_cover, data = aggregated_data)
+lm_summary <- tidy(lm_model)
+
+# Create scatter plot
+figure.2 <- ggplot(aggregated_data, aes(x = mean_kelp_cover, y = mean_fish_biomass)) +
+  geom_point() +
+  geom_smooth(method = "lm", se = TRUE, formula = y ~ x, color = "blue") +
+  geom_text(data = lm_summary, aes(label = paste("R² =", round(r.squared, 2), ", p =", signif(p.value, 2))),
+            x = Inf, y = Inf, hjust = 1, vjust = 1, size = 4, parse = TRUE) +
+  labs(x = "Kelp cover (%)", y = "Fish biomass (g/m2)",
+       title = "Fish Biomass vs Kelp Cover") +
+  theme_minimal();figure.2
+
+```

### 2024-05-29 scripts/manuscript_figures.Rmd
finished data and figure 2

--- 2024-05-24 scripts/manuscript_figures.Rmd
@@
 
+# remove control column, not needed
+kelp.filt <- kelp.filt[ -c(5) ]
+
 # Read the fish aggregated CSV file 2019-2023 and skip the first column
@@
 
+# remove control column, not needed
+fish.filt <- fish.filt[ -c(5) ]
+
 # Assuming you want to join by the 'transect' column
@@
 ```
+
+# Make Figure 2
+```{r}
+
+# Aggregate the data by year, site, and transect and compute mean values
+aggregated_data <- merged.dat %>%
+  group_by(YEAR, SITE, TRANSECT) %>%
+  summarize(mean_fish_biomass = mean(Fish.Biomass..g.m2., na.rm = TRUE),
+            mean_kelp_cover = mean(COVER, na.rm = TRUE)) %>%
+  ungroup()
+
+# Perform linear regression
+lm_model <- lm(mean_fish_biomass ~ mean_kelp_cover, data = aggregated_data)
+lm_summary <- tidy(lm_model)
+lm_summary_stats <- summary(lm_model)  # Get model summary
+
+# Extract R-squared and p-value
+r_squared <- round(lm_summary_stats$r.squared, 2)
+p_value <- signif(lm_summary$p.value[2], 2)
+
+# Create scatter plot
+figure.2 <- ggplot(aggregated_data, aes(x = mean_kelp_cover, y = mean_fish_biomass)) +
+  geom_point() +
+  geom_smooth(method = "lm", se = TRUE, formula = y ~ x, color = "blue") +
+  annotate("text", x = Inf, y = Inf, label = paste0("R^2 = ", r_squared, ", p = ", p_value), 
+           hjust = 2, vjust = 4, size = 4, parse = FALSE) +
+  labs(x = "Kelp cover (%)", y = "Fish biomass (g/m2)",
+       title = "Fish Biomass vs Kelp Cover") +
+  theme_classic(); figure.2
+
+```

### 2024-05-29 scripts/manuscript_figures.Rmd
finished data and figure 2

@@
 library(dplyr)
+library(MuMIn)
 library(tidyverse)
+library(lme4)
+library(broom.mixed)
 library(lubridate)
@@
 
-# Aggregate the data by year, site, and transect and compute mean values
-aggregated_data <- merged.dat %>%
-  group_by(YEAR, SITE, TRANSECT) %>%
-  summarize(mean_fish_biomass = mean(Fish.Biomass..g.m2., na.rm = TRUE),
-            mean_kelp_cover = mean(COVER, na.rm = TRUE)) %>%
-  ungroup()
+# Fit the linear mixed effects model
+lmm_model <- lmer(mean_fish_biomass ~ mean_kelp_cover + SITE + (1 | SITE/YEAR), data = aggregated_data)
 
-# Perform linear regression
-lm_model <- lm(mean_fish_biomass ~ mean_kelp_cover, data = aggregated_data)
-lm_summary <- tidy(lm_model)
-lm_summary_stats <- summary(lm_model)  # Get model summary
+# Summarize the model
+lmm_summary <- summary(lmm_model)
+lmm_tidy <- tidy(lmm_model)
 
-# Extract R-squared and p-value
-r_squared <- round(lm_summary_stats$r.squared, 2)
-p_value <- signif(lm_summary$p.value[2], 2)
+# Calculate R-squared for the mixed model
+r_squared <- r.squaredGLMM(lmm_model)
+
+# Extract marginal and conditional R-squared
+marginal_r2 <- round(r_squared[1], 2)
+conditional_r2 <- round(r_squared[2], 2)
+p_value <- signif(lmm_tidy$p.value[lmm_tidy$term == "mean_kelp_cover"], 2)
 
@@
   geom_smooth(method = "lm", se = TRUE, formula = y ~ x, color = "blue") +
-  annotate("text", x = Inf, y = Inf, label = paste0("R^2 = ", r_squared, ", p = ", p_value), 
-           hjust = 2, vjust = 4, size = 4, parse = FALSE) +
+  annotate("text", x = Inf, y = Inf, label = paste0("Marginal R^2 = ", marginal_r2, "\nConditional R^2 = ", conditional_r2, "\np = ", p_value), 
+           hjust = 1.1, vjust = 1.1, size = 4, parse = FALSE) +
   labs(x = "Kelp cover (%)", y = "Fish biomass (g/m2)",
        title = "Fish Biomass vs Kelp Cover") +
-  theme_classic(); figure.2
+  theme_classic();figure.2
 
 ```
+
+# Figure 3: Make the saame regression figure for fish biomass and kelp biomass (g/m2).
+To get kelp biomass, you will use the quad data which counts the number of kelp blades, and convert to biomass using the morphometric data.
+Take stipe length, blade width, blade length and convert to biomass.
+Then get a mean and SE of kelp frond biomass at that transect and multiply that by the number of fronds counted in quad.
+We want units to be g/m2 (I think).

### 2024-05-29 scripts/manuscript_figures.Rmd
finished data and figure 2

@@
 library(MuMIn)
+library(lmerTest)
 library(tidyverse)
@@
 
+# Aggregate the data by year, site, and transect and compute mean values
+aggregated_data <- merged.dat %>%
+  group_by(YEAR, SITE, TRANSECT) %>%
+  summarize(mean_fish_biomass = mean(Fish.Biomass..g.m2., na.rm = TRUE),
+            mean_kelp_cover = mean(COVER, na.rm = TRUE)) %>%
+  ungroup()
+
 # Fit the linear mixed effects model
@@
 conditional_r2 <- round(r_squared[2], 2)
-p_value <- signif(lmm_tidy$p.value[lmm_tidy$term == "mean_kelp_cover"], 2)
+
+# Extract p-value for mean_kelp_cover
+p_value_kelp_cover <- summary(lmm_model)$coefficients["mean_kelp_cover", "Pr(>|t|)"]
 
@@
   geom_smooth(method = "lm", se = TRUE, formula = y ~ x, color = "blue") +
-  annotate("text", x = Inf, y = Inf, label = paste0("Marginal R^2 = ", marginal_r2, "\nConditional R^2 = ", conditional_r2, "\np = ", p_value), 
-           hjust = 1.1, vjust = 1.1, size = 4, parse = FALSE) +
+annotate("text", x = Inf, y = Inf, label = paste0("Marginal R^2 = ", marginal_r2, "\nConditional R^2 = ", conditional_r2, "\np = ", signif(p_value_kelp_cover, 3)), hjust = 1.4, vjust = 1.1, size = 4, parse = FALSE) + # Adjust vjust
   labs(x = "Kelp cover (%)", y = "Fish biomass (g/m2)",
@@
 
+# Save the plot to a file
+ggsave("output/figures/figure_2.png", plot = figure.2, width = 8, height = 6)
 ```

### 2024-05-29 scripts/manuscript_figures.Rmd
calculatted kelp biommass using morph and quad data

@@
 We want units to be g/m2 (I think).
+
+```{r}
+# read in quad data with filtered for just kelp
+quad.filt <- QUAD %>%
+        filter(SP_CODE == "SUGK")
+
+# remove control column and two last columns, not needed
+quad.filt <- quad.filt[ -c(13:14) ] 
+quad.filt <- quad.filt[ -c(8) ] 
+quad.filt <- quad.filt[ -c(5:6) ]
+
+# I noticed that 06/30/23 dates were accidentily labeled 06/30/29
+# Replace "29" with "23" in the DATE column
+quad.filt$DATE <- sub("29$", "23", quad.filt$DATE)
+
+# Convert DATE to Date format
+quad.filt$DATE <- as.Date(quad.filt$DATE, format = "%m/%d/%y")
+
+# Convert Date format to "YYYY-MM-DD"
+quad.filt$DATE <- format(quad.filt$DATE, "%Y-%m-%d")
+
+# combine quad data with morph data to calculate kelp biomass
+
+#load in morph data and select for SUGK
+morph.filt <- morph %>%
+        filter(SP_CODE == "SL")
+
+#rename SL to SUGK
+morph.filt$SP_CODE[morph.filt$SP_CODE == "SL"] <- "SUGK"
+
+# Convert the integer column to character
+morph.filt$TRANSECT <- as.character(morph.filt$TRANSECT)
+
+# merge morph filt and quad.filt to calculate biomass
+merge.morph.quad <- left_join(morph.filt, quad.filt)
+
+#notice some dates and data missing, Austin has to look this over!
+
+#calculate kelp biomass
+# To get wet weight in grams (biomass) you use the equation: y=0.00421*x^1.951 where x is kelp blade + stipe length.
+
+# Calculate the kelp blade length
+merge.morph.quad$kelp_blade_length <- merge.morph.quad$BLADE_WIDTH_CM * merge.morph.quad$BLADE_LENGTH_CM
+
+# Calculate the sum of blade length and stipe length
+merge.morph.quad$sum_length <- merge.morph.quad$kelp_blade_length + merge.morph.quad$STIPE_LENGTH_CM
+
+# Define a function to calculate kelp biomass
+calculate_kelp_biomass <- function(x) {
+  y <- 0.00421 * x^1.951
+  return(y)
+}
+
+# Calculate kelp biomass using the sum of blade length and stipe length
+merge.morph.quad$kelp_biomass <- calculate_kelp_biomass(merge.morph.quad$sum_length)
+
+# Remove the intermediate columns if no longer needed
+merge.morph.quad$kelp_blade_length <- NULL
+merge.morph.quad$sum_length <- NULL
+
+```

### 2024-06-03 scripts/data_cleaning.Rmd
pulled temp data from bouys

@@
 
+#edit kings beach
+morph_data$SITE[morph_data$SITE == "Kings Beach"] <- "King's Beach"
+
 # output updated dataframe

### 2024-06-03 scripts/heatwave_script.Rmd
made figure 3

---
title: "RI_heatwave"
author: "daniellembecker"
date: "1/15/2022"
output:
  pdf_document: default
  html_document: default
adapted from: amandachiachi
---

#OISST_data
#Adapted by Amanda Chiachi
#11/13/2019

#### Setup ##########

#load libraries
```{r}
library(dplyr)
library(ggplot2)
library(ncdf4)
library(heatwaveR)# The packages we will use
library(dplyr) # A staple for modern data management in R
library(lubridate) # Useful functions for dealing with dates
library(ggplot2) # The preferred library for data visualisation
library(tidync) # For easily dealing with NetCDF data
library(rerddap) # For easily downloading subsets of data
library(doParallel) # For parallel processing
library(lubridate)
library(scales)
library(maps)

```

# The information for the NOAA OISST data
```{r}
rerddap::info(datasetid = "ncdcOisst21Agg_LonPM180", url = "https://coastwatch.pfeg.noaa.gov/erddap/")
```

# This function expects the user to provide it with a start and end date
# It then downloads and prepares the data
# we are downloading only the SST data out of several variables 
# spatial extent of latitude for ft wetheril
```{r}
# Define the OISST_sub_dl function
OISST_sub_dl <- function(time_df){
  print(time_df)  # Debug print to check the structure of time_df

  # Convert start and end dates to character strings
  start_date <- as.character(time_df$start)
  end_date <- as.character(time_df$end)

  print(paste("Start date:", start_date))
  print(paste("End date:", end_date))

  OISST_dat <- griddap(datasetx = "ncdcOisst21Agg_LonPM180", 
                       url = "https://coastwatch.pfeg.noaa.gov/erddap/", 
                       time = c(start_date, end_date), 
                       zlev = c(0, 0),
                       latitude = c(41.47944444, 41.47944444),
                       longitude = c(-71.35611111, -71.35555556),
                       fields = "sst")$data %>% 
    mutate(time = as.Date(stringr::str_remove(time, "T00:00:00Z"))) %>% 
    dplyr::rename(t = time, temp = sst) %>% 
    select(lon, lat, t, temp) %>% 
    na.omit()

  return(OISST_dat)
}

```

#server doesnt like more than 9 years of consecutive data 
#creates a data frame to download multiple batches of the data 
# allows us to automate the entire download 
# Date download range by start and end dates per year
```{r}
# Define the date ranges for data download
ft_weth_years <- data.frame(date_index = 1,
                            start = as.Date("2016-01-01"),
                            end = as.Date("2023-12-29"))

# Execute the data download and processing
system.time(
  OISST_data <- ft_weth_years %>% 
    group_by(date_index) %>% 
    group_modify(~OISST_sub_dl(.x)) %>% 
    ungroup() %>% 
    select(lon, lat, t, temp)
)

#format date and time, lubridate package format
OISST_data$t <- ymd(OISST_data$t)

#subset temp data if needed

heatwave.temps.OISST <- OISST_data %>%
  filter(lubridate::month(t) %in% c(3:4))

heatwave.temps_filter.OISST <- heatwave.temps.OISST %>%
  filter(lubridate::year(t) %in% c(2016:2024))

write.csv(heatwave.temps_filter.OISST, "output/ft.wetherill.heatwave.temps.OISST.csv")

```

# Detect the events in a time series
# "...must be at least three years to calculate thresholds"
# create your time series data 
```{r}
ts.OISST <- ts2clm(OISST_data, climatologyPeriod = c("2002-01-01", "2021-11-01"))
mhw.OISST <- detect_event(ts.OISST)
```

# View just a few metrics
# The function will return a line plot indicating the climatology, threshold and temperature, with the hot or cold events that meet the specifications of Hobday et al. (2016) shaded in as appropriate.
# An anomalously warm event is considered to be a MHW if it lasts for five or more days, with temperatures warmer than the 90th percentile based on a 30-year historical baseline period.
```{r}
#make data frame with marine heatwave events over time in Mo'orea

mhw.record.OISST <- mhw.OISST$event %>% 
  dplyr::ungroup() %>%
  dplyr::select(event_no, duration, date_start, date_end, date_peak, intensity_max, intensity_cumulative, intensity_mean, rate_onset, rate_decline) %>% 
  dplyr::arrange(-intensity_max); mhw.record.OISST

write.csv(mhw.record.OISST, "RAnalysis/output/heatwave/marine.heatwave.record.OISST.csv")

#format date and time
mhw.OISST$Date<- mdy(mhw.OISST$Date)

#create a line plot of marine heatwave data, can change date range for more specific times
#2019 heatwave scenarios
event_line(mhw.OISST, spread = 50, metric = "intensity_max", 
           start_date = "2019-03-01", end_date = "2019-05-01", x_axis_title = "Date") +
    scale_x_date(date_breaks = "1 day", labels = date_format("%b-%d")) + 
    theme_classic()+
    theme(axis.text.x = element_text(angle = 90))

```

### 2024-06-03 scripts/manuscript_figures.Rmd
made figure 3

@@
 output: html_document
+editor_options: 
+  chunk_output_type: console
 ---
@@
 library(dplyr)
+library(car)
 library(MuMIn)
@@
 
-# Figure 2: Fish biomass (g/m2) on y-axis and Kelp cover (%) on the x-axis.
-Point scatter plot with a linear regression and 95% CI bands.
-For all data points, you will have one value for fish biomass and a corresponding value for kelp cover.
-So it will be by: year x site x transect.
-
+# Figure 2: organize data to make fish biomass (g/m2) on y-axis and Kelp cover (%) figure
 ```{r}
@@
 # Assuming you want to join by the 'transect' column
-merged.dat <- left_join(fish.filt, kelp.filt)
+merged.dat.cover <- left_join(fish.filt, kelp.filt)
 
@@
 
-# Make Figure 2
+# Make Figure 2 and run statistical analysis
 ```{r}
-
 # Aggregate the data by year, site, and transect and compute mean values
-aggregated_data <- merged.dat %>%
+aggregated_data_cover <- merged.dat %>%
   group_by(YEAR, SITE, TRANSECT) %>%
@@
 
+# Filter dataframe for six top fish biomass above 150
+filt_data <- aggregated_data_cover %>%
+  filter(mean_fish_biomass > 150)
+
+# Make biomass a percent so * 100
+aggregated_data_cover$mean_kelp_cover <- aggregated_data_cover$mean_kelp_cover * 100
+
 # Fit the linear mixed effects model
-lmm_model <- lmer(mean_fish_biomass ~ mean_kelp_cover + SITE + (1 | SITE/YEAR), data = aggregated_data)
+lmm_model <- lmer(mean_fish_biomass ~ mean_kelp_cover + SITE + (1 | SITE/YEAR), data = aggregated_data_cover)
+#check normality
+qqPlot(residuals(lmm_model)) #slightly non normal values, log transform to improve
+
+# Fit the linear mixed effects model with log transformation
+lmm_model <- lmer(log(mean_fish_biomass) ~ mean_kelp_cover + SITE + (1 | SITE/YEAR), data = aggregated_data_cover)
+#check normality
+qqPlot(residuals(lmm_model)) #improved to normal
 
@@
 # Create scatter plot
-figure.2 <- ggplot(aggregated_data, aes(x = mean_kelp_cover, y = mean_fish_biomass)) +
+figure.2 <- ggplot(aggregated_data_cover, aes(x = mean_kelp_cover, y = mean_fish_biomass)) +
   geom_point() +
@@
 
-# Figure 3: Make the saame regression figure for fish biomass and kelp biomass (g/m2).
-To get kelp biomass, you will use the quad data which counts the number of kelp blades, and convert to biomass using the morphometric data.
-Take stipe length, blade width, blade length and convert to biomass.
-Then get a mean and SE of kelp frond biomass at that transect and multiply that by the number of fronds counted in quad.
-We want units to be g/m2 (I think).
-
+# Figure 3: organize and calculate kelp biomass data to make fish biomass/kelp biomass (g/m2) linear regression figure
 ```{r}
@@
 
-#notice some dates and data missing, Austin has to look this over!
-
 #calculate kelp biomass
 # To get wet weight in grams (biomass) you use the equation: y=0.00421*x^1.951 where x is kelp blade + stipe length.
-
-# Calculate the kelp blade length
-merge.morph.quad$kelp_blade_length <- merge.morph.quad$BLADE_WIDTH_CM * merge.morph.quad$BLADE_LENGTH_CM
-
 # Calculate the sum of blade length and stipe length
-merge.morph.quad$sum_length <- merge.morph.quad$kelp_blade_length + merge.morph.quad$STIPE_LENGTH_CM
+merge.morph.quad$sum_length <- merge.morph.quad$BLADE_LENGTH_CM + merge.morph.quad$STIPE_LENGTH_CM
 
@@
 
+# Replace NA values with 0
+merge.morph.quad$kelp_biomass <- ifelse(is.na(merge.morph.quad$kelp_biomass), 0, merge.morph.quad$kelp_biomass)
+
+# Sum kelp biomass at the transect level
+morph.filt <- merge.morph.quad %>%
+  group_by(TRANSECT, YEAR, SITE) %>%
+  summarize(total_kelp_biomass = sum(kelp_biomass))
+
+# Divide by 6 quadrats for correct g/m2 calculation
+morph.filt$average_kelp_biomass <- morph.filt$total_kelp_biomass / 6
+
 ```
+
+# Make figure 3 and run statistical analysis
+```{r}
+# Aggregate the data by year, site, and transect and compute mean values
+merged.dat.morph <- left_join(morph.filt, fish.filt)
+
+# Aggregate the data by year, site, and transect and compute mean values
+aggregated.data.morph <- merged.dat.morph %>%
+  group_by(YEAR, SITE, TRANSECT) %>%
+  summarize(mean_fish_biomass = mean(Fish.Biomass..g.m2., na.rm = TRUE),
+            mean_kelp_biomass = mean(average_kelp_biomass, na.rm = TRUE)) %>%
+  ungroup()
+
+# Fit the linear mixed effects model
+lmm_model <- lmer(mean_fish_biomass ~ mean_kelp_biomass + SITE + (1 | SITE/YEAR), data = aggregated.data.morph)
+#check normality
+qqPlot(residuals(lmm_model)) #slightly non normal values, log transform to improve
+
+# Fit the linear mixed effects model with log transformation
+lmm_model <- lmer(log(mean_fish_biomass) ~ mean_kelp_biomass + SITE + (1 | SITE/YEAR), data = aggregated.data.morph)
+#check normality
+qqPlot(residuals(lmm_model)) #improved to normal
+
+# Summarize the model
+lmm_summary <- summary(lmm_model)
+lmm_tidy <- tidy(lmm_model)
+
+# Calculate R-squared for the mixed model
+r_squared <- r.squaredGLMM(lmm_model)
+
+# Extract marginal and conditional R-squared
+marginal_r2 <- round(r_squared[1], 2)
+conditional_r2 <- round(r_squared[2], 2)
+
+# Extract p-value for mean_kelp_cover
+p_value_kelp_cover <- summary(lmm_model)$coefficients["mean_kelp_biomass", "Pr(>|t|)"]
+
+# Create scatter plot
+figure.3 <- ggplot(aggregated.data.morph, aes(x = mean_kelp_biomass, y = mean_fish_biomass)) +
+  geom_point() +
+  geom_smooth(method = "lm", se = TRUE, formula = y ~ x, color = "blue") +
+annotate("text", x = Inf, y = Inf, label = paste0("Marginal R^2 = ", marginal_r2, "\nConditional R^2 = ", conditional_r2, "\np = ", signif(p_value_kelp_cover, 3)), hjust = 1.4, vjust = 1.1, size = 4, parse = FALSE) + # Adjust vjust
+  labs(x = "Kelp Biomass (g/m2)", y = "Fish biomass (g/m2)",
+       title = "Fish Biomass vs Kelp Biomass") +
+  theme_classic();figure.3
+
+# Save the plot to a file
+ggsave("output/figures/figure_3.png", plot = figure.3, width = 8, height = 6)
+```

### 2024-06-03 scripts/manuscript_figures.Rmd
made figure 4

@@
 
+# Merge fish and morph data
+merged.dat.morph <- left_join(morph.filt, fish.filt)
+
 ```
@@
 ```{r}
-# Aggregate the data by year, site, and transect and compute mean values
-merged.dat.morph <- left_join(morph.filt, fish.filt)
-
 # Aggregate the data by year, site, and transect and compute mean values
@@
 # Extract p-value for mean_kelp_cover
-p_value_kelp_cover <- summary(lmm_model)$coefficients["mean_kelp_biomass", "Pr(>|t|)"]
+p_value_kelp_biomass <- summary(lmm_model)$coefficients["mean_kelp_biomass", "Pr(>|t|)"]
 
@@
   geom_smooth(method = "lm", se = TRUE, formula = y ~ x, color = "blue") +
-annotate("text", x = Inf, y = Inf, label = paste0("Marginal R^2 = ", marginal_r2, "\nConditional R^2 = ", conditional_r2, "\np = ", signif(p_value_kelp_cover, 3)), hjust = 1.4, vjust = 1.1, size = 4, parse = FALSE) + # Adjust vjust
+annotate("text", x = Inf, y = Inf, label = paste0("Marginal R^2 = ", marginal_r2, "\nConditional R^2 = ", conditional_r2, "\np = ", signif(p_value_kelp_biomass, 3)), hjust = 1.4, vjust = 1.1, size = 4, parse = FALSE) + # Adjust vjust
   labs(x = "Kelp Biomass (g/m2)", y = "Fish biomass (g/m2)",
@@
 ```
+
+# Figure 4: organize rugosity data and make linear regression figure for fish biomass and rugosity
+```{r}
+# Calculate mean rugosity complexity scores across depths for each transect, grouped by site and year
+mean_rugosity <- rugosity %>%
+  group_by(SITE, YEAR, TRANSECT) %>%
+  summarize(mean_rugosity = mean(c(X0M, X4M, X8M, X12M, X16M, X20M, X24M, X28M, X32M, X36M, X40M), na.rm = TRUE))
+
+# change transect from integer to chr
+mean_rugosity$TRANSECT <- as.character(mean_rugosity$TRANSECT)
+
+# Left jon fish and rugosity data
+merged.dat.rug <- left_join(mean_rugosity, fish.filt)
+
+```
+
+# Make figure 4 and run statistical analysis
+```{r}
+# Aggregate the data by year, site, and transect and compute mean values
+aggregated.data.rug <- merged.dat.rug %>%
+  group_by(YEAR, SITE, TRANSECT) %>%
+  summarize(mean_fish_biomass = mean(Fish.Biomass..g.m2., na.rm = TRUE),
+            mean_rugosity = mean(mean_rugosity, na.rm = TRUE)) %>%
+  ungroup()
+
+# Fit the linear mixed effects model
+lmm_model <- lmer(mean_fish_biomass ~ mean_rugosity + SITE + (1 | SITE/YEAR), data = aggregated.data.rug)
+#check normality
+qqPlot(residuals(lmm_model)) #data is normal no need to transform
+
+# Summarize the model
+lmm_summary <- summary(lmm_model)
+lmm_tidy <- tidy(lmm_model)
+
+# Calculate R-squared for the mixed model
+r_squared <- r.squaredGLMM(lmm_model)
+
+# Extract marginal and conditional R-squared
+marginal_r2 <- round(r_squared[1], 2)
+conditional_r2 <- round(r_squared[2], 2)
+
+# Extract p-value for mean_kelp_cover
+p_value_rugosity <- summary(lmm_model)$coefficients["mean_rugosity", "Pr(>|t|)"]
+
+# Create scatter plot
+figure.4 <- ggplot(aggregated.data.rug, aes(x = mean_rugosity, y = mean_fish_biomass)) +
+  geom_point() +
+  geom_smooth(method = "lm", se = TRUE, formula = y ~ x, color = "blue") +
+  annotate("text", x = Inf, y = Inf, label = paste0("Marginal R^2 = ", marginal_r2, "\nConditional R^2 = ", conditional_r2, "\np = ", signif(p_value_rugosity, 2)), 
+           hjust = 2.1, vjust = 1.1, size = 4, parse = FALSE) +
+  labs(x = "Rugosity Complexity", y = "Fish biomass (g/m2)",
+       title = "Fish Biomass vs Rugosity Complexity") +
+  theme_classic();figure.4
+
+# Save the plot to a file
+ggsave("output/figures/figure_4.png", plot = figure.4, width = 8, height = 6)
+```

### 2024-06-03 scripts/manuscript_figures.Rmd
makeing maps

@@
 library(dplyr)
+library(tigris)
 library(car)
@@
 library(broom.mixed)
+library(ggmap)
 library(lubridate)
@@
 library(broom)
-
+library(ggspatial)
+if ("ggsn" %in% rownames(installed.packages()) == 'FALSE') BiocManager::install("ggsn")
+library(ggsn)
+library(sf)
+library(rnaturalearth)
+library(rnaturalearthdata)
+library(ggspatial)
 ```
@@
 
+```
+
+# Figure 1: organize data and make study site map of a) larger RI and NarrBay with Ft Weatherill and Kings Beach areas in boxes; b) zoom in of the Ft Weatherill box with specific transects as dots; c) zoom in of the Kings Beach box with specific transects as dots.
+```{r}
+# Create a dataframe with site coordinates
+sites <- data.frame(
+  Site = c("Fort Weatherill", "King's Beach"),
+  Latitude = c(41.477, 41.452),  # Adjusted approximate values for the sites
+  Longitude = c(-71.364, -71.326) # Adjusted approximate values for the sites
+)
+
+# Convert sites dataframe to an sf object with CRS 4326
+sites_sf <- st_as_sf(sites, coords = c("Longitude", "Latitude"), crs = 4326)
+
+# Get Rhode Island state boundaries and set CRS to 4326
+ri_state <- st_as_sf(map("state", plot = FALSE, fill = TRUE)) %>%
+  filter(ID == "rhode island")
+
+# Define your rectangle sizes (adjust as needed)
+rect_size1 <- 0.01 # Smaller rectangle
+rect_size2 <- 0.02 # Larger rectangle
+
+# Function to create rectangles around each site
+create_rectangle <- function(site, rect_size) {
+  lon <- st_coordinates(site)[1]
+  lat <- st_coordinates(site)[2]
+  st_as_sf(st_sfc(st_polygon(list(rbind(
+    c(lon - rect_size, lat - rect_size), 
+    c(lon - rect_size, lat + rect_size), 
+    c(lon + rect_size, lat + rect_size), 
+    c(lon + rect_size, lat - rect_size), 
+    c(lon - rect_size, lat - rect_size)
+  )))), crs = st_crs(site))
+}
+
+# Apply the create_rectangle function to each site for both rectangle sizes
+site_rectangles <- lapply(st_geometry(sites_sf), function(site) {
+  rectangle1 <- create_rectangle(site, rect_size1)
+  rectangle2 <- create_rectangle(site, rect_size2)
+  list(rectangle1, rectangle2)
+})
+
+# Combine all rectangles into one sf object
+site_rectangles <- do.call(rbind, unlist(site_rectangles, recursive = FALSE))
+
+# Set CRS for combined rectangles
+st_crs(site_rectangles) <- st_crs(sites_sf)
+
+# Plot
+panelA <- ggplot() +
+  geom_sf(data = ri_state, fill = "lightgray") + # State boundaries
+  geom_sf(data = site_rectangles, color = "red2", fill = NA, size = 1) + # Add rectangles around sites
+  geom_sf(data = sites_sf, color = "blue", size = 3) + # Add site points for reference
+  coord_sf(xlim = c(-72, -71.0), ylim = c(41.1, 42.1), expand = FALSE) + # Zoom level
+  annotation_scale(location = "bl", width_hint = 0.5) +
+  annotation_north_arrow(location = "bl", which_north = "true", 
+                         pad_x = unit(1.1, "in"), pad_y = unit(0.4, "in"),
+                         style = north_arrow_fancy_orienteering) +
+  theme_classic() # Remove axes and background
+
+panelA
 ```

### 2024-06-03 scripts/manuscript_figures.Rmd
maade map panels a and b

@@
 
-# Convert sites dataframe to an sf object with CRS 4326
+# Convert sites dataframe to an sf object
 sites_sf <- st_as_sf(sites, coords = c("Longitude", "Latitude"), crs = 4326)
 
-# Get Rhode Island state boundaries and set CRS to 4326
-ri_state <- st_as_sf(map("state", plot = FALSE, fill = TRUE)) %>%
-  filter(ID == "rhode island")
+# Get Rhode Island state boundaries
+ri_state <- states(cb = TRUE) %>% 
+  filter(STUSPS == "RI") %>%
+  st_transform(crs = 4326)
 
-# Define your rectangle sizes (adjust as needed)
-rect_size1 <- 0.01 # Smaller rectangle
-rect_size2 <- 0.02 # Larger rectangle
+# Define the larger rectangle size (adjust as needed)
+rect_size <- 0.02
 
@@
 
-# Apply the create_rectangle function to each site for both rectangle sizes
+# Apply the create_rectangle function to each site
 site_rectangles <- lapply(st_geometry(sites_sf), function(site) {
-  rectangle1 <- create_rectangle(site, rect_size1)
-  rectangle2 <- create_rectangle(site, rect_size2)
-  list(rectangle1, rectangle2)
+  create_rectangle(site, rect_size)
 })
@@
 # Combine all rectangles into one sf object
-site_rectangles <- do.call(rbind, unlist(site_rectangles, recursive = FALSE))
+site_rectangles <- do.call(rbind, site_rectangles)
+st_crs(site_rectangles) <- st_crs(sites_sf) # Ensure CRS is set for combined rectangles
 
-# Set CRS for combined rectangles
-st_crs(site_rectangles) <- st_crs(sites_sf)
+# Define label data
+labels <- data.frame(
+  name = c("Narragansett\nBay", "Atlantic\nOcean"),
+  lat = c(41.6, 41.3),
+  lon = c(-71.35, -71.2)
+)
 
@@
   geom_sf(data = ri_state, fill = "lightgray") + # State boundaries
-  geom_sf(data = site_rectangles, color = "red2", fill = NA, size = 1) + # Add rectangles around sites
-  geom_sf(data = sites_sf, color = "blue", size = 3) + # Add site points for reference
+  geom_sf(data = site_rectangles, color = "red2", fill = NA, size = 2) + # Add rectangles around sites
+  #geom_sf(data = sites_sf, color = "blue", size = 3) + # Add site points for reference
+  geom_text(data = labels, aes(x = lon, y = lat, label = name), size = 5, color = "black") + # Add labels
   coord_sf(xlim = c(-72, -71.0), ylim = c(41.1, 42.1), expand = FALSE) + # Zoom level
@@
 panelA
+
+# Save the plot to a file
+ggsave("output/figures/figure_1A.png", plot = panelA, width = 8, height = 6)
+
+# Filter transect data by site
+transect_fw <- filter(site, Site == "Fort Wetherill")
+
+# Define label data
+labels <- data.frame(
+  name = c("Fort\nWetherill"),
+  lat = c(41.485),
+  lon = c(-71.368)
+)
+
+# Plot
+panelB <- ggplot() +
+  geom_sf(data = ri_state, fill = "lightgray") + # State boundaries
+  geom_point(data = transect_fw,
+             aes(x = Start.Longitude, y = Start.Latitude),
+            color = "red", size = 3) + # Add start points of transects
+  annotation_scale(location = "bl", width_hint = 0.5) +
+  annotation_north_arrow(location = "bl", which_north = "true", 
+                         pad_x = unit(1.5, "in"), pad_y = unit(0.2, "in"),
+                         style = north_arrow_fancy_orienteering) +
+    geom_text(data = labels, aes(x = lon, y = lat, label = name), size = 5, color = "black") + # Add labels
+  theme_classic() + # Remove axes and background
+  xlim(c(-71.45, -71.3)) + # Set longitude range
+  ylim(c(41.42, 41.5))   # Set latitude range
+
+panelB
+
+# Save the plot to a file
+ggsave("output/figures/figure_1B.png", plot = panelB, width = 8, height = 6)
+
 ```

### 2024-06-06 scripts/manuscript_figures.Rmd
map figure complete

@@
 library(ggspatial)
-if ("ggsn" %in% rownames(installed.packages()) == 'FALSE') BiocManager::install("ggsn")
-library(ggsn)
 library(sf)
@@
 library(ggspatial)
+
 ```
@@
 # Define label data
-labels <- data.frame(
+labels_tot <- data.frame(
   name = c("Narragansett\nBay", "Atlantic\nOcean"),
@@
 
-# Plot
+# Plot panel A zoommed out RI and sites in rectangles
 panelA <- ggplot() +
@@
   geom_sf(data = site_rectangles, color = "red2", fill = NA, size = 2) + # Add rectangles around sites
-  #geom_sf(data = sites_sf, color = "blue", size = 3) + # Add site points for reference
-  geom_text(data = labels, aes(x = lon, y = lat, label = name), size = 5, color = "black") + # Add labels
+  geom_text(data = labels_tot, aes(x = lon, y = lat, label = name), size = 5, color = "black") + # Add labels
   coord_sf(xlim = c(-72, -71.0), ylim = c(41.1, 42.1), expand = FALSE) + # Zoom level
@@
   annotation_north_arrow(location = "bl", which_north = "true", 
-                         pad_x = unit(1.1, "in"), pad_y = unit(0.4, "in"),
+                         pad_x = unit(0.8, "in"), pad_y = unit(0.4, "in"),
                          style = north_arrow_fancy_orienteering) +
+    # Change x and y axes names
+  labs(x = "Longitude", y = "Latitude") +
   theme_classic() # Remove axes and background
@@
 
+# Adjust latitude values by subtracting 0.0013
+transect_fw <- transect_fw %>%
+  mutate(
+    Start.Latitude = Start.Latitude - 0.0013
+  )
+
 # Define label data
@@
   name = c("Fort\nWetherill"),
-  lat = c(41.485),
+  lat = c(41.486),
   lon = c(-71.368)
@@
 
-# Plot
+# Plot panel B for fort wetherill transects
 panelB <- ggplot() +
@@
              aes(x = Start.Longitude, y = Start.Latitude),
-            color = "red", size = 3) + # Add start points of transects
+             color = "black", fill = "red", size = 3, shape = 21) + # Add start points of transects
   annotation_scale(location = "bl", width_hint = 0.5) +
   annotation_north_arrow(location = "bl", which_north = "true", 
-                         pad_x = unit(1.5, "in"), pad_y = unit(0.2, "in"),
+                         pad_x = unit(0.9, "in"), pad_y = unit(0.2, "in"),
                          style = north_arrow_fancy_orienteering) +
     geom_text(data = labels, aes(x = lon, y = lat, label = name), size = 5, color = "black") + # Add labels
+  # Change x and y axes names
+  labs(x = "Longitude", y = "Latitude") +
   theme_classic() + # Remove axes and background
@@
 ggsave("output/figures/figure_1B.png", plot = panelB, width = 8, height = 6)
+
+# Filter transect data by site
+transect_kb <- filter(site, Site == "King's Beach")
+
+# Define label data
+labels_kb <- data.frame(
+  name = c("King's\nBeach"),
+  lat = c(41.463),
+  lon = c(-71.33)
+)
+
+# Adjust latitude values by subtracting 0.0011
+transect_kb <- transect_kb %>%
+  mutate(
+    Start.Latitude = Start.Latitude - 0.0001
+  )
+
+# Plot panel C for kings beach transects
+panelC <- ggplot() +
+  geom_sf(data = ri_state, fill = "lightgray") + # State boundaries
+  geom_point(data = transect_kb,
+             aes(x = Start.Longitude, y = Start.Latitude),
+             color = "black", fill = "red", size = 3, shape = 21) + # Add start points of transects
+  annotation_scale(location = "bl", width_hint = 0.5) +
+  annotation_north_arrow(location = "bl", which_north = "true", 
+                         pad_x = unit(0.8, "in"), pad_y = unit(0.2, "in"),
+                         style = north_arrow_fancy_orienteering) +
+    geom_text(data = labels_kb, aes(x = lon, y = lat, label = name), size = 5, color = "black") + # Add labels
+  # Change x and y axes names
+  labs(x = "Longitude", y = "Latitude") +
+  theme_classic() + # Remove axes and background
+  xlim(c(-71.45, -71.3)) + # Set longitude range
+  ylim(c(41.42, 41.5))   # Set latitude range
+
+panelC
+
+# Save the plot to a file
+ggsave("output/figures/figure_1C.png", plot = panelC, width = 8, height = 6)
+
+```
+
+# Plot all three panels together using patchwork
+```{r}
+
+# Adjust width for individual plots
+panelB <- panelB + theme(plot.margin = margin(r = 0))  # Remove right margin
+panelC <- panelC + theme(plot.margin = margin(l = 0))  # Remove left margin
+
+# Combine the plots with patchwork
+combined_plot <- (panelA | (panelB / panelC)) +
+  plot_annotation(tag_levels = 'A') + 
+  theme(plot.tag = element_text(size = 20, face = "bold"))
+
+# Save the combined plot
+ggsave("output/figures/figure_1.png", combined_plot, width = 10, height = 6, units = "in", dpi = 300)
 

### 2024-06-06 scripts/manuscript_figures.Rmd
updated stats for correlation analysis and made plot

@@
 library(ggplot2)
+library(PerformanceAnalytics)
+library(reshape2)
+library(viridis)
+library(GGally)
+library(corrplot)
 library(patchwork)
@@
 # Aggregate the data by year, site, and transect and compute mean values
-aggregated_data_cover <- merged.dat %>%
+aggregated_data_cover <- merged.dat.cover %>%
   group_by(YEAR, SITE, TRANSECT) %>%
@@
 ```
+
+# Figure 5: Multiple regression (correlation analysis) between fish biomass (g/m2), kelp biomass (g/m2), kelp cover, rugosity, temperature in ggplot 
+# combine all dataframes
+```{r}
+# left join can only combine two at a time
+# combing kelp cover and fish biomass
+master.dat1 <- left_join(kelp.filt, fish.filt)
+
+#combine rugosity and kelp biomass
+master.dat2 <- left_join(mean_rugosity, morph.filt)
+
+# combine for final dataframe
+master.dat <- left_join(master.dat1, master.dat2)
+
+#delete unnecessary columns from the new joined data sheet 
+all.params.data <- master.dat[, -c(1:7, 9:11, 14)]
+
+#rename various columns in data frame
+names(all.params.data)[1] <- "Kelp Cover (%)"
+names(all.params.data)[2] <- "Fish Biomass (g/cm2)"
+names(all.params.data)[3] <- "Rugosity"
+names(all.params.data)[4] <- "Kelp Biomass (g/cm2)"
+
+#reorder columns to show biomass by eaxchother
+all.params.data <- all.params.data[, c(3, 1, 2, 4)]
+
+view(all.params.data)
+
+# Check for missing values
+any_missing <- any(is.na(all.params.data))
+if (any_missing) {
+  # Handle missing values
+  all.params.data <- na.omit(all.params.data)
+}
+
+# Compute the correlation matrix using Spearman correlation
+cormat <- round(cor(all.params.data, method = "spearman"), 4)
+head(cormat)
+
+#melt the correlation matrix means it reassembles data frame to be more effective to complete corr matrix
+#to long format
+melted_cormat <- melt(cormat)
+head(melted_cormat)
+
+#visulaize the correlation matrix in general
+ggplot(data = melted_cormat, aes(x=Var1, y=Var2, fill=value)) + 
+  geom_tile()
+
+# Get lower and upper triangle of the correlation matrix
+#Note that, a correlation matrix has redundant information. We’ll use the functions below to set half of it to NA
+get_lower_tri<-function(cormat){
+  cormat[upper.tri(cormat)] <- NA
+  return(cormat)
+}
+# Get upper triangle of the correlation matrix
+get_upper_tri <- function(cormat){
+  cormat[lower.tri(cormat)]<- NA
+  return(cormat)
+}
+
+#apply upper tri calculation to graphc
+upper_tri <- get_upper_tri(cormat)
+upper_tri
+
+#melt the correlation matrix
+#melt the correlation data and drop the rows with NA values 
+melted_cormat <- melt(upper_tri, na.rm = TRUE)
+
+#heatmap of correlation matrix
+#negative correlations are in purple color and positive correlations in red
+#scale_fill_gradient2 is used with the argument limit = c(-1,1) as correlation coefficients range from -1 to 1
+#coord_fixed() : this function ensures that one unit on the x-axis is the same length as one unit on the y-axis
+ggplot(data = melted_cormat, aes(Var2, Var1, fill = value))+
+  geom_tile(color = "white")+
+  scale_fill_gradient2(low = "midnightblue", high = "firebrick4", mid = "white", 
+                       midpoint = 0, limit = c(-1,1), space = "Lab", 
+                       name="Spearman's\nCorrelation") +
+  theme_minimal()+ 
+  theme(axis.text.x = element_text(angle = 45, vjust = 1, 
+                                   size = 12, hjust = 1))+
+  coord_fixed()
+
+#edit the number of sig figs listed in the value column
+melted_cormat <- melted_cormat %>% mutate_at(vars(starts_with("value")), funs(round(., 2)))
+
+# Create a ggheatmap with basic characteristics, etc. 
+ggheatmap <- ggplot(melted_cormat, aes(Var2, Var1, fill = value))+
+  geom_tile(color = "white")+
+  scale_fill_gradient2(low = "firebrick3", mid = "white", high = "dodgerblue3", 
+                       midpoint = 0, limit = c(-1,1), space = "Lab", 
+                       name="Spearman's\nCorrelation") +
+  theme_minimal()+ # minimal theme
+  theme(axis.text.x = element_text(angle = 45, vjust = 1, 
+                                   size = 12, hjust = 1))+
+  coord_fixed()
+
+# Print the heatmap
+print(ggheatmap)
+
+# Check for normality
+shapiro.test(all.params.data$`Kelp Cover (%)`)
+shapiro.test(all.params.data$`Fish Biomass (g/cm2)`) 
+shapiro.test(all.params.data$Rugosity)
+shapiro.test(all.params.data$`Kelp Biomass (g/cm2)`)
+
+# data is not normal so move forward using aa spearmans rank correlation test
+
+## Calculate p-values
+p_values <- cor.mtest(all.params.data, method = "spearman")$p
+
+## Convert the correlation matrix to a data frame
+cormat_df <- reshape2::melt(cormat)
+
+## Add p-values to the data frame
+cormat_df$p_value <- p_values[cbind(cormat_df$Var1, cormat_df$Var2)]
+
+## Create a new column to display stars based on p-values
+cormat_df$sig <- apply_rules(cormat_df$p_value, rules = c(0.001, 0.01, 0.05), symbols = c("***", "**", "*"))
+
+# Remove repeating rows
+corузыdated_cormat_df <- cormat_df[!duplicated(cormat_df[, c("Var1", "Var2")]), ]
+
+# Replace non-significant p-узыdues with "ns"
+cormat_df$sig[cormat_df$p_value > 0.05] <- "ns"
+
+# Print the modified data
+print(cormat_df)
+
+# Add correlation coefficients to the heatmap
+ggheatmap +
+  geom_text(data = cormat_df, aes(Var1, Var2, label = ifelse(Var1 == Var2, value, 
+                                                             ifelse(sig == "", paste0(value, "\nns"), 
+                                                             ifelse(duplicated(paste0(value, "\n", sig)), "", paste0(value, "\n", sig))))), 
+            color = "black", size = 6) +
+  theme(axis.title.x = element_blank(),
+        axis.title.y = element_blank(),
+        axis.text.x = element_text(size = 18, face="bold", color="black"),
+        axis.text.y = element_text(size = 18, face="bold", color="black"),
+        panel.grid.major = element_blank(),
+        panel.border = element_blank(),
+        panel.background = element_blank(),
+        axis.ticks = element_blank(),
+        legend.justification = c(0.8, 0),
+        legend.title = element_text(size = 18, face="bold", color="black"),
+        legend.text = element_text(size = 20, face="bold", color="black"),
+        legend.position = c(0.48, 0.75),
+        legend.direction = "horizontal") +
+  guides(fill = guide_colorbar(barwidth = 12, barheight = 2,
+                               title.position = "top", title.hjust = 0.5, title.vjust = 1.0))
+
+ggsave(filename = "output/figures/figure_5.png", device = "png", width = 10, height = 10)
+```

### 2024-06-06 scripts/manuscript_figures.Rmd
updated stats for correlation analysis and made plot

@@
 library(tigris)
+library(htmlwidgets)
 library(car)
@@
 library(lubridate)
+library(cowplot)
 library(ggplot2)
@@
 
-# Figure 5: Multiple regression (correlation analysis) between fish biomass (g/m2), kelp biomass (g/m2), kelp cover, rugosity, temperature in ggplot 
+# Figure 5: Correlation analysis between fish biomass (g/m2), kelp biomass (g/m2), kelp cover, rugosity, temperature in ggplot 
 # combine all dataframes
@@
 
-ggsave(filename = "output/figures/figure_5.png", device = "png", width = 10, height = 10)
+ggsave(filename = "output/figures/correlation_analysis_figure.png", device = "png", width = 10, height = 10)
+
 ```
+
+# Figure 5: Multiple regression between fish biomass (g/m2), kelp biomass (g/m2), kelp cover, rugosity, temperature in ggplot 
+# combine all dataframes
+```{r}
+install.packages("devtools")
+devtools::install_github("cardiomoon/ggiraphExtra", force = TRUE)
+library(ggiraphExtra)
+
+# Rename columns to remove spaces and special characters
+names(all.params.data) <- c("Fish_Biomass", "Rugosity", "Kelp_Cover", "Kelp_Biomass", "predictions")
+
+# Fit a multiple regression model
+fit4 <- lm(Fish_Biomass ~ Rugosity * Kelp_Biomass * Kelp_Cover, data = all.params.data)
+
+# Diagnostic plots for the regression model
+par(mfrow=c(2,2))
+plot(fit4)
+
+# Use ggPredict to visualize the multiple regression plot interactively
+interactive_plot <- ggPredict(fit4, interactive = TRUE)
+
+# Save the interactive plot as an HTML file
+htmlwidgets::saveWidget(interactive_plot, file = "output/figures/interactive_regression_plot.html")
+
+```
+
+Coefficients:
+                                                   Estimate Std. Error t value Pr(>|t|)  
+(Intercept)                                      -199.41221   96.09880  -2.075   0.0473 *
+Rugosity                                           98.82780   36.09379   2.738   0.0106 *
+`Kelp Biomass (g/cm2)`                              0.41643    0.18683   2.229   0.0340 *
+`Kelp Cover (%)`                                 -765.82428  612.85640  -1.250   0.2218  
+Rugosity:`Kelp Biomass (g/cm2)`                    -0.12496    0.05199  -2.403   0.0231 *
+Rugosity:`Kelp Cover (%)`                         375.38109  244.24549   1.537   0.1355  
+`Kelp Biomass (g/cm2)`:`Kelp Cover (%)`            -0.24978    0.91866  -0.272   0.7877  
+Rugosity:`Kelp Biomass (g/cm2)`:`Kelp Cover (%)`   -0.05647    0.24991  -0.226   0.8229
+
+# Try another type of multiple regession plot
+
+```{r}
+#fit multiple linear regression model
+model2 <- lm(Fish_Biomass ~ Rugosity * Kelp_Biomass * Kelp_Cover, data = all.params.data)
+
+#view results of model
+summary_text <- capture.output(summary(model2))
+writeLines(summary_text, "output/figures/summary_model.txt")
+
+# Produce added variable plots
+av_plots <- avPlots(model2)
+
+#The x-axis displays a single predictor variable and the y-axis displays the response variable.
+#The blue line shows the association between the predictor variable and the response variable, while holding the value of all other predictor variables constant.
+#The points that are labelled in each plot represent the 2 observations with the largest residuals and the 2 observations with the largest partial leverage.
+```

### 2024-06-12 scripts/manuscript_figures.Rmd
updated figures and models

--- 2024-06-06 scripts/manuscript_figures.Rmd
@@
 library(tigris)
+library(htmlwidgets)
 library(car)
+library(ggiraphExtra)
 library(MuMIn)
@@
 library(ggspatial)
+library(rslp)
 library(sf)
@@
   annotation_north_arrow(location = "bl", which_north = "true", 
-                         pad_x = unit(0.8, "in"), pad_y = unit(0.4, "in"),
+                         pad_x = unit(1.0, "in"), pad_y = unit(0.2, "in"),
                          style = north_arrow_fancy_orienteering) +
@@
   lat = c(41.486),
-  lon = c(-71.368)
+  lon = c(-71.37)
 )
@@
   annotation_north_arrow(location = "bl", which_north = "true", 
-                         pad_x = unit(0.9, "in"), pad_y = unit(0.2, "in"),
+                         pad_x = unit(0.5, "in"), pad_y = unit(0.2, "in"),
                          style = north_arrow_fancy_orienteering) +
@@
   theme_classic() + # Remove axes and background
-  xlim(c(-71.45, -71.3)) + # Set longitude range
-  ylim(c(41.42, 41.5))   # Set latitude range
+  xlim(c(-71.40, -71.35)) + # Set longitude range
+  ylim(c(41.46, 41.5))   # Set latitude range
 
@@
 # Save the plot to a file
-ggsave("output/figures/figure_1B.png", plot = panelB, width = 8, height = 6)
+ggsave("output/figures/figure_1B.png", plot = panelB, width = 8, height = 8)
 
@@
   mutate(
-    Start.Latitude = Start.Latitude - 0.0001
+    Start.Latitude = Start.Latitude - 0.0003
   )
@@
   annotation_north_arrow(location = "bl", which_north = "true", 
-                         pad_x = unit(0.8, "in"), pad_y = unit(0.2, "in"),
+                         pad_x = unit(0.6, "in"), pad_y = unit(0.2, "in"),
                          style = north_arrow_fancy_orienteering) +
@@
   theme_classic() + # Remove axes and background
-  xlim(c(-71.45, -71.3)) + # Set longitude range
-  ylim(c(41.42, 41.5))   # Set latitude range
+  xlim(c(-71.40, -71.3)) + # Set longitude range
+  ylim(c(41.44, 41.5))   # Set latitude range
 
@@
   plot_annotation(tag_levels = 'A') + 
-  theme(plot.tag = element_text(size = 20, face = "bold"))
+  theme(plot.tag = element_text(size = 20, face = "bold"));combined_plot
 
 # Save the combined plot
-ggsave("output/figures/figure_1.png", combined_plot, width = 10, height = 6, units = "in", dpi = 300)
+ggsave("output/figures/figure_1.png", combined_plot, width = 12, height = 8, units = "in", dpi = 300)
 
@@
 # Create scatter plot
-figure.2 <- ggplot(aggregated_data_cover, aes(x = mean_kelp_cover, y = mean_fish_biomass)) +
+figure.2 <- ggplot(aggregated_data_cover, aes(x = log(mean_kelp_cover), y = log(mean_fish_biomass))) +
   geom_point() +
@@
 annotate("text", x = Inf, y = Inf, label = paste0("Marginal R^2 = ", marginal_r2, "\nConditional R^2 = ", conditional_r2, "\np = ", signif(p_value_kelp_cover, 3)), hjust = 1.4, vjust = 1.1, size = 4, parse = FALSE) + # Adjust vjust
-  labs(x = "Kelp cover (%)", y = "Fish biomass (g/m2)",
+  labs(x = "Log Kelp cover (%)", y = "Log Fish biomass (g/m2)",
        title = "Fish Biomass vs Kelp Cover") +
@@
 # Create scatter plot
-figure.3 <- ggplot(aggregated.data.morph, aes(x = mean_kelp_biomass, y = mean_fish_biomass)) +
+figure.3 <- ggplot(aggregated.data.morph, aes(x = log(mean_kelp_biomass), y = log(mean_fish_biomass))) +
   geom_point() +
@@
 annotate("text", x = Inf, y = Inf, label = paste0("Marginal R^2 = ", marginal_r2, "\nConditional R^2 = ", conditional_r2, "\np = ", signif(p_value_kelp_biomass, 3)), hjust = 1.4, vjust = 1.1, size = 4, parse = FALSE) + # Adjust vjust
-  labs(x = "Kelp Biomass (g/m2)", y = "Fish biomass (g/m2)",
+  labs(x = "Log Kelp Biomass (g/m2)", y = "Log Fish biomass (g/m2)",
        title = "Fish Biomass vs Kelp Biomass") +
@@
 
-#delete unnecessary columns from the new joined data sheet 
-all.params.data <- master.dat[, -c(1:7, 9:11, 14)]
+#select necessary columns
+all.params.data <- master.dat %>% select("YEAR", "SITE", "COVER", "Fish.Biomass..g.m2.", "mean_rugosity", "average_kelp_biomass")
 
 #rename various columns in data frame
-names(all.params.data)[1] <- "Kelp Cover (%)"
-names(all.params.data)[2] <- "Fish Biomass (g/cm2)"
-names(all.params.data)[3] <- "Rugosity"
-names(all.params.data)[4] <- "Kelp Biomass (g/cm2)"
+names(all.params.data)[3] <- "Kelp Cover (%)"
+names(all.params.data)[4] <- "Fish Biomass (g/cm2)"
+names(all.params.data)[5] <- "Rugosity"
+names(all.params.data)[6] <- "Kelp Biomass (g/cm2)"
 
-#reorder columns to show biomass by eaxchother
-all.params.data <- all.params.data[, c(3, 1, 2, 4)]
+#reorder columns to show biomass by eachother
+all.params.data <- all.params.data[, c(1,2, 4, 6, 5, 3)]
 
@@
 
+# Define the custom apply_rules function
+apply_rules <- function(p_values, rules = c(0.001, 0.01, 0.05), symbols = c("***", "**", "*")) {
+  result <- rep("", length(p_values))
+  for (i in seq_along(rules)) {
+    result[p_values < rules[i]] <- symbols[i]
+  }
+  return(result)
+}
+
 ## Create a new column to display stars based on p-values
@@
 
-ggsave(filename = "output/figures/figure_5.png", device = "png", width = 10, height = 10)
+ggsave(filename = "output/figures/correlation_analysis_figure.png", device = "png", width = 10, height = 10)
+
 ```
+
+# Figure 5: Multiple regression between fish biomass (g/m2), kelp biomass (g/m2), kelp cover, rugosity, temperature in ggplot 
+# combine all dataframes
+```{r}
+# Rename columns to remove spaces and special characters
+names(all.params.data) <- c("YEAR", "SITE", "Fish_Biomass", "Rugosity", "Kelp_Cover", "Kelp_Biomass")
+
+# was getting an error: qr.default(X, tol = tol, LAPACK = FALSE) when running model below which indicates that there are still problematic values (NA, NaN, or Inf) in the data used for the model fitting
+
+# Check for zero or negative values
+zero_or_negative <- sapply(all.params.data[, c("Fish_Biomass", "Rugosity", "Kelp_Biomass", "Kelp_Cover")], function(x) sum(x <= 0))
+zero_or_negative
+
+# Remove rows with zero or negative values in relevant columns
+all.params.data <- all.params.data[all.params.data$Fish_Biomass > 0 & 
+                                   all.params.data$Rugosity > 0 & 
+                                   all.params.data$Kelp_Biomass > 0 & 
+                                   all.params.data$Kelp_Cover > 0, ]
+
+# Fit a multiple regression model
+fit1 <- lm(log(Fish_Biomass) ~ log(Rugosity) * log(Kelp_Biomass) * log(Kelp_Cover), data = all.params.data)
+
+# Diagnostic plots for the regression model
+par(mfrow=c(2,2))
+plot(fit1)
+
+# Produce added multivariable plots
+av_plots <- avPlots(fit1)
+
+#The x-axis displays a single predictor variable and the y-axis displays the response variable.
+#The blue line shows the association between the predictor variable and the response variable, while holding the value of all other predictor variables constant.
+#The points that are labelled in each plot represent the 2 observations with the largest residuals and the 2 observations with the largest partial leverage.
+
+#view results of model
+summary_text1 <- capture.output(summary(fit1))
+writeLines(summary_text1, "output/figures/summary_original_model.txt")
+
+# Fit a multiple regression model with site and year as interactions
+fit2 <- lm(Fish_Biomass ~ Rugosity * Kelp_Biomass * Kelp_Cover * SITE * YEAR, data = all.params.data)
+
+# Diagnostic plots for the regression model
+par(mfrow=c(2,2))
+plot(fit2)
+
+#view results of model
+summary_text2 <- capture.output(summary(fit2))
+writeLines(summary_text2, "output/figures/summary_model_site_year_fixed.txt")
+
+```
+
+# Try another type of multiple regession plot
+```{r}
+#fit multiple linear regression model WITH SITE AND YEAR
+fit3 <- lm(Fish_Biomass ~ SITE*YEAR, data = all.params.data)
+
+#view results of model
+summary_text3 <- capture.output(summary(fit3))
+writeLines(summary_text3, "output/figures/summary_model_site_year_fixed_fishbiomass.txt")
+
+# Fit the mixed-effects model
+fit4 <- lmer(log(Fish_Biomass) ~ log(Rugosity) * log(Kelp_Biomass) * log(Kelp_Cover) + 
+              (1 | SITE) + (1 | YEAR), data = all.params.data)
+
+#view results of model
+summary_text4 <- capture.output(summary(fit4))
+writeLines(summary_text4, "output/figures/summary_model_site_year_random.txt")
+
+# site and year no significant interaction or significant main effects
+
+# run with just site as random effect
+# Fit the mixed-effects model
+fit5 <- lmer(log(Fish_Biomass) ~ log(Rugosity) * log(Kelp_Biomass) * log(Kelp_Cover) + 
+              (1 | SITE), data = all.params.data)
+
+#view results of model
+summary_text5 <- capture.output(summary(fit5))
+writeLines(summary_text5, "output/figures/summary_model_site_random.txt")
+
+# site and year no significant main effects
+
+# run with just year as random effect
+# Fit the mixed-effects model
+fit6 <- lmer(log(Fish_Biomass) ~ log(Rugosity) * log(Kelp_Biomass) * log(Kelp_Cover) + 
+              (1 | YEAR), data = all.params.data)
+
+#view results of model
+summary_text6 <- capture.output(summary(fit6))
+writeLines(summary_text6, "output/figures/summary_model_year_random.txt")
+
+#year is significantly different for kelp cover
+
+```

### 2024-06-17 scripts/data_cleaning.Rmd
working on filtering QUAD data

@@
 
-# use rbind.fill() method in R to combine data frames with different columns
+)# use rbind.fill() method in R to combine data frames with different columns
 QUAD_data <- rbind.fill(QUAD_long, QUAD_2018, FtWeatherill_QUAD_2017, FtWeatherill_QUAD_2016_TALBENHORIN)
@@
 
+kelp_rug_2022 <- data.frame(YEAR = 2022, kelp_rugosity_2022)
+
+kelp_rug_2023 <- data.frame(YEAR = 2023, kelp_rugosity_2022)
+
 # use rbind.fill() method in R to combine data frames with different columns
-rug_data <- rbind.fill(rugosity_2019_DEM, rugosity_2020_DEM, rugosity_2021_DEM)
+rug_data <- rbind.fill(rugosity_2019_DEM, rugosity_2020_DEM, rugosity_2021_DEM, kelp_rug_2022, kelp_rug_2023)
 
 #remove 22-25 columns for observers not needed information
-rug_data <- rug_data[ -c(22:25) ]
+rug_data <- rug_data[ -c(21:25) ]
 
@@
 rug_data$SITE[rug_data$SITE == "KB"] <- "King's Beach"
+rug_data$SITE[rug_data$SITE == "Fort Weatherill"] <- "Fort Wetherill"
+rug_data$SITE[rug_data$SITE == "Ft. Weatherill"] <- "Fort Wetherill"
+rug_data$TRANSECT[rug_data$TRANSECT == "Control 1"] <- "1"
+rug_data$TRANSECT[rug_data$TRANSECT == "Control 2"] <- "2"
 

### 2024-06-17 scripts/manuscript_figures.Rmd
working on filtering QUAD data

@@
 
+# edited control transcet name
+mean_rugosity$TRANSECT[mean_rugosity$TRANSECT == "Control 2"] <- "2"
+
 # change transect from integer to chr
@@
 
+# Filter quad data for all inverts and use to correlate to fish biomass
+```{r}
+
+```
+
 # Figure 5: Correlation analysis between fish biomass (g/m2), kelp biomass (g/m2), kelp cover, rugosity, temperature in ggplot 
@@
 
-# Compute the correlation matrix using Spearman correlation
-cormat <- round(cor(all.params.data, method = "spearman"), 4)
+# Select the relevant columns
+selected_columns <- all.params.data[, c("Fish Biomass (g/cm2)", "Kelp Biomass (g/cm2)", "Rugosity", "Kelp Cover (%)")]
+
+# Calculate the Spearman correlation matrix
+cormat <- round(cor(selected_columns, method = "spearman", use = "complete.obs"), 4)
 head(cormat)
@@
 ## Calculate p-values
-p_values <- cor.mtest(all.params.data, method = "spearman")$p
+p_values <- cor.mtest(selected_columns, method = "spearman")$p
 
@@
 summary_text1 <- capture.output(summary(fit1))
-writeLines(summary_text1, "output/figures/summary_original_model.txt")
-
-# Fit a multiple regression model with site and year as interactions
-fit2 <- lm(Fish_Biomass ~ Rugosity * Kelp_Biomass * Kelp_Cover * SITE * YEAR, data = all.params.data)
-
-# Diagnostic plots for the regression model
-par(mfrow=c(2,2))
-plot(fit2)
-
-#view results of model
-summary_text2 <- capture.output(summary(fit2))
-writeLines(summary_text2, "output/figures/summary_model_site_year_fixed.txt")
+writeLines(summary_text1, "output/figures/summary_original.model.txt")
 
@@
 #fit multiple linear regression model WITH SITE AND YEAR
-fit3 <- lm(Fish_Biomass ~ SITE*YEAR, data = all.params.data)
+fit2 <- lm(Fish_Biomass ~ SITE*YEAR, data = all.params.data)
+
+#view results of model
+summary_text2 <- capture.output(summary(fit2))
+writeLines(summary_text2, "output/figures/summary_model_site_year_fixed_fishbiomass.txt")
+
+# Fit the mixed-effects model
+fit3 <- lmer(log(Fish_Biomass) ~ log(Rugosity) * log(Kelp_Biomass) * log(Kelp_Cover) + 
+              (1 | SITE) + (1 | YEAR), data = all.params.data)
 
@@
 summary_text3 <- capture.output(summary(fit3))
-writeLines(summary_text3, "output/figures/summary_model_site_year_fixed_fishbiomass.txt")
-
-# Fit the mixed-effects model
-fit4 <- lmer(log(Fish_Biomass) ~ log(Rugosity) * log(Kelp_Biomass) * log(Kelp_Cover) + 
-              (1 | SITE) + (1 | YEAR), data = all.params.data)
-
-#view results of model
-summary_text4 <- capture.output(summary(fit4))
-writeLines(summary_text4, "output/figures/summary_model_site_year_random.txt")
+writeLines(summary_text3, "output/figures/summary_model_site_year_random.txt")
 
@@
 # Fit the mixed-effects model
-fit5 <- lmer(log(Fish_Biomass) ~ log(Rugosity) * log(Kelp_Biomass) * log(Kelp_Cover) + 
+fit4 <- lmer(log(Fish_Biomass) ~ log(Rugosity) * log(Kelp_Biomass) * log(Kelp_Cover) + 
               (1 | SITE), data = all.params.data)
@@
 #view results of model
-summary_text5 <- capture.output(summary(fit5))
-writeLines(summary_text5, "output/figures/summary_model_site_random.txt")
+summary_text4 <- capture.output(summary(fit4))
+writeLines(summary_text4, "output/figures/summary_model_site_random.txt")
 
@@
 # Fit the mixed-effects model
-fit6 <- lmer(log(Fish_Biomass) ~ log(Rugosity) * log(Kelp_Biomass) * log(Kelp_Cover) + 
+fit5 <- lmer(log(Fish_Biomass) ~ log(Rugosity) * log(Kelp_Biomass) * log(Kelp_Cover) + 
               (1 | YEAR), data = all.params.data)
@@
 #view results of model
-summary_text6 <- capture.output(summary(fit6))
-writeLines(summary_text6, "output/figures/summary_model_year_random.txt")
+summary_text5 <- capture.output(summary(fit5))
+writeLines(summary_text5, "output/figures/summary_model_year_random.txt")
 

### 2024-06-17 scripts/data_cleaning.Rmd
output master datafrae

@@
 
+# Remove last two rows with kelp and oar weed totals
+QUAD <- QUAD[ -c(128:130) ]
+
 # Melt only columns AGH:WHEL into long format
 QUAD_long <- QUAD %>%
-  pivot_longer(cols = AGH:Oar.Weed.Total, names_to = "SP_CODE", values_to = "COUNT")
+  pivot_longer(cols = AGH:NP, names_to = "SP_CODE", values_to = "COUNT")
 
-)# use rbind.fill() method in R to combine data frames with different columns
+# Select only the code and type columns from species.metadata
+species_metadata_type <- species.metadata[, c("code", "type")]
+
+# Merge quad_long dataframe with species.metadata to include type information
+quad_long_with_type <- merge(QUAD_long, species.metadata[, c("code", "type")], by.x = "SP_CODE", by.y = "code", all.x = TRUE)
+
+# use rbind.fill() method in R to combine data frames with different columns
 QUAD_data <- rbind.fill(QUAD_long, QUAD_2018, FtWeatherill_QUAD_2017, FtWeatherill_QUAD_2016_TALBENHORIN)

### 2024-06-17 scripts/manuscript_figures.Rmd
output master datafrae

@@
 # Fit the linear mixed effects model
-lmm_model <- lmer(mean_fish_biomass ~ mean_kelp_cover + SITE + (1 | SITE/YEAR), data = aggregated_data_cover)
+lm_model <- lmer(mean_fish_biomass ~ mean_kelp_cover + SITE + (1 | SITE/YEAR), data = aggregated_data_cover)
 #check normality
-qqPlot(residuals(lmm_model)) #slightly non normal values, log transform to improve
+qqPlot(residuals(lm_model)) #slightly non normal values, log transform to improve
 
@@
 ```{r}
+invert.filt <-
 
@@
 master.dat <- left_join(master.dat1, master.dat2)
+
+# output to .csv
+write.csv("output/cleaned_data/mast_output_dataframe.csv")
 
@@
 ```
+
+# Run model on temperature by date between two sites 
+```{r}
+# Add a 'Site' column to each dataframe
+FW_temperature$Site <- "FW"
+KB_temperature$Site <- "KB"
+
+# Combine data frames
+temperature_data <- rbind(FW_temperature, KB_temperature)
+
+# Run normality on the temp data
+qqPlot(temperature_data$Temp, main = "QQ Plot of Temperature Data")
+
+# Data is not normal so log transform
+qqPlot(temperature_data$log(Temp), main = "QQ Plot of Temperature Data")
+
+# Fit linear model
+lm_model <- lm(log(Temp) ~ Date + Site, data = temperature_data)
+
+# Summary of the model
+summary(lm_model)
+
+```

### 2024-06-19 scripts/manuscript_figures.Rmd
made temp analysis and table

@@
 library(plyr)
+library(kableExtra)
+library(gt)
 library(dplyr)
 library(tigris)
+library(MASS)
 library(htmlwidgets)
@@
 library(broom)
+library(knitr)
 library(ggspatial)
@@
 # Read the fish aggregated CSV file 2019-2023 and skip the first column
-fish <- read.csv("data/kelp_timeseries_data/fish_Aggregated.csv") %>%
-          select(-1)
+fish <- read.csv("data/kelp_timeseries_data/fish_Aggregated.csv")
+
+# remove Xcolumn, not needed
+fish <- fish[ -c(1) ]
 
@@
 # Fit the linear mixed effects model with log transformation
-lmm_model <- lmer(log(mean_fish_biomass) ~ mean_kelp_cover + SITE + (1 | SITE/YEAR), data = aggregated_data_cover)
+lmm_model <- lmer(log(mean_fish_biomass + 1.1) ~ log(mean_kelp_cover + 1.1) + SITE + (1 | SITE/YEAR), data = aggregated_data_cover)
 #check normality
@@
 # Create scatter plot
-figure.2 <- ggplot(aggregated_data_cover, aes(x = log(mean_kelp_cover), y = log(mean_fish_biomass))) +
+figure.2 <- ggplot(aggregated_data_cover, aes(x = log(mean_kelp_cover + 1.1), y = log(mean_fish_biomass + 1.1))) +
   geom_point() +
@@
 # Fit the linear mixed effects model with log transformation
-lmm_model <- lmer(log(mean_fish_biomass) ~ mean_kelp_biomass + SITE + (1 | SITE/YEAR), data = aggregated.data.morph)
+lmm_model <- lmer(log(mean_fish_biomass + 1.1) ~ log(mean_kelp_biomass + 1.1) + SITE + (1 | SITE/YEAR), data = aggregated.data.morph)
 #check normality
@@
 # Create scatter plot
-figure.3 <- ggplot(aggregated.data.morph, aes(x = log(mean_kelp_biomass), y = log(mean_fish_biomass))) +
+figure.3 <- ggplot(aggregated.data.morph, aes(x = log(mean_kelp_biomass + 1.1), y = log(mean_fish_biomass + 1.1))) +
   geom_point() +
@@
 
+# Fit the linear mixed effects model with log transformation
+lmm_model <- lmer(log(mean_fish_biomass + 1.1) ~ log(mean_rugosity + 1.1) + SITE + (1 | SITE/YEAR), data = aggregated.data.morph)
+
+#check normality
+qqPlot(residuals(lmm_model)) #improved to normal
+
 # Summarize the model
@@
 # Create scatter plot
-figure.4 <- ggplot(aggregated.data.rug, aes(x = mean_rugosity, y = mean_fish_biomass)) +
+figure.4 <- ggplot(aggregated.data.rug, aes(x = log(mean_rugosity + 1.1), y = log(mean_fish_biomass + 1.1))) +
   geom_point() +
@@
 # output to .csv
-write.csv("output/cleaned_data/mast_output_dataframe.csv")
+write.csv(master.dat, "output/cleaned_data/master_output_dataframe.csv")
 
@@
 
-ggsave(filename = "output/figures/correlation_analysis_figure.png", device = "png", width = 10, height = 10)
+ggsave(filename = "output/figures/figure_5.png", device = "png", width = 10, height = 10)
 
@@
 
-# Figure 5: Multiple regression between fish biomass (g/m2), kelp biomass (g/m2), kelp cover, rugosity, temperature in ggplot 
+# Multiple regression between fish biomass (g/m2), kelp biomass (g/m2), kelp cover, rugosity, temperature in ggplot 
 # combine all dataframes
@@
 # Fit a multiple regression model
-fit1 <- lm(log(Fish_Biomass) ~ log(Rugosity) * log(Kelp_Biomass) * log(Kelp_Cover), data = all.params.data)
+fit1 <- lm(log(Fish_Biomass + 1.1) ~ log(Rugosity + 1.1) * log(Kelp_Biomass + 1.1) * log(Kelp_Cover + 1.1), data = all.params.data)
 
@@
 # Fit the mixed-effects model
-fit4 <- lmer(log(Fish_Biomass) ~ log(Rugosity) * log(Kelp_Biomass) * log(Kelp_Cover) + 
+fit4 <- lmer(log(Fish_Biomass + 1.1) ~ log(Rugosity + 1.1) * log(Kelp_Biomass + 1.1) * log(Kelp_Cover + 1.1) + 
               (1 | SITE), data = all.params.data)
@@
 # Fit the mixed-effects model
-fit5 <- lmer(log(Fish_Biomass) ~ log(Rugosity) * log(Kelp_Biomass) * log(Kelp_Cover) + 
+fit5 <- lmer(log(Fish_Biomass + 1.1) ~ log(Rugosity + 1.1) * log(Kelp_Biomass + 1.1) * log(Kelp_Cover + 1.1) + 
               (1 | YEAR), data = all.params.data)
@@
 
-# Run normality on the temp data
-qqPlot(temperature_data$Temp, main = "QQ Plot of Temperature Data")
+# Check normality for each site
+shapiro.test(temperature_data$Temp[temperature_data$Site == "KB"])
+shapiro.test(temperature_data$Temp[temperature_data$Site == "FW"])
 
-# Data is not normal so log transform
-qqPlot(temperature_data$log(Temp), main = "QQ Plot of Temperature Data")
+# data not normal
+
+# Perform Mann-Whitney U Test for site
+mann_whitney_test <- wilcox.test(Temp ~ Site, data = temperature_data, exact = FALSE)
+print(mann_whitney_test)
+
+# Convert Date column to Date class
+temperature_data$Date <- as.Date(temperature_data$Date, format = "%m/%d/%y")
+
+# Extract Year and Month
+temperature_data <- temperature_data %>%
+  mutate(
+    Year = year(Date),
+    Month = month(Date)
+  )
+
+# Calculate monthly averages
+monthly_averages <- temperature_data %>%
+  group_by(Year, Month, Site) %>%
+  summarize(AverageTemp = mean(Temp), .groups = 'drop')
 
 # Fit linear model
-lm_model <- lm(log(Temp) ~ Date + Site, data = temperature_data)
+model <- lm(log(AverageTemp +1.1) ~ Month + Year + Site, data = monthly_averages)
 
-# Summary of the model
-summary(lm_model)
+#check normality
+qqPlot(residuals(model)) #improved to normal
+
+#not normal
+
+# Remove specific outliers
+cleaned_data <- monthly_averages[-c(95, 96), ]
+
+# Refit the model without outliers
+cleaned_model <- lm(log(AverageTemp + 1.1) ~ Month + Year + Site, data = cleaned_data)
+qqPlot(residuals(cleaned_model))
+summary(cleaned_model)
+
+#month is sig different but year and site are not
+
+# make a table for temperature values (monthly mean plus/minus one standard error) for each site (for supplement)
+
+# Calculate monthly mean and standard error
+summary_table <- temperature_data %>%
+  group_by(Site, Month) %>%
+  summarize(
+    MeanTemp = mean(Temp),
+    StdDev = sd(Temp),
+    N = n(),
+    StdError = StdDev / sqrt(N)
+  ) %>%
+  mutate(
+    MeanPlusSE = MeanTemp + StdError,
+    MeanMinusSE = MeanTemp - StdError
+  ) %>%
+  ungroup() %>%
+  dplyr::select(Site, Month, MeanTemp, MeanPlusSE, MeanMinusSE)
+
+# Mapping month numbers to month names
+month_names <- c("January", "February", "March", "April", "May", "June", 
+                 "July", "August", "September", "October", "November", "December")
+
+# Mapping site abbreviations to full names
+site_names <- c("FW" = "Fort Wetherill", "KB" = "King's Beach")
+
+# Replace numeric month values with month names
+summary_table$Month <- month_names[summary_table$Month]
+
+# Replace site abbreviations with full names
+summary_table$Site <- site_names[summary_table$Site]
+
+# Create a gt table
+table_gt <- gt(summary_table) %>%
+  tab_header(
+    title = "Monthly Mean Temperatures by Site",
+    subtitle = "Summary Table"
+  ) %>%
+  fmt_number(
+    columns = vars(MeanTemp, MeanPlusSE, MeanMinusSE),
+    decimals = 1
+  ) %>%
+  tab_style(
+    style = list(
+      cell_text(color = "black", weight = "bold")
+    ),
+    locations = cells_body()
+  );table_gt
+
+# Save as HTML
+gtsave(table_gt, file = "output/temperature_table.html")
 
 ```
+Temperature between sites is not significantly different
+
+	Wilcoxon rank sum test with continuity correction
+
+data:  Temp by Site
+W = 4273000, p-value = 0.9511
+alternative hypothesis: true location shift is not equal to 0

### 2024-06-20 scripts/manuscript_figures.Rmd
reran all scripts without rugosity

@@
 
-# Extract p-value for mean_kelp_cover
-p_value_kelp_cover <- summary(lmm_model)$coefficients["mean_kelp_cover", "Pr(>|t|)"]
+# Extract the p-value for the kelp biomass coefficient
+p_value_kelp_cover <- lmm_summary$coefficients["log(mean_kelp_cover + 1.1)", "Pr(>|t|)"]
 
@@
   geom_smooth(method = "lm", se = TRUE, formula = y ~ x, color = "blue") +
-annotate("text", x = Inf, y = Inf, label = paste0("Marginal R^2 = ", marginal_r2, "\nConditional R^2 = ", conditional_r2, "\np = ", signif(p_value_kelp_cover, 3)), hjust = 1.4, vjust = 1.1, size = 4, parse = FALSE) + # Adjust vjust
+
+  # R-squared values
+  annotate("text", x = Inf, y = Inf, 
+           label = paste0("Marginal R^2 = ", marginal_r2, "\nConditional R^2 = ", conditional_r2), 
+           hjust = 1.1, vjust = 1.1, size = 4) +
+
+  # p-value
+  annotate("text", x = Inf, y = -Inf, 
+           label = paste0("p = ", format(p_value_kelp_cover, scientific = FALSE, digits = 2)), 
+           hjust = 1.3, vjust = -38.5, size = 4) +
+
   labs(x = "Log Kelp cover (%)", y = "Log Fish biomass (g/m2)",
        title = "Fish Biomass vs Kelp Cover") +
-  theme_classic();figure.2
+  theme_classic() +
+  theme(plot.margin = margin(t = 10, r = 30, b = 10, l = 10, unit = "pt")) +
+  coord_cartesian(clip = 'off');figure.2
 
@@
 
-# I noticed that 06/30/23 dates were accidentily labeled 06/30/29
+# I noticed that 06/30/23 dates were accidentally labeled 06/30/29
 # Replace "29" with "23" in the DATE column
@@
 
+# Identify the outliers
+outliers <- c(45, 37, 25, 33)
+
+# remove outliers
+data_clean <- aggregated.data.morph[-outliers, ]
+
+# rerun model with log and without outliers
+lmm_model <- lmer(log(mean_fish_biomass + 1.1) ~ log(mean_kelp_biomass + 1.1) + SITE + (1 | SITE/YEAR), data = data_clean)
+#check normality
+qqPlot(residuals(lmm_model)) #improved to normal
+
 # Summarize the model
@@
 
-# Extract p-value for mean_kelp_cover
-p_value_kelp_biomass <- summary(lmm_model)$coefficients["mean_kelp_biomass", "Pr(>|t|)"]
+# Extract the p-value for the kelp biomass coefficient
+p_value_kelp_biomass <- lmm_summary$coefficients["log(mean_kelp_biomass + 1.1)", "Pr(>|t|)"]
 
@@
   geom_smooth(method = "lm", se = TRUE, formula = y ~ x, color = "blue") +
-annotate("text", x = Inf, y = Inf, label = paste0("Marginal R^2 = ", marginal_r2, "\nConditional R^2 = ", conditional_r2, "\np = ", signif(p_value_kelp_biomass, 3)), hjust = 1.4, vjust = 1.1, size = 4, parse = FALSE) + # Adjust vjust
+
+  # R-squared values
+  annotate("text", x = Inf, y = Inf, 
+           label = paste0("Marginal R^2 = ", marginal_r2, "\nConditional R^2 = ", conditional_r2), 
+           hjust = 1.1, vjust = 1.1, size = 4) +
+
+  # p-value
+  annotate("text", x = Inf, y = -Inf, 
+           label = paste0("p = ", format(p_value_kelp_biomass, scientific = FALSE, digits = 2)), 
+           hjust = 1.3, vjust = -38.5, size = 4) +
+
   labs(x = "Log Kelp Biomass (g/m2)", y = "Log Fish biomass (g/m2)",
        title = "Fish Biomass vs Kelp Biomass") +
-  theme_classic();figure.3
+  theme_classic() +
+  theme(plot.margin = margin(t = 10, r = 30, b = 10, l = 10, unit = "pt")) +
+  coord_cartesian(clip = 'off');figure.3
 
@@
 # Fit the linear mixed effects model with log transformation
-lmm_model <- lmer(log(mean_fish_biomass + 1.1) ~ log(mean_rugosity + 1.1) + SITE + (1 | SITE/YEAR), data = aggregated.data.morph)
+lmm_model <- lmer(log(mean_fish_biomass + 1.1) ~ log(mean_rugosity + 1.1) + SITE + (1 | SITE/YEAR), data = aggregated.data.rug)
+#check normality
+qqPlot(residuals(lmm_model)) #improved to normal
 
+# Identify the outliers
+outliers <- c(25, 37, 24, 35)
+
+# remove outliers
+data_clean.rug <- aggregated.data.rug[-outliers, ]
+
+# rerun model with log and without outliers
+lmm_model <- lmer(log(mean_fish_biomass + 1.1) ~ log(mean_rugosity + 1.1) + SITE + (1 | SITE/YEAR), data = data_clean.rug)
 #check normality
@@
 
-# Extract p-value for mean_kelp_cover
-p_value_rugosity <- summary(lmm_model)$coefficients["mean_rugosity", "Pr(>|t|)"]
+# Extract the p-value for the rugosity
+p_value_rugosity <- lmm_summary$coefficients["log(mean_rugosity + 1.1)", "Pr(>|t|)"]
 
-# Create scatter plot
+#create scatter plot
 figure.4 <- ggplot(aggregated.data.rug, aes(x = log(mean_rugosity + 1.1), y = log(mean_fish_biomass + 1.1))) +
@@
   geom_smooth(method = "lm", se = TRUE, formula = y ~ x, color = "blue") +
-  annotate("text", x = Inf, y = Inf, label = paste0("Marginal R^2 = ", marginal_r2, "\nConditional R^2 = ", conditional_r2, "\np = ", signif(p_value_rugosity, 2)), 
-           hjust = 2.1, vjust = 1.1, size = 4, parse = FALSE) +
+
+  # R-squared values
+  annotate("text", x = Inf, y = Inf, 
+           label = paste0("Marginal R^2 = ", marginal_r2, "\nConditional R^2 = ", conditional_r2), 
+           hjust = 1.1, vjust = 1.1, size = 4) +
+
+  # p-value
+  annotate("text", x = Inf, y = -Inf, 
+           label = paste0("p = ", format(p_value_rugosity, scientific = FALSE, digits = 2)), 
+           hjust = 1.3, vjust = -38.5, size = 4) +
+
   labs(x = "Rugosity Complexity", y = "Fish biomass (g/m2)",
        title = "Fish Biomass vs Rugosity Complexity") +
-  theme_classic();figure.4
+  theme_classic() +
+  theme(plot.margin = margin(t = 10, r = 30, b = 10, l = 10, unit = "pt")) +
+  coord_cartesian(clip = 'off');figure.4
 
@@
 ```{r}
-invert.filt <-
+#invert.filt <-
 
@@
 #select necessary columns
-all.params.data <- master.dat %>% select("YEAR", "SITE", "COVER", "Fish.Biomass..g.m2.", "mean_rugosity", "average_kelp_biomass")
+all.params.data <- master.dat %>% dplyr::select("YEAR", "SITE", "COVER", "Fish.Biomass..g.m2.", "mean_rugosity", "average_kelp_biomass")
 
@@
 fit1 <- lm(log(Fish_Biomass + 1.1) ~ log(Rugosity + 1.1) * log(Kelp_Biomass + 1.1) * log(Kelp_Cover + 1.1), data = all.params.data)
+summary(fit1)
 
@@
 summary_text1 <- capture.output(summary(fit1))
-writeLines(summary_text1, "output/figures/summary_original.model.txt")
+writeLines(summary_text1, "output/model.result.tables/rugosity.included/summary_original.model.txt")
 
@@
 fit2 <- lm(Fish_Biomass ~ SITE*YEAR, data = all.params.data)
+summary(fit2)
 
@@
 summary_text2 <- capture.output(summary(fit2))
-writeLines(summary_text2, "output/figures/summary_model_site_year_fixed_fishbiomass.txt")
+writeLines(summary_text2, "output/model.result.tables/rugosity.included/summary_model_site_year_fixed_fishbiomass.txt")
 
@@
               (1 | SITE) + (1 | YEAR), data = all.params.data)
+summary(fit3)
 
@@
 summary_text3 <- capture.output(summary(fit3))
-writeLines(summary_text3, "output/figures/summary_model_site_year_random.txt")
+writeLines(summary_text3, "output/model.result.tables/rugosity.included/summary_model_site_year_random.txt")
 
@@
               (1 | SITE), data = all.params.data)
+summary(fit4)
 
@@
 summary_text4 <- capture.output(summary(fit4))
-writeLines(summary_text4, "output/figures/summary_model_site_random.txt")
+writeLines(summary_text4, "output/model.result.tables/rugosity.included/summary_model_site_random.txt")
 
@@
               (1 | YEAR), data = all.params.data)
+summary(fit5)
 
@@
 summary_text5 <- capture.output(summary(fit5))
-writeLines(summary_text5, "output/figures/summary_model_year_random.txt")
+writeLines(summary_text5, "output/model.result.tables/rugosity.included/summary_model_year_random.txt")
 
 #year is significantly different for kelp cover
+
+```
+
+# Run all above models without rugosity
+
+```{r}
+
+# Fit a multiple regression model
+fit6 <- lm(log(Fish_Biomass + 1.1) ~ log(Kelp_Biomass + 1.1) * log(Kelp_Cover + 1.1), data = all.params.data)
+summary(fit6)
+
+#view results of model
+summary_text6 <- capture.output(summary(fit6))
+writeLines(summary_text6, "output/model.result.tables/rugosity.not.included/summary_original.model.txt")
+
+#no sig diff effects
+
+# Fit the mixed-effects model
+fit7 <- lmer(log(Fish_Biomass) ~ log(Kelp_Biomass) * log(Kelp_Cover) + (1 | SITE) + (1 | YEAR), data = all.params.data)
+summary(fit7)
+
+#view results of model
+summary_text7 <- capture.output(summary(fit7))
+writeLines(summary_text7, "output/model.result.tables/rugosity.not.included/summary_model_site_year_random.txt")
+
+# site and year no significant interaction or significant main effects
+
+# run with just site as random effect
+# Fit the mixed-effects model
+fit8 <- lmer(log(Fish_Biomass + 1.1) ~ log(Kelp_Biomass + 1.1) * log(Kelp_Cover + 1.1) + (1 | SITE), data = all.params.data)
+summary(fit8)
+
+#view results of model
+summary_text8 <- capture.output(summary(fit8))
+writeLines(summary_text8, "output/model.result.tables/rugosity.not.included/summary_model_site_random.txt")
+
+# site not sig different
+
+# run with just year as random effect
+# Fit the mixed-effects model
+fit9 <- lmer(log(Fish_Biomass + 1.1) ~  log(Kelp_Biomass + 1.1) * log(Kelp_Cover + 1.1) + (1 | YEAR), data = all.params.data)
+summary(fit9)
+
+#view results of model
+summary_text9 <- capture.output(summary(fit9))
+writeLines(summary_text9, "output/model.result.tables/rugosity.not.included/summary_model_year_random.txt")
+
+#year is not sig different
 
@@
 
-# make a table for temperature values (monthly mean plus/minus one standard error) for each site (for supplement)
+```
 
+# Make a table for temperature values (monthly mean plus/minus one standard error) for each site (for supplement)
+```{r}
 # Calculate monthly mean and standard error

### 2024-06-20 scripts/data_cleaning.Rmd
made temp analysis and table
updated rugosity models

@@
 rug_data$SITE[rug_data$SITE == "Ft. Weatherill"] <- "Fort Wetherill"
-rug_data$TRANSECT[rug_data$TRANSECT == "Control 1"] <- "1"
-rug_data$TRANSECT[rug_data$TRANSECT == "Control 2"] <- "2"
+rug_data$TRANSECT[rug_data$TRANSECT == "Control 1"] <- "5"
+rug_data$TRANSECT[rug_data$TRANSECT == "Control 2"] <- "6"
 

### 2024-06-21 scripts/manuscript_figures.Rmd
updated rugosity models
made proportion change plots

@@
 
-# edited control transcet name
-mean_rugosity$TRANSECT[mean_rugosity$TRANSECT == "Control 2"] <- "2"
-
 # change transect from integer to chr
@@
 alternative hypothesis: true location shift is not equal to 0
+
+Production to biomass ratios (P:B) for each fish species
+```{r}
+
+```
+
+Proportional change per year in kelp cover on y-axis, year on x-axis (by site; one graph for KB and one graph for FW)
+```{r}
+# use kelp filt dataframe that was filtered from quad data for sugar kelp
+# Convert COVER to percent and make new dataaframe
+prop.dat <- kelp.filt %>% mutate(COVER = COVER * 100)
+
+# Function to calculate proportional change and standard error per year
+calculate_proportional_change <- function(data) {
+  data <- data %>%
+    # First, aggregate data by SITE and YEAR
+    group_by(SITE, YEAR) %>%
+    summarise(COVER = mean(COVER, na.rm = TRUE), .groups = "drop") %>%
+
+    # Then calculate proportional change
+    arrange(SITE, YEAR) %>%
+    group_by(SITE) %>%
+    mutate(Proportional_Change = ifelse(lag(COVER) == 0, NA, (COVER - lag(COVER)) / lag(COVER))) %>%
+
+    # Calculate standard error for each site
+    mutate(Proportional_Change_SE = sd(Proportional_Change, na.rm = TRUE) / sqrt(sum(!is.na(Proportional_Change)))) %>%
+    ungroup()
+
+  return(data)
+}
+
+# Calculate proportional change for each site
+prop.dat <- calculate_proportional_change(prop.dat)
+
+# Filter the data for each site
+data_kb <- prop.dat %>% filter(SITE == "King's Beach")
+data_fw <- prop.dat %>% filter(SITE == "Fort Wetherill")
+
+# function to plot
+plot_proportional_change <- function(data, site_name, x_limits = c(-2, 20), y_limits = c(0, 30)) {
+  ggplot(data, aes(x = Proportional_Change, y = COVER)) +
+    geom_point() +
+    geom_errorbarh(aes(xmin = Proportional_Change - Proportional_Change_SE, 
+                       xmax = Proportional_Change + Proportional_Change_SE),
+                  height = 0) +
+    # geom_smooth(method = "lm", se = FALSE, color = "blue") +  # Add trend line
+    geom_vline(xintercept = 0, linetype = "dashed") +
+    geom_text(aes(label = YEAR), vjust = -0.5, size = 3) +
+    scale_x_continuous(limits = x_limits) +
+    scale_y_continuous(limits = y_limits) +
+    labs(title = paste("Proportional Change in Kelp Cover at", site_name),
+         x = expression("Proportional change " ~ yr^-1),
+         y = "Mean Percent Cover (%)") +
+    theme_bw() +
+    theme(plot.title = element_text(hjust = 0.5))
+}
+
+# Assuming data_kb and data_fw are already filtered datasets for each site
+data_kb <- prop.dat %>% filter(SITE == "King's Beach")
+data_fw <- prop.dat %>% filter(SITE == "Fort Wetherill")
+
+# Plot for King's Beach with customized settings
+plot_kb <- plot_proportional_change(data_kb, site_name = "King's Beach", x_limits = c(-1.2, 1.2), y_limits = c(0, 10))
+print(plot_kb)
+
+# Plot for Fort Wetherill with customized settings
+plot_fw <- plot_proportional_change(data_fw, site_name = "Fort Wetherill", x_limits = c(-0.6, 0.6), y_limits = c(7, 12))
+print(plot_fw)
+
+# Save the plots as PNG files
+ggsave("output/figures/Kings_Beach_proportional_change.png", plot = plot_kb, width = 10, height = 6, dpi = 300)
+ggsave("output/figures/Fort_Wetherill_proportional_change.png", plot = plot_fw, width = 10, height = 6, dpi = 300)
+
+```

### 2024-06-25 scripts/manuscript_figures.Rmd
updated and added temp, rate of change, kelp biomass plots

--- 2024-06-20 scripts/manuscript_figures.Rmd
@@
 
-# Make figure 3 and run statistical analysis
+# Make figure 3 fish biomass x kelp biomass and run statistical analysis
 ```{r}
@@
 
-  labs(x = "Log Kelp Biomass (g/m2)", y = "Log Fish biomass (g/m2)",
-       title = "Fish Biomass vs Kelp Biomass") +
+  labs(x = expression("Log Kelp Biomass (g/" ~ m^2 ~ ")"), y = expression("Log Fish Biomass (g/" ~ m^2 ~ ")")) +
   theme_classic() +
   theme(plot.margin = margin(t = 10, r = 30, b = 10, l = 10, unit = "pt")) +
-  coord_cartesian(clip = 'off');figure.3
+  coord_cartesian(clip = 'off') +
+  theme(
+    axis.title.x = element_text(size = 18),
+    axis.title.y = element_text(size = 18),
+    axis.text.x = element_text(size = 18),
+    axis.text.y = element_text(size = 18));figure.3
 
@@
 
-# Figure 4: organize rugosity data and make linear regression figure for fish biomass and rugosity
+# Make kelp biomass time series (a) 
+```{r}
+# Function to calculate standard error
+std_error <- function(x) {
+  sd(x, na.rm = TRUE) / sqrt(length(na.omit(x)))
+}
+
+# Aggregate the data by year and site and compute mean values and standard error
+aggregated.kelp.morph <- morph.filt %>%
+  group_by(YEAR, SITE) %>%
+  summarize(
+    mean_kelp_biomass = mean(average_kelp_biomass, na.rm = TRUE),
+    se_kelp_biomass = std_error(average_kelp_biomass)
+  ) %>%
+  ungroup()
+
+# Transform mean kelp biomass and standard error to log scale
+aggregated.kelp.morph <- aggregated.kelp.morph %>%
+  mutate(
+    log_mean_kelp_biomass = log(mean_kelp_biomass + 1.1),
+    log_se_kelp_biomass = log(mean_kelp_biomass + 1.1 + se_kelp_biomass) - log(mean_kelp_biomass + 1.1)
+  )
+
+# Define shapes for sites
+shape_values <- c("Fort Wetherill" = 16, "King's Beach" = 17) # Circle for Fort Wetherill, Triangle for King's Beach
+
+# Time series plot for kelp biomass with standard error bars from 2016-2023
+kelp.timeseries.2016 <- ggplot(aggregated.kelp.morph, aes(x = YEAR, y = log_mean_kelp_biomass, shape = SITE)) +
+  geom_line(aes(group = SITE), color = "black") +
+  geom_point(size = 3) +
+  geom_errorbar(aes(ymin = log_mean_kelp_biomass - log_se_kelp_biomass, 
+                    ymax = log_mean_kelp_biomass + log_se_kelp_biomass), 
+                width = 0.2) +
+  scale_shape_manual(values = shape_values) +
+  labs(x = "Year", y = expression("Log Mean Kelp Biomass (g/" ~ m^2 ~ ")"), shape = "Site") +
+  theme_bw() +
+  scale_x_continuous(breaks = unique(aggregated.kelp.morph$YEAR)) +
+  theme(
+    axis.title.x = element_text(size = 18),
+    axis.title.y = element_text(size = 18),
+    axis.text.x = element_text(size = 18),
+    axis.text.y = element_text(size = 18),
+    legend.text = element_text(size = 16),
+    legend.title = element_text(size = 18)
+  );kelp.timeseries.2016
+
+# Save the plot to a file
+ggsave("output/figures/kelp_timeseries_2016.png", plot = kelp.timeseries.2016, width = 8, height = 6)
+
+# Filter for the years 2019-2023
+filtered_data <- aggregated.kelp.morph %>%
+  filter(YEAR >= 2019 & YEAR <= 2023)
+
+# Time series plot for kelp biomass with standard error bars for 2019-2023
+kelp.timeseries.2019_2023 <- ggplot(filtered_data, aes(x = YEAR, y = log_mean_kelp_biomass, shape = SITE)) +
+  geom_line(aes(group = SITE), color = "black") +
+  geom_point(size = 3) +
+  geom_errorbar(aes(ymin = log_mean_kelp_biomass - log_se_kelp_biomass, 
+                    ymax = log_mean_kelp_biomass + log_se_kelp_biomass), 
+                width = 0.2) +
+  scale_shape_manual(values = shape_values) +
+  labs(x = "Year", y = expression("Log Mean Kelp Biomass (g/" ~ m^2 ~ ")"), shape = "Site") +
+  theme_bw() +
+  theme(
+    axis.title.x = element_text(size = 18),
+    axis.title.y = element_text(size = 18),
+    axis.text.x = element_text(size = 18),
+    axis.text.y = element_text(size = 18),
+    legend.text = element_text(size = 16),
+    legend.title = element_text(size = 18)
+  );kelp.timeseries.2019_2023
+
+# Save the plot to a file
+ggsave("output/figures/kelp_timeseries_2019.png", plot = kelp.timeseries.2019_2023, width = 8, height = 6)
+
+```
+
+# Rate of change (b) by site figure
+```{r}
+# Fit the linear model
+lm_model <- lm(average_kelp_biomass ~ YEAR, data = morph.filt)
+#check normality
+qqPlot(residuals(lm_model)) #data is normal no need to transform
+
+# Fit log transformed linear model
+lm_model <- lm(log(average_kelp_biomass + 1.1) ~ YEAR, data = morph.filt)
+#check normality
+qqPlot(residuals(lm_model)) #data is normal no need to transform
+
+# Convert YEAR to numeric (if necessary)
+morph.filt$YEAR <- as.numeric(as.character(morph.filt$YEAR))
+
+# Define shapes for sites
+shape_values <- c("Fort Wetherill" = 16, "King's Beach" = 17)  # Circle for Fort Wetherill, Triangle for King's Beach
+
+# Fit linear regression models for each SITE separately
+slope_results <- morph.filt %>%
+  mutate(log_avg_kelp_biomass = log(average_kelp_biomass + 1.1)) %>%
+  group_by(SITE) %>%
+  summarize(
+    mean_slope = coef(lm(log_avg_kelp_biomass ~ YEAR))[2],  # Mean slope
+    se_slope = summary(lm(log_avg_kelp_biomass ~ YEAR))$coefficients[2, "Std. Error"],  # Standard error of slope
+    mean_log_avg_kelp_biomass = mean(log_avg_kelp_biomass)  # Mean log average kelp biomass
+  )
+
+# Calculate rate of change per year
+slope_results <- slope_results %>%
+  mutate(
+    rate_of_change_per_year = mean_slope / mean_log_avg_kelp_biomass,  # Rate of change per year
+    se_rate_of_change = se_slope / mean_log_avg_kelp_biomass  # Standard error of rate of change per year
+  )
+
+# Plotting the mean rates of change with error bars (standard error horizontally)
+rate_of_change <- ggplot(slope_results, aes(x = rate_of_change_per_year, y = mean_log_avg_kelp_biomass, shape = SITE)) +
+  geom_point(size = 3) +
+  geom_errorbarh(aes(xmin = rate_of_change_per_year - se_rate_of_change, xmax = rate_of_change_per_year + se_rate_of_change), height = 0.2) +
+  labs(
+    x = expression("Rate of Change ("~yr^-1 ~")"),
+    y = expression("Log Mean Kelp Biomass (g/"~m^2~")"), shape = "Site") +
+  theme_bw() +
+  scale_shape_manual(values = shape_values, guide = guide_legend(title = "Site")) +   # Use shape values and customize legend
+  theme(
+    axis.title.x = element_text(size = 18),
+    axis.title.y = element_text(size = 18),
+    axis.text.x = element_text(size = 18),
+    axis.text.y = element_text(size = 18),
+    legend.text = element_text(size = 16),
+    legend.title = element_text(size = 18));rate_of_change
+
+```
+
+# Plot two panels together using patchwork
+```{r}
+# Combine the plots with patchwork for all years of kelp and rate of change
+kelp.rate.2016 <- kelp.timeseries.2016 | rate_of_change +
+  plot_annotation(tag_levels = 'A') + 
+  theme(plot.tag = element_text(size = 20, face = "bold"));kelp.rate.2016
+
+# Save the combined plot
+ggsave("output/figures/kelp.2016.rate.png", kelp.rate.2016, width = 20, height = 8, units = "in", dpi = 300)
+
+# Combine the plots with patchwork for all years of kelp and rate of change
+kelp.rate.2019 <- kelp.timeseries.2019_2023 | rate_of_change +
+  plot_annotation(tag_levels = 'A') + 
+  theme(plot.tag = element_text(size = 20, face = "bold"));kelp.rate.2019
+
+# Save the combined plot
+ggsave("output/figures/kelp.2019.rate.png", kelp.rate.2019, width = 20, height = 8, units = "in", dpi = 300)
+
+```
+
+# Organize rugosity data and make linear regression figure for fish biomass and rugosity
 ```{r}
@@
   summarize(mean_rugosity = mean(c(X0M, X4M, X8M, X12M, X16M, X20M, X24M, X28M, X32M, X36M, X40M), na.rm = TRUE))
-
-# edited control transcet name
-mean_rugosity$TRANSECT[mean_rugosity$TRANSECT == "Control 2"] <- "2"
 
@@
 # Run all above models without rugosity
-
 ```{r}
@@
 
+```
+
+# Make line plot for temperature time series by site per month
+```{r}
+# Calculate average temperature per month and site
+avg_temp <- temperature_data %>%
+  mutate(MonthAbbrev = month.abb[Month]) %>%  # Add month abbreviations
+  group_by(MonthAbbrev, Site) %>%
+  summarize(avg_temp = mean(Temp))
+
+# Map site codes to full names
+avg_temp$Site <- ifelse(avg_temp$Site == "FW", "Fort Wetherill", "King's Beach")
+
+# Plotting average temperature by site as lines and points per month
+temperature_plot <- ggplot(avg_temp, aes(x = MonthAbbrev, y = avg_temp, color = Site, group = Site)) +
+  geom_line(size = 1.5) +  # Lines for average temperature by site
+  geom_point(size = 3) +  # Points for mean values by site
+  labs(x = "Month", y = expression("Average Temperature (°C)"), color = "Site") +
+  scale_color_manual(values = c("Fort Wetherill" = "grey", "King's Beach" = "black")) +  # Custom colors
+  theme_bw()+
+    theme(
+    axis.title.x = element_text(size = 18),
+    axis.title.y = element_text(size = 18),
+    axis.text.x = element_text(size = 18),
+    axis.text.y = element_text(size = 18),
+    legend.text = element_text(size = 16),
+    legend.title = element_text(size = 18));temperature_plot
+
+# Save the plot to a file
+ggsave("output/figures/temp.plot.png", plot = temperature_plot, width = 20, height = 10)
 ```
@@
 alternative hypothesis: true location shift is not equal to 0
+
+Production to biomass ratios (P:B) for each fish species
+```{r}
+
+```

### 2024-06-25 scripts/manuscript_figures.Rmd
updated and added temp, rate of change, kelp biomass plots

@@
 # Identify the outliers
-outliers <- c(45, 37, 25, 33)
+outliers <- c(45, 37, 25, 33, 37, 22)
 
@@
 # rerun model with log and without outliers
-lmm_model <- lmer(log(mean_fish_biomass + 1.1) ~ log(mean_kelp_biomass + 1.1) + SITE + (1 | SITE/YEAR), data = data_clean)
+lmm_model <- lmer(log(mean_fish_biomass + 1.1) ~ log(mean_kelp_biomass + 1.1) * YEAR + (1 | SITE), data = data_clean)
 #check normality
@@
 lmm_tidy <- tidy(lmm_model)
+
+# Export to CSV file
+write.csv(lmm_tidy, "output/lmm_model_results.csv", row.names = FALSE)
 
@@
 
+# site not sig different
+```
+
+# Run model for fish_biomass ~ kelp_biomass * year (randomize site)
+```{r}
+# Fit the mixed-effects model
+fit10 <- lmer(log(Fish_Biomass + 1.1) ~ log(Kelp_Biomass + 1.1) * YEAR + (1 | SITE), data = all.params.data)
+summary(fit10)
+
+# Example: Standardizing numeric variables in your dataset
+all.params.data$Kelp_Biomass_standardized <- scale(all.params.data$Kelp_Biomass)
+all.params.data$YEAR_standardized <- scale(all.params.data$YEAR)
+
+# Fit the model with standardized predictors
+fit10 <- lmer(log(Fish_Biomass + 1.1) ~ log(Kelp_Biomass_standardized + 1.1) * YEAR_standardized + (1 | SITE), data = all.params.data)
+summary(fit10)
+
+#view results of model
+summary_text10 <- capture.output(summary(fit10))
+writeLines(summary_text9, "output/model.result.tables/rugosity.not.included/summary_model_year_random.txt")
+
 ```

### 2024-06-26 scripts/manuscript_figures.Rmd
calcualating pb and hpi ratios

@@
 library(kableExtra)
+library(flextable)
 library(gt)
@@
 
+# Identify the outliers
+outliers <- c(27,7)
+
+# remove outliers
+data_clean <- aggregated_data_cover[-outliers, ]
+
+# rerun model with log and without outliers
+lmm_model <- lmer(log(mean_fish_biomass + 1.1) ~ log(mean_kelp_cover + 1.1) + YEAR + (1 | SITE), data = data_clean)
+qqPlot(residuals(lmm_model)) #improved to normal
+summary(lmm_model)
+
+# Extract coefficients and their standard errors
+model_summary <- tidy(lmm_model)
+
+# Create a data frame with coefficients, standard errors, and p-values
+df <- data.frame(
+  Coefficient = model_summary$term,
+  Estimate = model_summary$estimate,
+  Std.Error = model_summary$std.error,
+  P.Value = model_summary$p.value
+)
+
+# Create flextable object
+flex <- flextable(df)
+
+# Save flextable as PNG
+png_file <- "output/model.result.tables/rugosity.not.included/summary_fish_kelp_cover_year_fixed.png"
+save_as_image(flex, png_file)
+
+# rerun model with log and without outliers
+lmm_model <- lmer(log(mean_fish_biomass + 1.1) ~ log(mean_kelp_cover + 1.1) * YEAR + (1 | SITE), data = data_clean)
+qqPlot(residuals(lmm_model)) #improved to normal
+summary(lmm_model)
+
+# Extract coefficients and their standard errors
+model_summary <- tidy(lmm_model)
+
+# Create a data frame with coefficients, standard errors, and p-values
+df <- data.frame(
+  Coefficient = model_summary$term,
+  Estimate = model_summary$estimate,
+  Std.Error = model_summary$std.error,
+  P.Value = model_summary$p.value
+)
+
+# Create flextable object
+flex <- flextable(df)
+
+# Save flextable as PNG
+png_file <- "output/model.result.tables/rugosity.not.included/summary_fish_kelp_cover_year_interaction.png"
+save_as_image(flex, png_file)
+
+# rerun model with log and without outliers
+lmm_model <- lmer(log(mean_fish_biomass + 1.1) ~ log(mean_kelp_cover + 1.1) + SITE + (1 | SITE/YEAR), data = data_clean)
+qqPlot(residuals(lmm_model)) #improved to normal
+summary(lmm_model)
+
+# Extract coefficients and their standard errors
+model_summary <- tidy(lmm_model)
+
+# Create a data frame with coefficients, standard errors, and p-values
+df <- data.frame(
+  Coefficient = model_summary$term,
+  Estimate = model_summary$estimate,
+  Std.Error = model_summary$std.error,
+  P.Value = model_summary$p.value
+)
+
+# Create flextable object
+flex <- flextable(df)
+
+# Save flextable as PNG
+png_file <- "output/model.result.tables/rugosity.not.included/summary_fish_kelp_cover.png"
+save_as_image(flex, png_file)
+
 # Summarize the model
@@
 ```{r}
-# Calculate average temperature per month and site
+# Calculate mean and standard error of temperature by month and site
 avg_temp <- temperature_data %>%
@@
   group_by(MonthAbbrev, Site) %>%
-  summarize(avg_temp = mean(Temp))
+  summarize(
+    avg_temp = mean(Temp),
+    std_error = sd(Temp) / sqrt(n()))  # Calculate standard error
 
@@
 
-# Plotting average temperature by site as lines and points per month
+# Plotting average temperature by site as lines and points per month with error bars
 temperature_plot <- ggplot(avg_temp, aes(x = MonthAbbrev, y = avg_temp, color = Site, group = Site)) +
@@
   geom_point(size = 3) +  # Points for mean values by site
+  geom_errorbar(aes(ymin = avg_temp - std_error, ymax = avg_temp + std_error), width = 0.2) +  # Error bars for SEM
   labs(x = "Month", y = expression("Average Temperature (°C)"), color = "Site") +
   scale_color_manual(values = c("Fort Wetherill" = "grey", "King's Beach" = "black")) +  # Custom colors
-  theme_bw()+
+  theme_bw() +
   theme(
@@
 ```{r}
-
-```
+# load fish data from 2016
+PB <- read.csv("data/fish.data.clean.calc.csv")
+
+# remove habitat and control, month, day, date, not needed
+PB <- PB[ -c(1:2, 4:7) ]
+
+# Calculate fish biomass 'W' using the length-weight relationship
+PB$biomass <- with(PB, coef.a * (actual_size)^coef.b)
+
+# Divide biomass by 80 m to get g/m2
+#calculate biomass/m2
+PB$biomass.g.m2 <- PB$biomass / 80
+
+# Constants from the P/B ratio equation
+coefficient <- 2.64
+exponent <- -0.35
+
+# Calculate P/B ratio
+PB$P_B_ratio <- coefficient * PB$biomass.g.m2^exponent
+
+# Calculate Habitat Productivity Index (HPI)
+PB$HPI <- PB$biomass.g.m2 * PB$P_B_ratio
+
+# Aggregate by Transect, Site, Year and calculate mean HPI
+hpi_summary <- PB %>%
+  group_by(TRANSECT, SITE, YEAR) %>%
+  summarise(mean_HPI = mean(HPI))
+
+ggplot(hpi_summary, aes(x = paste(TRANSECT, SITE, sep = " - "), y = mean_HPI, fill = as.factor(YEAR))) +
+  geom_bar(stat = "identity", position = "dodge", color = "black") +
+  labs(x = "Transect - Site", y = "Mean Habitat Productivity Index (HPI)") +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
+```

### 2024-06-27 scripts/manuscript_figures.Rmd
updated HPI

@@
 
-# Calculate P/B ratio
+# Calculate P/B ratio, the predictive equation is P/B (per year) = 2.64*Wmat^-035 , where Wmat is weight-at-maturity (Randall and Minns 2000)
 PB$P_B_ratio <- coefficient * PB$biomass.g.m2^exponent
 
-# Calculate Habitat Productivity Index (HPI)
+# Calculate Habitat Productivity Index (HPI), habitat productivity index (HPI), a proposed measure of habitat suitability and productive capacity, can be calculated as the product of average fish biomass and the species P/B coefficient (HPI = B × P/B) (Randall and Minns 2000)
 PB$HPI <- PB$biomass.g.m2 * PB$P_B_ratio
@@
   group_by(TRANSECT, SITE, YEAR) %>%
-  summarise(mean_HPI = mean(HPI))
-
-ggplot(hpi_summary, aes(x = paste(TRANSECT, SITE, sep = " - "), y = mean_HPI, fill = as.factor(YEAR))) +
+  summarise(mean_HPI = mean(HPI, na.rm = TRUE))
+
+# Make faceted line plot of HPI index over the years
+hpi.plot < - ggplot(hpi_summary, aes(x = interaction(TRANSECT, SITE, sep = " - "), y = mean_HPI, fill = as.factor(YEAR))) +
   geom_bar(stat = "identity", position = "dodge", color = "black") +
-  labs(x = "Transect - Site", y = "Mean Habitat Productivity Index (HPI)") +
+  labs(x = "Transect - Site", y = "Mean Habitat Productivity Index (HPI)", fill = "Year") +
   theme_minimal() +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
-
-```
+  theme(axis.text.x = element_text(angle = 45, hjust = 1));hpi.plot
+
+```

### 2024-06-27 scripts/data_cleaning.Rmd
updated HPI

@@
 # Function to rename columns using metadata
-# Get the unique column names from the QUAD ataframe
+# Get the unique column names from the QUAD dataframe
 column_names.quad <- colnames(QUAD)
-
 # Create a lookup table with name-to-code mappings
@@
 
+# Rename Kings Beach to King's Beach
+quad_long_with_type$SITE[quad_long_with_type$SITE == "Kings Beach"] <- "King's Beach"
+
 # use rbind.fill() method in R to combine data frames with different columns
-QUAD_data <- rbind.fill(QUAD_long, QUAD_2018, FtWeatherill_QUAD_2017, FtWeatherill_QUAD_2016_TALBENHORIN)
+QUAD_data <- rbind.fill(quad_long_with_type, QUAD_2018, FtWeatherill_QUAD_2017, FtWeatherill_QUAD_2016_TALBENHORIN)
 
 #remove 131-138 columns for observers not needed information
-QUAD_data <- QUAD_data[ -c(14:17) ]
+QUAD_data <- QUAD_data[ -c(14:19) ]
 

### 2024-06-27 scripts/manuscript_figures.Rmd
updated HPI

@@
 library(kableExtra)
+library(vegan)
 library(flextable)
@@
 
-# Rename fort weatherill
+# Rename kings beach
 fish.filt$SITE[fish.filt$SITE == "Kings Beach"] <- "King's Beach"
@@
 
-Production to biomass ratios (P:B) for each fish species
+Production to biomass ratios (P:B) and habitat productivity index for each fish species
 ```{r}
@@
 ```
+
+nMDS or PCA on benthic community structure
+```{r}
+# Load in QUAD data and filter for invertebrates
+invert.filt <- QUAD %>%
+  filter(type %in% c("molluscs", "echinoderms", "crustacean"))
+
+# I noticed that 06/30/23 dates were accidentally labeled 06/30/29
+# Replace "29" with "23" in the DATE column
+invert.filt$DATE <- sub("29$", "23", invert.filt$DATE)
+
+# Rename KB q to transect 6
+invert.filt$TRANSECT[invert.filt$TRANSECT == "KB q"] <- "6"
+
+# Sum count data at the transect level
+invert.count <- invert.filt %>%
+  group_by(TRANSECT, YEAR, SITE, type) %>%
+  summarize(count.sum = sum(COUNT))
+
+# Divide by 6 quadrats for correct abundance g/m2 calculation
+invert.count$abundance <- invert.count$count.sum / 6
+
+# Check for normality
+shapiro.test(invert.count$abundance) #not normal so use nMDS
+
+#make community matrix - extract columns with abundance information
+com = invert.count[,6:ncol(invert.count)]
+
+# Calculate Bray-Curtis dissimilarity
+m_com <- vegdist(com, method = "bray")
+
+# Check for warnings and errors
+if (any(warnings())) {
+  warning("Warnings during distance calculation:", warnings())
+}
+
+# Remove empty rows (if applicable)
+m_com <- as.matrix(m_com[complete.cases(m_com), ])
+
+# Check dimensions of m_com after removing empty rows
+dim(m_com)
+
+# Perform nMDS with Bray-Curtis dissimilarity
+nmds <- metaMDS(m_com, distance = "bray")
+
+# Create a dataframe for plotting
+plot_data <- data.frame(
+  NMDS1 = nmds$points[, 1],  # NMDS dimension 1
+  NMDS2 = nmds$points[, 2],  # NMDS dimension 2
+  type = invert.count$type,  # Type of invertebrate
+  abundance = invert.count$abundance  # Abundance of invertebrate
+)
+
+# Plot nMDS with type labels and scaled points by abundance
+ggplot(plot_data, aes(x = NMDS1, y = NMDS2, color = type, size = abundance)) +
+  geom_point(alpha = 0.8) +
+  scale_size_continuous(range = c(2, 8)) +  # Adjust point size range as needed
+  labs(title = "nMDS Plot of Invertebrate Types by Abundance",
+       x = "NMDS 1", y = "NMDS 2") +
+  theme_minimal() +
+  theme(legend.position = "right") +
+  geom_text(aes(label = type), nudge_x = 0.05, nudge_y = 0.05, size = 3)  # Add type labels
+```

### 2024-06-28 scripts/manuscript_figures.Rmd
updated nmds script

@@
 
-#make community matrix - extract columns with abundance information
-com = invert.count[,6:ncol(invert.count)]
+# Convert data from long to wide for nmds
+wide_data <- invert.count %>%
+  dplyr::select(TRANSECT, YEAR, SITE, type, abundance) %>%
+  pivot_wider(names_from = type, values_from = abundance, values_fill = list(abundance = 0))
+
+# Ensure abundance columns are numeric
+wide_data <- wide_data %>%
+  mutate(across(crustacean:molluscs, as.numeric))
 
 # Calculate Bray-Curtis dissimilarity
-m_com <- vegdist(com, method = "bray")
-
-# Check for warnings and errors
-if (any(warnings())) {
-  warning("Warnings during distance calculation:", warnings())
-}
-
-# Remove empty rows (if applicable)
-m_com <- as.matrix(m_com[complete.cases(m_com), ])
-
-# Check dimensions of m_com after removing empty rows
-dim(m_com)
-
-# Perform nMDS with Bray-Curtis dissimilarity
+m_com <- vegdist(wide_data[, -c(4:6)], method = "bray")
+
+# Perform nMDS
 nmds <- metaMDS(m_com, distance = "bray")

### 2024-06-28 scripts/data_cleaning.Rmd
updated nmds script and plot

@@
 ```{r}
-
 #make QUAD data long

### 2024-06-28 scripts/manuscript_figures.Rmd
updated nmds script and plot

@@
 # remove control column and two last columns, not needed
-quad.filt <- quad.filt[ -c(13:14) ] 
-quad.filt <- quad.filt[ -c(8) ] 
-quad.filt <- quad.filt[ -c(5:6) ]
+quad.filt <- quad.filt[ -c(9) ] 
+#quad.filt <- quad.filt[ -c(8) ] 
+#quad.filt <- quad.filt[ -c(5:6) ]
 
@@
 
+# Make line plot for temperature time series by site per year
+```{r}
+# Calculate mean and standard error of temperature by month, site, and year
+avg_temp <- temperature_data %>%
+  mutate(MonthAbbrev = month.abb[Month]) %>%  # Add month abbreviations
+  group_by(Year, MonthAbbrev, Site) %>%
+  summarize(
+    avg_temp = mean(Temp),
+    std_error = sd(Temp) / sqrt(n()),  # Calculate standard error
+    .groups = 'drop'  # Drop grouping after summarization
+  )
+
+# Map site codes to full names
+avg_temp$Site <- ifelse(avg_temp$Site == "FW", "Fort Wetherill", "King's Beach")
+
+# Ensure MonthAbbrev is a factor with levels in the correct chronological order
+avg_temp$MonthAbbrev <- factor(avg_temp$MonthAbbrev, levels = month.abb)
+
+# Plotting average temperature by site as lines and points per month with error bars
+temperature_plot <- ggplot(avg_temp, aes(x = MonthAbbrev, y = avg_temp, color = Site, group = interaction(Site, Year))) +
+  geom_line(size = 1.5) +  # Lines for average temperature by site and year
+  geom_point(size = 3) +  # Points for mean values by site and year
+  geom_errorbar(aes(ymin = avg_temp - std_error, ymax = avg_temp + std_error), width = 0.2) +  # Error bars for SEM
+  labs(x = "Month", y = expression("Average Temperature (°C)"), color = "Site") +
+  facet_wrap(~ Year) +  # Facet by year
+  scale_color_manual(values = c("Fort Wetherill" = "grey", "King's Beach" = "black")) +  # Custom colors
+  theme_bw() +
+  theme(
+    axis.title.x = element_text(size = 18),
+    axis.title.y = element_text(size = 18),
+    axis.text.x = element_text(size = 18),
+    axis.text.y = element_text(size = 18),
+    legend.text = element_text(size = 16),
+    legend.title = element_text(size = 18));temperature_plot
+
+# Save the plot to a file
+ggsave("output/figures/temp.plot.year.png", plot = temperature_plot, width = 20, height = 10)
+```
+
 # Make line plot for temperature time series by site per month
 ```{r}
-# Calculate mean and standard error of temperature by month and site
-avg_temp <- temperature_data %>%
+# Calculate mean and standard error of temperature by month, site, and year
+avg_temp_month <- temperature_data %>%
   mutate(MonthAbbrev = month.abb[Month]) %>%  # Add month abbreviations
@@
     avg_temp = mean(Temp),
-    std_error = sd(Temp) / sqrt(n()))  # Calculate standard error
+    std_error = sd(Temp) / sqrt(n()))  # Calculate standard error  # Drop grouping after summarization
 
 # Map site codes to full names
-avg_temp$Site <- ifelse(avg_temp$Site == "FW", "Fort Wetherill", "King's Beach")
+avg_temp_month$Site <- ifelse(avg_temp_month$Site == "FW", "Fort Wetherill", "King's Beach")
+
+# Ensure MonthAbbrev is a factor with levels in the correct chronological order
+avg_temp_month$MonthAbbrev <- factor(avg_temp_month$MonthAbbrev, levels = month.abb)
 
 # Plotting average temperature by site as lines and points per month with error bars
-temperature_plot <- ggplot(avg_temp, aes(x = MonthAbbrev, y = avg_temp, color = Site, group = Site)) +
+temperature_plot_month <- ggplot(avg_temp_month, aes(x = MonthAbbrev, y = avg_temp, color = Site, group = Site)) +
   geom_line(size = 1.5) +  # Lines for average temperature by site
@@
     legend.text = element_text(size = 16),
-    legend.title = element_text(size = 18));temperature_plot
+    legend.title = element_text(size = 18));temperature_plot_month
 
 # Save the plot to a file
-ggsave("output/figures/temp.plot.png", plot = temperature_plot, width = 20, height = 10)
+ggsave("output/figures/temp.plot.month.png", plot = temperature_plot_month, width = 20, height = 10)
 ```
@@
 
-# Ensure abundance columns are numeric
-wide_data <- wide_data %>%
-  mutate(across(crustacean:molluscs, as.numeric))
+# Ensure only numeric columns are selected for vegdist
+numeric_data <- wide_data[, c("crustacean", "echinoderms", "molluscs")]
+
+# Remove rows where all abundance values are zero
+numeric_data <- numeric_data[rowSums(numeric_data) > 0, ]
 
 # Calculate Bray-Curtis dissimilarity
-m_com <- vegdist(wide_data[, -c(4:6)], method = "bray")
+m_com <- vegdist(numeric_data, method = "bray")
 
@@
 
-# Create a dataframe for plotting
-plot_data <- data.frame(
-  NMDS1 = nmds$points[, 1],  # NMDS dimension 1
-  NMDS2 = nmds$points[, 2],  # NMDS dimension 2
-  type = invert.count$type,  # Type of invertebrate
-  abundance = invert.count$abundance  # Abundance of invertebrate
-)
-
-# Plot nMDS with type labels and scaled points by abundance
-ggplot(plot_data, aes(x = NMDS1, y = NMDS2, color = type, size = abundance)) +
-  geom_point(alpha = 0.8) +
-  scale_size_continuous(range = c(2, 8)) +  # Adjust point size range as needed
-  labs(title = "nMDS Plot of Invertebrate Types by Abundance",
-       x = "NMDS 1", y = "NMDS 2") +
-  theme_minimal() +
-  theme(legend.position = "right") +
-  geom_text(aes(label = type), nudge_x = 0.05, nudge_y = 0.05, size = 3)  # Add type labels
-```
+# Extract scores
+scores <- as.data.frame(scores(nmds))
+
+# Add row numbers as a unique identifier
+scores$ID <- row.names(scores)
+
+# Create a data frame with SITE and YEAR information
+site_year_info <- wide_data[rowSums(wide_data[, c("crustacean", "echinoderms", "molluscs")]) > 0, c("SITE", "YEAR")]
+site_year_info$ID <- row.names(site_year_info)
+
+# Merge SITE and YEAR information into scores
+scores <- merge(scores, site_year_info, by = "ID")
+
+# Plotting nMDS results
+nmds_plot <- ggplot(scores, aes(x = NMDS1, y = NMDS2, color = SITE, shape = as.factor(YEAR))) +
+  geom_point(size = 8) +
+  labs(x = "NMDS1", y = "NMDS2", color = "Site", shape = "Year") +
+  stat_ellipse(aes(group = YEAR), level = 0.95, linetype = 2, size = 1, alpha = 0.2)+
+  scale_color_manual(values = c("Fort Wetherill" = "grey", "King's Beach" = "black")) +
+  theme_bw() +
+  theme(
+    axis.title.x = element_text(size = 18),
+    axis.title.y = element_text(size = 18),
+    axis.text.x = element_text(size = 18),
+    axis.text.y = element_text(size = 18),
+    legend.text = element_text(size = 16),
+    legend.title = element_text(size = 18));nmds_plot
+
+```

### 2024-06-28 scripts/manuscript_figures.Rmd
calculated PERMANOVA and SIMPER results

@@
 
-```
+# Save the plot to a file
+ggsave("output/figures/nmds.plot.png", plot = nmds_plot, width = 15, height = 10)
+
+```
+
+PERMANOVA and SIMPER analysis
+```{r}
+# Remove rows where all abundance values are zero
+non_zero_rows <- rowSums(community_data) > 0
+community_data <- community_data[non_zero_rows, ]
+environmental_data <- environmental_data[non_zero_rows, ]
+
+# Convert YEAR to factor
+environmental_data$YEAR <- as.factor(environmental_data$YEAR)
+
+# Perform PERMANOVA
+permanova_result <- adonis2(community_data ~ SITE + YEAR, data = environmental_data, permutations = 999, method = "bray")
+
+# View the results
+print(permanova_result)
+
+# For the SIMPER analysis, we'll use the original (non-transformed) data, but still excluding the all-zero rows
+simper_result <- simper(community_data, group = environmental_data$SITE, permutations = 999)
+
+# View summary of results
+summary(simper_result)
+
+# Extract the relevant information from the SIMPER results
+simper_data <- data.frame(
+  Species = simper_result$`Fort Wetherill_King's Beach`$species,
+  Contribution = simper_result$`Fort Wetherill_King's Beach`$average,
+  Fort_Wetherill = simper_result$`Fort Wetherill_King's Beach`$ava,
+  Kings_Beach = simper_result$`Fort Wetherill_King's Beach`$avb
+)
+
+# Calculate the percentage contribution
+simper_data$Percentage <- simper_data$Contribution / sum(simper_data$Contribution) * 100
+
+# Reshape the data for plotting
+simper_long <- tidyr::pivot_longer(simper_data, 
+                                   cols = c(Fort_Wetherill, Kings_Beach),
+                                   names_to = "Site", 
+                                   values_to = "Abundance")
+
+# Create the plot
+ggplot(simper_long, aes(x = Species, y = Abundance, fill = Site)) +
+  geom_bar(stat = "identity", position = "dodge") +
+  geom_text(data = simper_data, 
+            aes(x = Species, y = max(simper_long$Abundance) + 0.5, 
+                label = sprintf("%.1f%%", Percentage)),
+            position = position_dodge(width = 0.9),
+            vjust = -0.5) +
+  labs(title = "Species Contributions to Differences Between Sites",
+       subtitle = "Based on SIMPER analysis",
+       y = "Average Abundance",
+       x = "Species") +
+  scale_fill_manual(values = c("Fort_Wetherill" = "grey", "Kings_Beach" = "black")) +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1),
+        legend.title = element_blank())
+
+```
+
+PERMANOVA results:
+adonis2(formula = community_data ~ SITE + YEAR, data = environmental_data, permutations = 999, method = "bray")
+         Df SumOfSqs      R2      F Pr(>F)    
+SITE      1   0.2632 0.02501 1.3160  0.235    
+YEAR      4   2.8619 0.27191 3.5774  0.001 ***
+Residual 37   7.4000 0.70308                  
+Total    42  10.5251 1.00000
+
+SIMPER results:
+Contrast: Fort Wetherill_King's Beach 
+            average      sd   ratio     ava     avb cumsum     p  
+molluscs    0.56710 0.30146 1.88120 7.64700 5.84300  0.894 0.803  
+echinoderms 0.04820 0.10148 0.47480 0.14700 0.34300  0.970 0.067 .
+crustacean  0.01920 0.04652 0.41350 0.30000 0.03700  1.000 0.713  
+---
+Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
+Permutation: free
+Number of permutations: 999

### 2024-06-28 scripts/manuscript_figures.Rmd
calculated PERMANOVA and SIMPER results

@@
 library(ggspatial)
+library(gridExtra)
+library(grid)
 

### 2024-06-28 scripts/manuscript_figures.Rmd
updated table output for permanova and simper

@@
 
+# Create dataframe from PERMANOVA results
+permanova_df <- data.frame(
+  Factor = c("SITE", "YEAR", "Residual", "Total"),
+  Df = c(1, 4, 37, 42),
+  SumOfSqs = c(0.2632, 2.8619, 7.4000, 10.5251),
+  R2 = c(0.02501, 0.27191, 0.70308, 1.00000),
+  F = c(1.3160, 3.5774, NA, NA),
+  Pr_F = c(0.235, 0.001, NA, NA)  # Changed column name to Pr_F
+)
+
+# Create flextable object
+flex_perm <- flextable(permanova_df)
+
+# Customize the flextable
+flex_perm <- flex_perm %>%
+  set_header_labels(
+    Factor = "Factor",
+    Df = "Df",
+    SumOfSqs = "Sum of Squares",
+    R2 = "R2",
+    F = "F",
+    Pr_F = "Pr(>F)"  # Changed to match the new column name
+  ) %>%
+  colformat_double(digits = 4) %>%
+  bold(i = ~ Factor == "YEAR", j = "Pr_F") %>%  # Changed to match the new column name
+  add_header_lines(values = c(
+    "Permutation test for adonis under reduced model",
+    "Terms added sequentially (first to last)",
+    "Permutation: free",
+    "Number of permutations: 999"
+  )) %>%
+  add_footer_lines(c(
+    "Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1",
+    "adonis2(formula = community_data ~ SITE + YEAR, data = environmental_data, permutations = 999, method = \"bray\")"
+  )) %>%
+  autofit() %>%
+  theme_box()
+
+# Save flextable as PNG
+png_file <- "output/model.result.tables/permanova.png"
+save_as_image(flex_perm, path = png_file)
+
 # For the SIMPER analysis, we'll use the original (non-transformed) data, but still excluding the all-zero rows
@@
 summary(simper_result)
+
+# Create dataframe from SIMPER results
+simp <- data.frame(
+  Species = rownames(summary(simper_result)$`Fort Wetherill_King's Beach`),
+  Average = summary(simper_result)$`Fort Wetherill_King's Beach`$average,
+  Ratio = summary(simper_result)$`Fort Wetherill_King's Beach`$ratio,
+  Fort_Wetherill = summary(simper_result)$`Fort Wetherill_King's Beach`$ava,
+  Kings_Beach = summary(simper_result)$`Fort Wetherill_King's Beach`$avb,
+  Cumulative_Sum = summary(simper_result)$`Fort Wetherill_King's Beach`$cumsum,
+  p_value = summary(simper_result)$`Fort Wetherill_King's Beach`$p
+)
+
+# Customize the flextable
+flex_simp <- flex_simp %>%
+  set_header_labels(
+    Species = "Species",
+    Average = "Avg. Contribution",
+    Ratio = "Ratio",
+    Fort_Wetherill = "Fort Wetherill Abundance",
+    Kings_Beach = "King's Beach Abundance",
+    Cumulative_Sum = "Cumulative Sum",
+    p_value = "p-value"
+  ) %>%
+  autofit() %>%
+  theme_box()
+
+# Save flextable as PNG
+png_file <- "output/model.result.tables/simper_results.png"
+save_as_image(flex_simp, path = png_file)
 

### 2024-07-05 scripts/manuscript_figures.Rmd
updated figures and models

@@
 # rerun model with log and without outliers
-lmm_model <- lmer(log(mean_fish_biomass + 1.1) ~ log(mean_kelp_biomass + 1.1) * YEAR + (1 | SITE), data = data_clean)
+lmm_model <- lmer(log(mean_fish_biomass + 1.1) ~ log(mean_kelp_biomass + 1.1) + (1 | YEAR) + (1 | SITE), data = data_clean)
 #check normality
 qqPlot(residuals(lmm_model)) #improved to normal
+summary(lmm_model)
+
+# Extract coefficients and their standard errors
+model_summary <- tidy(lmm_model)
+
+# Create a data frame with coefficients, standard errors, and p-values
+df <- data.frame(
+  Coefficient = model_summary$term,
+  Estimate = model_summary$estimate,
+  Std.Error = model_summary$std.error,
+  P.Value = model_summary$p.value
+)
+
+# Create flextable object
+flex <- flextable(df)
+
+# Save flextable as PNG
+png_file <- "output/model.result.tables/rugosity.not.included/summary_fish_kelp_biomass.png"
+save_as_image(flex, png_file)
 
@@
 
-# Export to CSV file
-write.csv(lmm_tidy, "output/lmm_model_results.csv", row.names = FALSE)
-
 # Calculate R-squared for the mixed model
@@
 # Create scatter plot
-figure.3 <- ggplot(aggregated.data.morph, aes(x = log(mean_kelp_biomass + 1.1), y = log(mean_fish_biomass + 1.1))) +
+figure.3 <- ggplot(aggregated.data.morph, aes(mean_kelp_biomass, mean_fish_biomass)) +
   geom_point() +
@@
            label = paste0("p = ", format(p_value_kelp_biomass, scientific = FALSE, digits = 2)), 
-           hjust = 1.3, vjust = -38.5, size = 4) +
-
-  labs(x = expression("Log Kelp Biomass (g/" ~ m^2 ~ ")"), y = expression("Log Fish Biomass (g/" ~ m^2 ~ ")")) +
+           hjust = 1.3, vjust = -39, size = 4) +
+
+  labs(x = expression("Kelp Biomass (g/" ~ m^2 ~ ")"), y = expression("Fish Biomass (g/" ~ m^2 ~ ")")) +
   theme_classic() +
@@
 ggsave("output/figures/figure_3.png", plot = figure.3, width = 8, height = 6)
+
 ```
@@
 # Time series plot for kelp biomass with standard error bars from 2016-2023
-kelp.timeseries.2016 <- ggplot(aggregated.kelp.morph, aes(x = YEAR, y = log_mean_kelp_biomass, shape = SITE)) +
+kelp.timeseries.2016 <- ggplot(aggregated.kelp.morph, aes(x = YEAR, y = mean_kelp_biomass, shape = SITE)) +
   geom_line(aes(group = SITE), color = "black") +
-  geom_point(size = 3) +
-  geom_errorbar(aes(ymin = log_mean_kelp_biomass - log_se_kelp_biomass, 
-                    ymax = log_mean_kelp_biomass + log_se_kelp_biomass), 
+  geom_point(size = 6) +
+  geom_errorbar(aes(ymin = mean_kelp_biomass - se_kelp_biomass, 
+                    ymax = mean_kelp_biomass + se_kelp_biomass), 
                 width = 0.2) +
   scale_shape_manual(values = shape_values) +
-  labs(x = "Year", y = expression("Log Mean Kelp Biomass (g/" ~ m^2 ~ ")"), shape = "Site") +
+  labs(x = "Year", y = expression("Mean Kelp Biomass (g/" ~ m^2 ~ ")"), shape = "Site") +
   theme_bw() +
@@
     legend.text = element_text(size = 16),
-    legend.title = element_text(size = 18)
+    legend.title = element_text(size = 18),
+    legend.position = c(1, 1),  # Positioning legend inside the plot area at the top right
+    legend.justification = c(1, 1),  # Anchoring the legend to the top right corner
+    legend.background = element_rect(fill = "white", color = "black")  # Optional: adding a background for better visibility
   );kelp.timeseries.2016
@@
   geom_line(aes(group = SITE), color = "black") +
-  geom_point(size = 3) +
+  geom_point(size = 8) +
   geom_errorbar(aes(ymin = log_mean_kelp_biomass - log_se_kelp_biomass, 
@@
     se_slope = summary(lm(log_avg_kelp_biomass ~ YEAR))$coefficients[2, "Std. Error"],  # Standard error of slope
-    mean_log_avg_kelp_biomass = mean(log_avg_kelp_biomass)  # Mean log average kelp biomass
+    mean_log_avg_kelp_biomass = mean(log_avg_kelp_biomass),  # Mean log average kelp biomass
+    mean_nonlog_avg_kelp_biomass = mean(average_kelp_biomass)  # Mean non-log average kelp biomass
   )
@@
 # Plotting the mean rates of change with error bars (standard error horizontally)
-rate_of_change <- ggplot(slope_results, aes(x = rate_of_change_per_year, y = mean_log_avg_kelp_biomass, shape = SITE)) +
-  geom_point(size = 3) +
-  geom_errorbarh(aes(xmin = rate_of_change_per_year - se_rate_of_change, xmax = rate_of_change_per_year + se_rate_of_change), height = 0.2) +
+rate_of_change <- ggplot(slope_results, aes(x = rate_of_change_per_year, y = mean_nonlog_avg_kelp_biomass, shape = SITE)) +
+  geom_point(size = 8) +
+  geom_errorbarh(aes(xmin = rate_of_change_per_year - se_rate_of_change, xmax = rate_of_change_per_year + se_rate_of_change), height = 0.9) +
   labs(
     x = expression("Rate of Change ("~yr^-1 ~")"),
-    y = expression("Log Mean Kelp Biomass (g/"~m^2~")"), shape = "Site") +
+    y = expression("Mean Kelp Biomass (g/"~m^2~")"), shape = "Site") +
   theme_bw() +
@@
     legend.text = element_text(size = 16),
-    legend.title = element_text(size = 18));rate_of_change
+    legend.title = element_text(size = 18), ,
+    legend.position = "none");rate_of_change
 
@@
 
+#filter out data from 2016 and 2017 due to size bins
+PB_filtered <- PB %>%
+  filter(YEAR != 2016 & YEAR != 2017)
+
 # Calculate fish biomass 'W' using the length-weight relationship
-PB$biomass <- with(PB, coef.a * (actual_size)^coef.b)
+PB_filtered$biomass <- with(PB, coef.a * (actual_size)^coef.b)
 
@@
 
-# Plotting nMDS results
+# Define custom shape values for more than 8 categories
+custom_shape_values <- c(16, 17, 18, 19, 20, 21, 22, 23)
+
+# Use these custom shapes in your plot
 nmds_plot <- ggplot(scores, aes(x = NMDS1, y = NMDS2, color = SITE, shape = as.factor(YEAR))) +
   geom_point(size = 8) +
+  #stat_ellipse(aes(group = YEAR), level = 0.95, linetype = 2, size = 1, alpha = 0.5) + # Add ellipses
   labs(x = "NMDS1", y = "NMDS2", color = "Site", shape = "Year") +
-  stat_ellipse(aes(group = YEAR), level = 0.95, linetype = 2, size = 1, alpha = 0.2)+
   scale_color_manual(values = c("Fort Wetherill" = "grey", "King's Beach" = "black")) +
+  scale_shape_manual(values = custom_shape_values) +  # Specify the custom shapes
   theme_bw() +
@@
     legend.text = element_text(size = 16),
-    legend.title = element_text(size = 18));nmds_plot
+    legend.title = element_text(size = 18)
+  );nmds_plot
 
@@
 ```{r}
+# Ensure only numeric columns are selected for vegdist
+community_data <- wide_data[, c("crustacean", "echinoderms", "molluscs")]
+
 # Remove rows where all abundance values are zero
@@
 community_data <- community_data[non_zero_rows, ]
-environmental_data <- environmental_data[non_zero_rows, ]
+
+# Create environmental data frame
+environmental_data <- wide_data[non_zero_rows, c("TRANSECT", "YEAR", "SITE")]
 
@@
   Factor = c("SITE", "YEAR", "Residual", "Total"),
-  Df = c(1, 4, 37, 42),
-  SumOfSqs = c(0.2632, 2.8619, 7.4000, 10.5251),
-  R2 = c(0.02501, 0.27191, 0.70308, 1.00000),
-  F = c(1.3160, 3.5774, NA, NA),
-  Pr_F = c(0.235, 0.001, NA, NA)  # Changed column name to Pr_F
+  Df = c(1, 7, 44, 52),
+  SumOfSqs = c(0.5649, 5.0492, 8.7778, 14.3919),
+  R2 = c(0.03925, 0.35084, 0.60991, 1.00000),
+  F = c(2.8317, 3.6157, NA, NA),
+  Pr_F = c(0.050, 0.001, NA, NA)  # Changed column name to Pr_F
 )
@@
   colformat_double(digits = 4) %>%
-  bold(i = ~ Factor == "YEAR", j = "Pr_F") %>%  # Changed to match the new column name
+  bold(i = ~ Pr_F <= 0.05, j = "Pr_F") %>%  # Bold values if Pr_F is 0.05 or lower
   add_header_lines(values = c(
     "Permutation test for adonis under reduced model",
-    "Terms added sequentially (first to last)",
-    "Permutation: free",
     "Number of permutations: 999"
-  )) %>%
-  add_footer_lines(c(
-    "Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1",
-    "adonis2(formula = community_data ~ SITE + YEAR, data = environmental_data, permutations = 999, method = \"bray\")"
   )) %>%
@@
 
-# For the SIMPER analysis, we'll use the original (non-transformed) data, but still excluding the all-zero rows
+# Perform SIMPER analysis
 simper_result <- simper(community_data, group = environmental_data$SITE, permutations = 999)
@@
 )
+
+# Create flextable object
+flex_simp <- flextable(simp)
 
@@
 simper_data <- data.frame(
-  Species = simper_result$`Fort Wetherill_King's Beach`$species,
-  Contribution = simper_result$`Fort Wetherill_King's Beach`$average,
-  Fort_Wetherill = simper_result$`Fort Wetherill_King's Beach`$ava,
-  Kings_Beach = simper_result$`Fort Wetherill_King's Beach`$avb
+  Species = rownames(summary(simper_result)$`Fort Wetherill_King's Beach`),
+  Contribution = summary(simper_result)$`Fort Wetherill_King's Beach`$average,
+  Fort_Wetherill = summary(simper_result)$`Fort Wetherill_King's Beach`$ava,
+  Kings_Beach = summary(simper_result)$`Fort Wetherill_King's Beach`$avb
 )
@@
 # Create the plot
-ggplot(simper_long, aes(x = Species, y = Abundance, fill = Site)) +
+simper_plot <- ggplot(simper_long, aes(x = Species, y = Abundance, fill = Site)) +
   geom_bar(stat = "identity", position = "dodge") +
-  geom_text(data = simper_data, 
-            aes(x = Species, y = max(simper_long$Abundance) + 0.5, 
-                label = sprintf("%.1f%%", Percentage)),
+  geom_text(aes(label = sprintf("%.1f%%", Percentage[match(Species, simper_data$Species)])),
             position = position_dodge(width = 0.9),
-            vjust = -0.5) +
+            vjust = -0.5, 
+            size = 3) +
   labs(title = "Species Contributions to Differences Between Sites",
@@
   theme(axis.text.x = element_text(angle = 45, hjust = 1),
-        legend.title = element_blank())
+        legend.title = element_blank());simper_plot
+
+# Save the plot to a file
+ggsave("output/figures/simper.plot.png", plot = simper_plot, width = 15, height = 10)
 
@@
          Df SumOfSqs      R2      F Pr(>F)    
-SITE      1   0.2632 0.02501 1.3160  0.235    
-YEAR      4   2.8619 0.27191 3.5774  0.001 ***
-Residual 37   7.4000 0.70308                  
-Total    42  10.5251 1.00000
+SITE      1   0.5649 0.03925 2.8317  0.043 *  
+YEAR      7   5.0492 0.35084 3.6157  0.001 ***
+Residual 44   8.7778 0.60991                  
+Total    52  14.3919 1.00000
 
 SIMPER results:
-Contrast: Fort Wetherill_King's Beach 
+ontrast: Fort Wetherill_King's Beach
+
             average      sd   ratio     ava     avb cumsum     p
-molluscs    0.56710 0.30146 1.88120 7.64700 5.84300  0.894 0.803  
-echinoderms 0.04820 0.10148 0.47480 0.14700 0.34300  0.970 0.067 .
-crustacean  0.01920 0.04652 0.41350 0.30000 0.03700  1.000 0.713  
----
-Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
+molluscs    0.58750 0.29784 1.97250 5.57600 5.84300  0.863 0.576
+echinoderms 0.07140 0.12811 0.55760 0.21400 0.34300  0.968 0.547
+crustacean  0.02160 0.05018 0.43030 0.21900 0.09300  1.000 0.526
 Permutation: free

### 2024-07-05 scripts/data_cleaning.Rmd
updasted HPI and production plots

@@
 
+# Rename columns in species_metadata_type to match those in QUAD_data
+colnames(species_metadata_type) <- c("SP_CODE", "type")
+
+# Perform a left join to update the type column in QUAD_data
+QUAD_data <- QUAD_data %>%
+  left_join(species_metadata_type, by = "SP_CODE", suffix = c("", ".new"))
+
+# Update the original type column with the new type values where the original type is NA
+QUAD_data$type <- ifelse(is.na(QUAD_data$type), QUAD_data$type.new, QUAD_data$type)
+
+# Remove the temporary new type column using base R
+QUAD_data$type.new <- NULL
+
 # output updated dataframe

### 2024-07-05 scripts/manuscript_figures.Rmd
updasted HPI and production plots

@@
 
+# Rename kings beach and fort wetherill
+PB$SITE[PB$SITE == "Kings Beach"] <- "King's Beach"
+PB$SITE[PB$SITE == "Fort Weatherill"] <- "Fort Wetherill"
+
 #filter out data from 2016 and 2017 due to size bins
@@
 # Calculate fish biomass 'W' using the length-weight relationship
-PB_filtered$biomass <- with(PB, coef.a * (actual_size)^coef.b)
-
-# Divide biomass by 80 m to get g/m2
-#calculate biomass/m2
-PB$biomass.g.m2 <- PB$biomass / 80
-
-# Constants from the P/B ratio equation
+PB_filtered$Wmat <- PB_filtered$coef.a * (PB_filtered$Lmat^PB_filtered$coef.b)
+
+# Constants for the PB.ratio calculation
 coefficient <- 2.64
@@
 
-# Calculate P/B ratio, the predictive equation is P/B (per year) = 2.64*Wmat^-035 , where Wmat is weight-at-maturity (Randall and Minns 2000)
-PB$P_B_ratio <- coefficient * PB$biomass.g.m2^exponent
+# Calculate PB.ratio column
+PB_filtered$PB.ratio <- coefficient * (PB_filtered$Wmat ^ exponent)
+
+# Calculate fish biomass
+# Calculate total weight for each fish species
+PB_filtered$biomass <- PB_filtered$Wmat * PB_filtered$COUNT
+
+# Calculate biomass per m²
+PB_filtered$biomass_per_m2 <- PB_filtered$biomass / 80
+
+# Calculate production for each species in each transect
+production_data <- PB_filtered %>%
+  group_by(TRANSECT, SP_CODE, YEAR) %>%
+  summarise(
+    Total_Biomass = sum(biomass_per_m2, na.rm = TRUE),  # Sum of biomass for the species in the transect
+    Total_Production = sum(biomass_per_m2 * PB.ratio, na.rm = TRUE)  # Total production for the species in the transect
+  )
+
+# Determine the maximum and minimum values for Total_Production
+max_prod <- max(production_data$Total_Production)
+min_prod <- min(production_data$Total_Production)
+
+# Plotting using ggplot2
+production.plot <- ggplot(production_data, aes(x = YEAR, y = Total_Production, fill = SP_CODE)) +
+  geom_bar(stat = "identity", position = "dodge") +
+  labs(x = "Year",
+       y = "Fish Production",
+       fill = "Species") +
+  theme_bw() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1),  # Rotate x-axis labels if needed
+        axis.title.x = element_text(vjust = -0.5),  # Adjust x-axis title position
+        axis.title.y = element_text(vjust = 1.5)) + # Adjust y-axis title position
+  scale_y_continuous(limits = c(0, 650), expand = c(0, 0));production.plot  # Set y-axis limits and remove expansion
 
 # Calculate Habitat Productivity Index (HPI), habitat productivity index (HPI), a proposed measure of habitat suitability and productive capacity, can be calculated as the product of average fish biomass and the species P/B coefficient (HPI = B × P/B) (Randall and Minns 2000)
-PB$HPI <- PB$biomass.g.m2 * PB$P_B_ratio
+# Calculate average biomass per SP_CODE
+average_biomass_df <- PB_filtered %>%
+  group_by(SP_CODE, PB.ratio, YEAR, TRANSECT, SITE) %>%
+  summarise(average_biomass = mean(biomass_per_m2, na.rm = TRUE))
+
+# Calculate Habitat Productivity Index (HPI)
+average_biomass_df$HPI <- average_biomass_df$average_biomass * average_biomass_df$PB.ratio
 
 # Aggregate by Transect, Site, Year and calculate mean HPI
-hpi_summary <- PB %>%
-  group_by(TRANSECT, SITE, YEAR) %>%
+hpi_summary <- average_biomass_df %>%
+  group_by(SITE, YEAR) %>%
   summarise(mean_HPI = mean(HPI, na.rm = TRUE))
 
-# Make faceted line plot of HPI index over the years
-hpi.plot < - ggplot(hpi_summary, aes(x = interaction(TRANSECT, SITE, sep = " - "), y = mean_HPI, fill = as.factor(YEAR))) +
-  geom_bar(stat = "identity", position = "dodge", color = "black") +
-  labs(x = "Transect - Site", y = "Mean Habitat Productivity Index (HPI)", fill = "Year") +
-  theme_minimal() +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1));hpi.plot
+# Make faceted line plot of HPI index over the years as a area plot
+hpi.plot <- ggplot(hpi_summary, aes(x = YEAR, y = mean_HPI, fill = SITE)) +
+  geom_area(position = "stack") +
+  labs(x = "Year",
+       y = "Mean Habitat productivity index",
+       fill = "Site") +
+  theme_minimal(); hpi.plot
 
@@
 # Define custom shape values for more than 8 categories
-custom_shape_values <- c(16, 17, 18, 19, 20, 21, 22, 23)
+custom_shape_values <- c(0, 1, 2, 3, 4, 5, 6, 23)
 

### 2024-07-06 scripts/manuscript_figures.Rmd
updasted HPI and production plots

--- 2024-07-05 scripts/manuscript_figures.Rmd
@@
 
-# Calculate P/B ratio, the predictive equation is P/B (per year) = 2.64*Wmat^-035 , where Wmat is weight-at-maturity (Randall and Minns 2000)
-PB$P_B_ratio <- coefficient * PB$biomass.g.m2^exponent
+# Calculate PB.ratio column
+PB_filtered$PB.ratio <- coefficient * (PB_filtered$Wmat ^ exponent)
+
+# Calculate fish biomass
+PB_filtered$biomass <- PB_filtered$Wmat * PB_filtered$COUNT
+
+# Calculate biomass per m²
+PB_filtered$biomass_per_m2 <- PB_filtered$biomass / 80
+
+# Plotting using ggplot2
+production.plot <- ggplot(PB_filtered, aes(x = YEAR, y = PB.ratio, fill = SP_CODE)) +
+  geom_bar(stat = "identity", position = "dodge") +
+  labs(x = "Species",
+       y = "P/B",
+       fill = "Species") +
+  theme_bw() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 18),  # Rotate x-axis labels if needed
+        axis.title.x = element_text(vjust = -0.5, size = 18),  # Adjust x-axis title position
+        axis.title.y = element_text(vjust = 1.5, size = 18),
+        axis.text.y = element_text(size = 18),
+        legend.text = element_text(size = 16),
+        legend.title = element_text(size = 18)) + # Adjust y-axis title position
+  scale_y_continuous(limits = c(0, 4), expand = c(0, 0));production.plot  # Set y-axis limits and remove expansion
 
 # Calculate Habitat Productivity Index (HPI), habitat productivity index (HPI), a proposed measure of habitat suitability and productive capacity, can be calculated as the product of average fish biomass and the species P/B coefficient (HPI = B × P/B) (Randall and Minns 2000)
-PB$HPI <- PB$biomass.g.m2 * PB$P_B_ratio
-
-# Aggregate by Transect, Site, Year and calculate mean HPI
-hpi_summary <- PB %>%
-  group_by(TRANSECT, SITE, YEAR) %>%
-  summarise(mean_HPI = mean(HPI, na.rm = TRUE))
-
-# Make faceted line plot of HPI index over the years
-hpi.plot < - ggplot(hpi_summary, aes(x = interaction(TRANSECT, SITE, sep = " - "), y = mean_HPI, fill = as.factor(YEAR))) +
-  geom_bar(stat = "identity", position = "dodge", color = "black") +
-  labs(x = "Transect - Site", y = "Mean Habitat Productivity Index (HPI)", fill = "Year") +
+# Calculate HPI
+HPI_calculated <- PB_filtered %>%
+  group_by(YEAR, SITE, TRANSECT) %>%
+  summarise(HPI = sum(biomass_per_m2 * PB.ratio, na.rm = TRUE),
+            .groups = 'drop')
+
+# We add a small constant (1e-6) to avoid issues with log(0)
+HPI_calculated <- HPI_calculated %>%
+  mutate(log_HPI = log(HPI + 1.1))
+
+# Now plot with the log-transformed HPI
+species.hpi.plot <- ggplot(HPI_calculated, aes(x = SITE, y = log_HPI, fill = factor(YEAR))) +
+  geom_bar(stat = "identity", position = "dodge") +
+  labs(x = "Site", 
+       y = "Log Habitat Productivity Index", 
+       fill = "Year") +
   theme_minimal() +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1));hpi.plot
+  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 18),
+        axis.text.y = element_text(size = 18),
+        axis.title.y = element_text(size = 18),
+         axis.title.x = element_text(size = 18),
+        legend.text = element_text(size = 16),
+        legend.title = element_text(size = 18)) +
+  scale_y_continuous(expand = expansion(mult = c(0, .05)),
+    labels = scales::label_number(accuracy = 0.1)) +
+  coord_cartesian(clip = "off"); species.hpi.plot
+
+```
+
+# Plot two plots together using patchwork for production and HPI
+```{r}
+# Combine the plots with patchwork for all years of kelp and rate of change
+production.hpi <- production.plot | species_hpi.plot +
+  plot_annotation(tag_levels = 'A') + 
+  theme(plot.tag = element_text(size = 20, face = "bold"));production.hpi
+
+# Save the combined plot
+ggsave("output/figures/kelp.2016.rate.png", kelp.rate.2016, width = 20, height = 8, units = "in", dpi = 300)
 

### 2024-07-06 scripts/manuscript_figures.Rmd
added production and HPI data

@@
 library(vegan)
+library(RColorBrewer)
 library(flextable)
@@
 
+# add 2018 fish data to fish.filt dataframe
+fish.2018 <- read.csv("data/fish.data.clean.calc.csv")
+
+# Filter for only 2018 data
+fish.2018.filtered <- subset(fish.2018, YEAR == 2018)
+
+# Rename kings beach and fort wetherill
+fish.2018.filtered$SITE[fish.2018.filtered$SITE == "Kings Beach"] <- "King's Beach"
+fish.2018.filtered$SITE[fish.2018.filtered$SITE == "Fort Weatherill"] <- "Fort Wetherill"
+
+# Calculate fish biomass 'W' using the length-weight relationship
+fish.2018.filtered$Wmat <- fish.2018.filtered$coef.a * (fish.2018.filtered$Lmat^fish.2018.filtered$coef.b)
+
+# Calculate fish biomass
+fish.2018.filtered$Fish.Biomass..g.80m2. <- fish.2018.filtered$Wmat * fish.2018.filtered$COUNT
+
+# Calculate biomass per m²
+fish.2018.filtered$Fish.Biomass..g.m2. <- fish.2018.filtered$Fish.Biomass..g.80m2. / 80
+
+# selcet specific columns needed to merge
+fish.2018.selected <- fish.2018.filtered[, c("YEAR", "HABITAT", "REGION", "SITE", "TRANSECT", "COUNT", "Fish.Biomass..g.m2.", "Fish.Biomass..g.80m2.")]
+
+# Convert the integer column to character
+fish.2018.selected$TRANSECT <- as.character(fish.2018.selected$TRANSECT)
+
+#add area and n column
+fish.2018.selected$AREA <- 80
+fish.2018.selected <- fish.2018.selected %>%
+  rename(n = COUNT)
+
+# Merge (rbind) fish.2018.selected with fish.filt
+fish.selected <- rbind(fish.filt, fish.2018.selected)
+
 # Assuming you want to join by the 'transect' column
-merged.dat.cover <- left_join(fish.filt, kelp.filt)
+merged.dat.cover <- left_join(fish.selected, kelp.filt)
 
@@
 quad.filt <- QUAD %>%
-        filter(SP_CODE == "SUGK")
+  filter(SP_CODE == "SUGK" | SP_CODE == "LADI")
 
@@
 # Merge fish and morph data
-merged.dat.morph <- left_join(morph.filt, fish.filt)
+merged.dat.morph <- left_join(morph.filt, fish.selected)
 
@@
 qqPlot(residuals(lmm_model)) #improved to normal
+summary(lmm_model)
 
 # Identify the outliers
-outliers <- c(45, 37, 25, 33, 37, 22)
+outliers <- c(45, 37, 29,37)
 
@@
 
-# Plotting using ggplot2
-production.plot <- ggplot(PB_filtered, aes(x = YEAR, y = PB.ratio, fill = SP_CODE)) +
+# Convert YEAR to factor (if it's numeric)
+PB_filtered$YEAR <- factor(PB_filtered$YEAR)
+
+# Calculate production by multiplying biomass_per_m2 by PB raatio
+PB_filtered$Production <- PB_filtered$PB.ratio * PB_filtered$biomass_per_m2
+
+# make commmon name column
+PB_filtered <- PB_filtered %>%
+  mutate(Species = case_when(
+    SP_CODE == "BRTY" ~ "Atlantic menhaden",
+    SP_CODE == "CEST" ~ "Black sea bass",
+    SP_CODE == "MOSA" ~ "Striped bass",
+    SP_CODE == "PADE" ~ "Summer flounder",
+    SP_CODE == "STCH" ~ "Scup",
+    SP_CODE == "TAAD" ~ "Cunner",
+    SP_CODE == "TAON" ~ "Tautog",
+    TRUE ~ SP_CODE  # This will keep the original code if it doesn't match any of the above
+  ))
+
+# Now you can use this new 'Species' column in your ggplot
+production_plot <- ggplot(PB_filtered, aes(x = YEAR, y = Production, fill = Species)) +
   geom_bar(stat = "identity", position = "dodge") +
-  labs(x = "Species",
-       y = "P/B",
-       fill = "Species") +
+  labs(x = "Year",  # X-axis label
+       y = expression("Fish Production" ~ (g ~ m^-2 ~ yr^-1)),  # Y-axis label
+       fill = "Species") +  # This should now correctly set the legend title
+  scale_fill_viridis_d() +  # Use viridis colorblind-friendly palette
   theme_bw() +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 18),  # Rotate x-axis labels if needed
-        axis.title.x = element_text(vjust = -0.5, size = 18),  # Adjust x-axis title position
-        axis.title.y = element_text(vjust = 1.5, size = 18),
-        axis.text.y = element_text(size = 18),
-        legend.text = element_text(size = 16),
-        legend.title = element_text(size = 18)) + # Adjust y-axis title position
-  scale_y_continuous(limits = c(0, 4), expand = c(0, 0));production.plot  # Set y-axis limits and remove expansion
+  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 12),  # Rotate x-axis labels if needed
+        axis.title.x = element_text(vjust = -0.5, size = 14),  # Adjust x-axis title position
+        axis.title.y = element_text(vjust = 1.5, size = 14),
+        axis.text.y = element_text(size = 12),
+        legend.text = element_text(size = 12),
+        legend.title = element_text(size = 14)) +
+  scale_y_continuous(limits = c(0, 4), expand = c(0, 0));production_plot  # Additional attempt to change legend title  # Use viridis colorblind-friendly palette
 
@@
 
-# We add a small constant (1e-6) to avoid issues with log(0)
+# Log transform HPI 
 HPI_calculated <- HPI_calculated %>%
@@
 
-# Now plot with the log-transformed HPI
-species.hpi.plot <- ggplot(HPI_calculated, aes(x = SITE, y = log_HPI, fill = factor(YEAR))) +
+# Define the color palette
+my_palette <- brewer.pal(n = 6, name = "Paired")  # Adjust 'n' as per the number of years in your data
+
+# Plotting code
+hpi.plot <- ggplot(HPI_calculated, aes(x = SITE, y = log_HPI, fill = factor(YEAR))) +
   geom_bar(stat = "identity", position = "dodge") +
@@
        fill = "Year") +
+  scale_fill_manual(values = my_palette) +  # Use the colorblind-friendly palette
   theme_minimal() +
@@
                      labels = scales::label_number(accuracy = 0.1)) +
-  coord_cartesian(clip = "off"); species.hpi.plot
+  coord_cartesian(clip = "off");hpi.plot
 
@@
 # Combine the plots with patchwork for all years of kelp and rate of change
-production.hpi <- production.plot | species_hpi.plot +
+production.hpi <- production_plot | hpi.plot +
   plot_annotation(tag_levels = 'A') + 
@@
 # Save the combined plot
-ggsave("output/figures/kelp.2016.rate.png", kelp.rate.2016, width = 20, height = 8, units = "in", dpi = 300)
+ggsave("output/figures/production.HPI.png", production.hpi, width = 20, height = 10, units = "in", dpi = 300)
 

### 2024-07-09 scripts/manuscript_figures.Rmd
updated habitat productivity index plots

@@
   theme_bw() +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 12),  # Rotate x-axis labels if needed
-        axis.title.x = element_text(vjust = -0.5, size = 14),  # Adjust x-axis title position
-        axis.title.y = element_text(vjust = 1.5, size = 14),
-        axis.text.y = element_text(size = 12),
-        legend.text = element_text(size = 12),
-        legend.title = element_text(size = 14)) +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 18),  # Rotate x-axis labels if needed
+        axis.title.x = element_text(vjust = -0.5, size = 18),  # Adjust x-axis title position
+        axis.title.y = element_text(vjust = 1.5, size = 18),
+        axis.text.y = element_text(size = 18),
+        legend.text = element_text(size = 16),
+        legend.title = element_text(size = 18)) +
   scale_y_continuous(limits = c(0, 4), expand = c(0, 0));production_plot  # Additional attempt to change legend title  # Use viridis colorblind-friendly palette
@@
 
-# Log transform HPI 
+# Add log_HPI column to the dataframe
 HPI_calculated <- HPI_calculated %>%
-  mutate(log_HPI = log(HPI + 1.1))
-
-# Define the color palette
-my_palette <- brewer.pal(n = 6, name = "Paired")  # Adjust 'n' as per the number of years in your data
-
-# Plotting code
-hpi.plot <- ggplot(HPI_calculated, aes(x = SITE, y = log_HPI, fill = factor(YEAR))) +
-  geom_bar(stat = "identity", position = "dodge") +
-  labs(x = "Site", 
-       y = "Log Habitat Productivity Index", 
-       fill = "Year") +
-  scale_fill_manual(values = my_palette) +  # Use the colorblind-friendly palette
-  theme_minimal() +
+  mutate(log_HPI = log(HPI))
+
+# Summarize by year and site for plot
+HPI_summary <- HPI_calculated %>%
+  group_by(YEAR, SITE) %>%
+  summarise(
+    mean_HPI = mean(HPI, na.rm = TRUE),
+    se_HPI = sd(HPI, na.rm = TRUE) / sqrt(n()),
+    mean_log_HPI = mean(log_HPI, na.rm = TRUE),
+    se_log_HPI = sd(log_HPI, na.rm = TRUE) / sqrt(n()),
+    .groups = 'drop'
+  )
+
+# Create the plot with nonlogged values
+hpi.plot.nonlog <- ggplot(HPI_summary, aes(x = factor(YEAR), y = mean_HPI, group = SITE)) +
+  geom_line() +
+  geom_point(aes(shape = SITE), size = 3) +
+  geom_errorbar(aes(ymin = mean_HPI - se_HPI, ymax = mean_HPI + se_HPI), width = 0.2) +
+  labs(x = "Year", 
+       y = "Habitat Productivity Index", 
+       shape = "Site") +
+  theme_bw() +
   theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 18),
         axis.text.y = element_text(size = 18),
-        axis.title.y = element_text(size = 18),
-        axis.title.x = element_text(size = 18),
+        axis.title = element_text(size = 18),
         legend.text = element_text(size = 16),
         legend.title = element_text(size = 18)) +
-  scale_y_continuous(expand = expansion(mult = c(0, .05)),
-                     labels = scales::label_number(accuracy = 0.1)) +
-  coord_cartesian(clip = "off");hpi.plot
-
-```
+  scale_y_continuous(expand = expansion(mult = c(0.05, 0.05)),
+                     labels = scales::label_number(accuracy = 0.1));hpi.plot.nonlog
+
+# Create the plot with log_HPI values
+hpi_plot_log <- ggplot(HPI_summary, aes(x = factor(YEAR), y = mean_log_HPI, group = SITE)) +
+  geom_line() +
+  geom_point(aes(shape = SITE), size = 3) +
+  geom_errorbar(aes(ymin = mean_log_HPI - se_log_HPI, ymax = mean_log_HPI + se_log_HPI), width = 0.2) +
+  labs(x = "Year", 
+       y = "Log Habitat Productivity Index", 
+       shape = "Site") +
+  theme_bw() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 18),
+        axis.text.y = element_text(size = 18),
+        axis.title = element_text(size = 18),
+        legend.text = element_text(size = 16),
+        legend.title = element_text(size = 18)) +
+  scale_y_continuous(expand = expansion(mult = c(0.05, 0.05)),
+                     labels = scales::label_number(accuracy = 0.1));hpi_plot_log
+
+# Define the color palette
+#my_palette <- brewer.pal(n = 6, name = "Paired")  # Adjust 'n' as per the number of years in your data
+
+# Test for normality and perform appropriate statistical test
+# Check for normality using Shapiro-Wilk test for each site
+shapiro_test_fort_wetherill <- shapiro.test(HPI_calculated$HPI[HPI_calculated$SITE == "Fort Wetherill"])
+shapiro_test_kings_beach <- shapiro.test(HPI_calculated$HPI[HPI_calculated$SITE == "King's Beach"])
+
+# Print results of normality tests
+print(shapiro_test_fort_wetherill)
+print(shapiro_test_kings_beach)
+
+# Randomize the YEAR column
+set.seed(123)  # Set seed for reproducibility
+HPI_calculated$YEAR <- sample(HPI_calculated$YEAR)
+
+# Perform the Mann-Whitney U test on logged HPI values
+wilcox_test_log_result <- wilcox.test(HPI ~ SITE, data = HPI_calculated)
+
+# Print the result of the Wilcoxon rank-sum test
+print(wilcox_test_log_result)
+```
+Wilcoxon rank sum exact test
+
+data:  HPI by SITE
+W = 527, p-value = 0.03417
+alternative hypothesis: true location shift is not equal to 0
 
@@
 # Combine the plots with patchwork for all years of kelp and rate of change
-production.hpi <- production_plot | hpi.plot +
+production.hpi.nonlog <- production_plot | hpi.plot.nonlog +
   plot_annotation(tag_levels = 'A') + 
-  theme(plot.tag = element_text(size = 20, face = "bold"));production.hpi
+  theme(plot.tag = element_text(size = 20, face = "bold"));production.hpi.nonlog
 
 # Save the combined plot
-ggsave("output/figures/production.HPI.png", production.hpi, width = 20, height = 10, units = "in", dpi = 300)
+ggsave("output/figures/production.HPI.nonlog.png", production.hpi.nonlog, width = 20, height = 10, units = "in", dpi = 300)
+
+# Combine the plots with patchwork for all years of kelp and rate of change
+production.hpi.log <- production_plot | hpi_plot_log +
+  plot_annotation(tag_levels = 'A') + 
+  theme(plot.tag = element_text(size = 20, face = "bold"));production.hpi.log
+
+# Save the combined plot
+ggsave("output/figures/production.HPI.log.png", production.hpi.log, width = 20, height = 10, units = "in", dpi = 300)
 

### 2024-07-10 scripts/manuscript_figures.Rmd
updated mmanuscript figures and habitat productivity index

@@
 morph.filt <- morph %>%
-        filter(SP_CODE == "SL")
+        filter(SP_CODE == "SL" | SP_CODE == "LADI")
 
@@
 # merge morph filt and quad.filt to calculate biomass
-merge.morph.quad <- left_join(morph.filt, quad.filt)
+merge.morph.quad <- rbind.fill(morph.filt, quad.filt)
 
@@
 # Identify the outliers
-outliers <- c(45, 37, 29,37)
+outliers <- c(48,19)
 
@@
 # Extract the p-value for the kelp biomass coefficient
-p_value_kelp_biomass <- lmm_summary$coefficients["log(mean_kelp_biomass + 1.1)", "Pr(>|t|)"]
+p_value_kelp_biomass <- ""
 
@@
   annotate("text", x = Inf, y = -Inf, 
-           label = paste0("p = ", format(p_value_kelp_biomass, scientific = FALSE, digits = 2)), 
+           label = paste0("p < 0.0001", format(p_value_kelp_biomass, scientific = FALSE, digits = 2)), 
            hjust = 1.3, vjust = -39, size = 4) +
@@
   scale_shape_manual(values = shape_values) +
-  labs(x = "Year", y = expression("Mean Kelp Biomass (g/" ~ m^2 ~ ")"), shape = "Site") +
+  labs(x = "Year", y = expression("Kelp Biomass (g/" ~ m^2 ~ ")"), shape = "Site") +
   theme_bw() +
@@
   scale_shape_manual(values = shape_values) +
-  labs(x = "Year", y = expression("Log Mean Kelp Biomass (g/" ~ m^2 ~ ")"), shape = "Site") +
+  labs(x = "Year", y = expression("Kelp Biomass (g/" ~ m^2 ~ ")"), shape = "Site") +
   theme_bw() +
@@
     x = expression("Rate of Change ("~yr^-1 ~")"),
-    y = expression("Mean Kelp Biomass (g/"~m^2~")"), shape = "Site") +
+    y = expression("Kelp Biomass (g/"~m^2~")"), shape = "Site") +
   theme_bw() +
@@
     legend.title = element_text(size = 18), ,
-    legend.position = "none");rate_of_change
+    legend.position = "none") +
+  scale_x_continuous(limits = c(-0.5, 0.0), expand = c(0, 0));rate_of_change
 
@@
 ggsave("output/figures/kelp.2019.rate.png", kelp.rate.2019, width = 20, height = 8, units = "in", dpi = 300)
-
-```
-
-# Organize rugosity data and make linear regression figure for fish biomass and rugosity
-```{r}
-# Calculate mean rugosity complexity scores across depths for each transect, grouped by site and year
-mean_rugosity <- rugosity %>%
-  group_by(SITE, YEAR, TRANSECT) %>%
-  summarize(mean_rugosity = mean(c(X0M, X4M, X8M, X12M, X16M, X20M, X24M, X28M, X32M, X36M, X40M), na.rm = TRUE))
-
-# change transect from integer to chr
-mean_rugosity$TRANSECT <- as.character(mean_rugosity$TRANSECT)
-
-# Left jon fish and rugosity data
-merged.dat.rug <- left_join(mean_rugosity, fish.filt)
-
-```
-
-# Make figure 4 and run statistical analysis
-```{r}
-# Aggregate the data by year, site, and transect and compute mean values
-aggregated.data.rug <- merged.dat.rug %>%
-  group_by(YEAR, SITE, TRANSECT) %>%
-  summarize(mean_fish_biomass = mean(Fish.Biomass..g.m2., na.rm = TRUE),
-            mean_rugosity = mean(mean_rugosity, na.rm = TRUE)) %>%
-  ungroup()
-
-# Fit the linear mixed effects model
-lmm_model <- lmer(mean_fish_biomass ~ mean_rugosity + SITE + (1 | SITE/YEAR), data = aggregated.data.rug)
-#check normality
-qqPlot(residuals(lmm_model)) #data is normal no need to transform
-
-# Fit the linear mixed effects model with log transformation
-lmm_model <- lmer(log(mean_fish_biomass + 1.1) ~ log(mean_rugosity + 1.1) + SITE + (1 | SITE/YEAR), data = aggregated.data.rug)
-#check normality
-qqPlot(residuals(lmm_model)) #improved to normal
-
-# Identify the outliers
-outliers <- c(25, 37, 24, 35)
-
-# remove outliers
-data_clean.rug <- aggregated.data.rug[-outliers, ]
-
-# rerun model with log and without outliers
-lmm_model <- lmer(log(mean_fish_biomass + 1.1) ~ log(mean_rugosity + 1.1) + SITE + (1 | SITE/YEAR), data = data_clean.rug)
-#check normality
-qqPlot(residuals(lmm_model)) #improved to normal
-
-# Summarize the model
-lmm_summary <- summary(lmm_model)
-lmm_tidy <- tidy(lmm_model)
-
-# Calculate R-squared for the mixed model
-r_squared <- r.squaredGLMM(lmm_model)
-
-# Extract marginal and conditional R-squared
-marginal_r2 <- round(r_squared[1], 2)
-conditional_r2 <- round(r_squared[2], 2)
-
-# Extract the p-value for the rugosity
-p_value_rugosity <- lmm_summary$coefficients["log(mean_rugosity + 1.1)", "Pr(>|t|)"]
-
-#create scatter plot
-figure.4 <- ggplot(aggregated.data.rug, aes(x = log(mean_rugosity + 1.1), y = log(mean_fish_biomass + 1.1))) +
-  geom_point() +
-  geom_smooth(method = "lm", se = TRUE, formula = y ~ x, color = "blue") +
-
-  # R-squared values
-  annotate("text", x = Inf, y = Inf, 
-           label = paste0("Marginal R^2 = ", marginal_r2, "\nConditional R^2 = ", conditional_r2), 
-           hjust = 1.1, vjust = 1.1, size = 4) +
-
-  # p-value
-  annotate("text", x = Inf, y = -Inf, 
-           label = paste0("p = ", format(p_value_rugosity, scientific = FALSE, digits = 2)), 
-           hjust = 1.3, vjust = -38.5, size = 4) +
-
-  labs(x = "Rugosity Complexity", y = "Fish biomass (g/m2)",
-       title = "Fish Biomass vs Rugosity Complexity") +
-  theme_classic() +
-  theme(plot.margin = margin(t = 10, r = 30, b = 10, l = 10, unit = "pt")) +
-  coord_cartesian(clip = 'off');figure.4
-
-# Save the plot to a file
-ggsave("output/figures/figure_4.png", plot = figure.4, width = 8, height = 6)
-```
-
-# Filter quad data for all inverts and use to correlate to fish biomass
-```{r}
-#invert.filt <-
 
@@
 
-# Plot two plots together using patchwork for production and HPI
+# Create box plot x axis as type (control and treatment) and y is HPI for kelp 
+```{r}
+
+```
+
+# Plot three plots together using patchwork for production and HPI for fish and kelp biomass
 ```{r}

### 2024-07-10 scripts/manuscript_figures.Rmd
calcualting kelp HPI

@@
 ```{r}
+# First, need to designate control and kelp treatments in the data frame for each year at each site, first filter 2016-2018 years because no 5 and 6 transects where used
+# Filter out 2016-2018 YEAR data
+HPI.kelp <- subset(morph.filt, YEAR < 2016 | YEAR > 2018)
+
+# Rename KB q to transect 6
+HPI.kelp$TRANSECT[HPI.kelp$TRANSECT == "KB q"] <- "6"
+
+# Now create a treatment column that assigns TRANSECTs 5 and 6 from each year as TREATMENT = control and all other transectes TREATMENT = kelp
+HPI.kelp <- HPI.kelp %>%
+  mutate(TREATMENT = ifelse(TRANSECT %in% c(5, 6), "control", "kelp"))
+
+# Calculate habitat productivity index for all transects
 

### 2024-07-10 scripts/manuscript_figures.Rmd
added functional group change over time data

@@
 library(plyr)
+library(dunn.test)
 library(kableExtra)
@@
 
-# Calculate habitat productivity index for all transects
+# Calculate habitat productivity index for all transects using HPI = average kelp biomass/area surveyed (80)
+# Define the area surveyed (80 square meters for each transects)
+area_surveyed <- 80
+
+# Calculate the HPI for each transect
+HPI.kelp <- HPI.kelp %>%
+  mutate(HPI = total_kelp_biomass / area_surveyed)
+
+# Summarize mean and standard error of HPI by treatment
+HPI_summary <- HPI.kelp %>%
+  group_by(TREATMENT) %>%
+  summarise(
+    mean_HPI = mean(HPI, na.rm = TRUE),
+    se_HPI = sd(HPI, na.rm = TRUE) / sqrt(n()),
+    .groups = 'drop'
+  )
+
+# Create the box plot
+hpi_box_plot <- ggplot(HPI.kelp, aes(x = TREATMENT, y = HPI, fill = TREATMENT)) +
+  geom_boxplot() +
+  geom_point(data = HPI_summary, aes(x = TREATMENT, y = mean_HPI), shape = 21, size = 3, position = position_dodge(0.75)) +
+  labs(x = "Treatment", y = "Habitat Productivity Index (HPI)") +  # Rename legend text
+  scale_fill_manual(values = c("Control" = "#4E79A7", "Kelp" = "#F28E2B")) +  # Manual fill colors and labels
+  theme_bw() +
+  theme(axis.text.x = element_text(size = 18),
+        axis.text.y = element_text(size = 18),
+        axis.title = element_text(size = 18),
+        legend.text = element_text(size = 16),
+        legend.title = element_text(size = 18));hpi_box_plot
 
@@
 # Combine the plots with patchwork for all years of kelp and rate of change
-production.hpi.nonlog <- production_plot | hpi.plot.nonlog +
+production.hpi.nonlog <- (production_plot | hpi.plot.nonlog | hpi_box_plot) +
   plot_annotation(tag_levels = 'A') + 
@@
 # Combine the plots with patchwork for all years of kelp and rate of change
-production.hpi.log <- production_plot | hpi_plot_log +
+production.hpi.log <- (production_plot | hpi_plot_log | hpi_box_plot) +
   plot_annotation(tag_levels = 'A') + 
@@
 Number of permutations: 999
+
+# Figure for functional group change in abundance over time using invert count data
+```{r}
+# Check for normality of residuals
+# Log transform the abundance data
+invert.count$log_abundance <- log(invert.count$abundance + 1.1)
+
+# Check for normality of transformed data
+anova_model <- aov(log_abundance ~ type * YEAR, data = invert.count)
+residuals <- residuals(anova_model)
+
+# Plot Q-Q plot
+qqnorm(residuals)
+qqline(residuals, col = "blue")
+
+# Perform Shapiro-Wilk test
+shapiro_test <- shapiro.test(residuals)
+shapiro_test
+
+# still not normal, use non-parametric test
+
+# Perform Kruskal-Wallis test for interaction type and YEAR
+interaction_data <- invert.count %>%
+  mutate(interaction = interaction(type, YEAR))
+
+kruskal_test_interaction <- kruskal.test(abundance ~ interaction, data = interaction_data)
+print(kruskal_test_interaction)
+
+# Extract relevant information from the test result
+kruskal_summary <- data.frame(
+  Test = "Kruskal-Wallis",
+  Interaction = as.character(kruskal_test_interaction$method),
+  ChiSquare = kruskal_test_interaction$statistic,
+  DF = kruskal_test_interaction$parameter,
+  P.Value = kruskal_test_interaction$p.value
+)
+
+# Create flextable object
+flex <- flextable(kruskal_summary)
+
+# Save flextable as PNG
+png_file <- "output/model.result.tables/kruskal_test_interaction_summary.png"
+save_as_image(flex, png_file)
+
+# Assuming invert.count is your dataframe and abundance_g_per_m2 is already calculated
+total_abundance <- invert.count %>%
+  group_by(YEAR) %>%
+  summarise(total_abundance_g_per_m2 = sum(abundance))
+
+# merge total abundance with original data
+invert.abun <- merge(invert.count, total_abundance, by = "YEAR")
+
+# calculate relative abundance
+invert.abun <- invert.abun %>%
+  mutate(relative_abundance = (abundance / total_abundance_g_per_m2) * 100)
+
+# Calculate mean and standard error per functional group and year
+mean_std_data <- invert.abun %>%
+  group_by(YEAR, type) %>%
+  summarise(
+    mean_rel_abundance = mean(relative_abundance),
+    std_error = sd(relative_abundance) / sqrt(n()))
+
+# Convert YEAR to factor
+mean_std_data$YEAR <- factor(mean_std_data$YEAR)
+
+# Example: Renaming groups in mean_std_data dataframe
+mean_std_data <- mean_std_data %>%
+  mutate(type = case_when(
+    type == "crustacean" ~ "Crustaceans",
+    type == "echinoderms" ~ "Echinoderms",
+    type == "molluscs" ~ "Molluscs",
+    TRUE ~ type  # Keep other cases unchanged
+  ))
+
+# Plot mean abundance with error bars
+func.group.plot.line <- ggplot(mean_std_data, aes(x = YEAR, y = mean_rel_abundance, color = type, group = type)) +
+  geom_errorbar(aes(ymin = mean_rel_abundance - std_error, ymax = mean_rel_abundance + std_error), width = 0.2) +
+  geom_line() +
+  geom_point(size = 3) +
+  labs(x = "Year", y = "Relative Abundance (%)", color = "Functional Group") +
+  theme_bw() +
+  theme(
+    axis.title.x = element_text(size = 18),
+    axis.title.y = element_text(size = 18),
+    axis.text.x = element_text(size = 18),
+    axis.text.y = element_text(size = 18),
+    legend.text = element_text(size = 16),
+    legend.title = element_text(size = 18));func.group.plot.line
+
+# Save the plot to a file
+ggsave("output/figures/functional.plot.line.png", plot = func.group.plot.line, width = 15, height = 10)
+
+# Plot stacked bar chart
+stacked_bar_plot <- ggplot(mean_std_data, aes(x = factor(YEAR), y = mean_rel_abundance, fill = type)) +
+  geom_bar(stat = "identity", position = "stack", color = "black") +
+  labs(x = "Year", y = "Relative Abundance (%)", fill = "Functional Group") +
+  theme_bw() +
+  theme(
+    axis.title.x = element_text(size = 18),
+    axis.title.y = element_text(size = 18),
+    axis.text.x = element_text(size = 14, angle = 45, hjust = 1),
+    axis.text.y = element_text(size = 14),
+    legend.text = element_text(size = 14),
+    legend.title = element_text(size = 16),
+    panel.grid.major = element_blank(),
+    panel.grid.minor = element_blank()) +
+  scale_y_continuous(limits = c(0, 50), expand = c(0, 0));stacked_bar_plot
+
+# Save the plot to a file
+ggsave("output/figures/functional.plot.stacked.png", plot = stacked_bar_plot, width = 15, height = 10)
+
+```
+	Kruskal-Wallis rank sum test
+
+data:  abundance by interaction
+Kruskal-Wallis chi-squared = 104.02, df = 23, p-value = 2.83e-12

### 2024-07-12 scripts/manuscript_figures.Rmd
updated HPI and fish production code

@@
 quad.filt <- QUAD %>%
-  filter(SP_CODE == "SUGK" | SP_CODE == "LADI")
+  filter(SP_CODE == "SUGK" | SP_CODE == "SUGK.1")
 
@@
 morph.filt <- morph %>%
-        filter(SP_CODE == "SL" | SP_CODE == "LADI")
+        filter(SP_CODE == "SL")
 
@@
 
-# Calculate fish biomass 'W' using the length-weight relationship
+# Calculate fish weight at maturity using the length-weight relationship
 PB_filtered$Wmat <- PB_filtered$coef.a * (PB_filtered$Lmat^PB_filtered$coef.b)
@@
 
-# Calculate fish biomass
-PB_filtered$biomass <- PB_filtered$Wmat * PB_filtered$COUNT
+# add an individual fish.mass column where we take (PB_filtered$coef.a * (PB_filtered$actual_size ^ PB_filtered$coef.b)
+PB_filtered$ind.fish.mass <- (PB_filtered$coef.a * (PB_filtered$actual_size ^ PB_filtered$coef.b))
+
+# Calculate fish biomass: coefa * actual size ^b * COUNT
+PB_filtered$biomass <- PB_filtered$ind.fish.mass * PB_filtered$COUNT
+
+# add total biomass for each transect per species
+PB.aggregated <- PB_filtered %>%
+  group_by(YEAR, SITE, TRANSECT, SP_CODE) %>%
+  summarise(tot.biomass = sum(biomass, na.rm = TRUE),
+            .groups = 'drop')
 
 # Calculate biomass per m²
-PB_filtered$biomass_per_m2 <- PB_filtered$biomass / 80
+PB.aggregated$biomass_per_m2 <- PB.aggregated$tot.biomass / 80
 

### 2024-07-16 scripts/manuscript_figures.Rmd
make fish production plots and updated calculations

@@
 
-Production to biomass ratios (P:B) and habitat productivity index for each fish species
+Production to biomass ratios (P:B) and fish produciton plots
 ```{r}
@@
 PB.aggregated <- PB_filtered %>%
-  group_by(YEAR, SITE, TRANSECT, SP_CODE) %>%
+  group_by(YEAR, SITE, TRANSECT, SP_CODE, PB.ratio, growth.coef) %>%
   summarise(tot.biomass = sum(biomass, na.rm = TRUE),
@@
 PB_filtered$YEAR <- factor(PB_filtered$YEAR)
-
-# Calculate production by multiplying biomass_per_m2 by PB raatio
-PB_filtered$Production <- PB_filtered$PB.ratio * PB_filtered$biomass_per_m2
+PB.aggregated$YEAR <- factor(PB.aggregated$YEAR)
+
+# Calculate production by methods of Ricker (1971) where P = G x B (P: production; G: instantaneous growth coefficient; B: average biomass)
+# Step 1: Calculate average biomass
+calc.prod <- PB.aggregated %>%
+  group_by(YEAR, SITE, TRANSECT, SP_CODE, PB.ratio, growth.coef) %>%
+  summarize(avg_biomass = mean(biomass_per_m2))
+
+# Step 2: Calculate production
+calc.prod$production <- calc.prod$growth.coef * calc.prod$avg_biomass
 
 # make commmon name column
-PB_filtered <- PB_filtered %>%
+calc.prod <- calc.prod %>%
   mutate(Species = case_when(
@@
 
+# Now create a treatment column that assigns TRANSECTs 5 and 6 from each year as TREATMENT = control and all other transectes TREATMENT = kelp
+calc.prod <- calc.prod %>%
+  mutate(TREATMENT = ifelse(TRANSECT %in% c(5, 6), "control", "kelp"))
+
+# create summary data for per year and treatment
+pb_summary <- calc.prod  %>%
+  group_by(YEAR, TREATMENT) %>%
+  summarize(mean_production = mean(production, na.rm = TRUE),
+            se_production = sd(production, na.rm = TRUE) / sqrt(n()))
+
 # Now you can use this new 'Species' column in your ggplot
-production_plot <- ggplot(PB_filtered, aes(x = YEAR, y = Production, fill = Species)) +
-  geom_bar(stat = "identity", position = "dodge") +
+production_plot <- ggplot(pb_summary, aes(x = YEAR, y = mean_production, color = TREATMENT, group = TREATMENT)) +
+  geom_point(size = 2) +         # Add points
+  geom_line() +          # Add lines
+  geom_errorbar(aes(ymin = mean_production - se_production, ymax = mean_production + se_production), width = 0.2) +  # Add error bars
   labs(x = "Year",  # X-axis label
        y = expression("Fish Production" ~ (g ~ m^-2 ~ yr^-1)),  # Y-axis label
-       fill = "Species") +  # This should now correctly set the legend title
-  scale_fill_viridis_d() +  # Use viridis colorblind-friendly palette
+       color = "Habitat") +  # Legend title
   theme_bw() +
@@
         legend.title = element_text(size = 18)) +
-  scale_y_continuous(limits = c(0, 4), expand = c(0, 0));production_plot  # Additional attempt to change legend title  # Use viridis colorblind-friendly palette
+  scale_y_continuous(limits = c(0, 8), expand = c(0, 0));production_plot # Additional attempt to change legend title  # Use viridis colorblind-friendly palette
+
+# Save the plot
+ggsave("output/figures/production.habitat.png", production_plot_A, width = 20, height = 10, units = "in", dpi = 300)
+
+# Subset the data for control and kelp treatments
+control_data <- subset(pb_summary, TREATMENT == "control")
+kelp_data <- subset(pb_summary, TREATMENT == "kelp")
+
+# Merge to align control and kelp data by year
+merged_data <- merge(control_data, kelp_data, by = "YEAR", suffixes = c("_control", "_kelp"))
+
+# Calculate the differences
+merged_data$production_difference <- merged_data$mean_production_kelp - merged_data$mean_production_control
+merged_data$se_difference <- sqrt(merged_data$se_production_kelp^2 + merged_data$se_production_control^2)
+
+# Step 4: Plot the data
+production_diff_plot <- ggplot(merged_data, aes(x = YEAR, y = production_difference, group = TREATMENT_kelp)) +
+  geom_point(size = 2) +  # Points with light blue outline
+  geom_line() +  # Lines with light blue color
+  geom_errorbar(aes(ymin = production_difference - se_difference, ymax = production_difference + se_difference), 
+                width = 0.2, color = "black", size = 0.7) +  # Error bars
+  labs(x = "Year",  # X-axis label
+       y = expression("Kelp Habitat Fish Production" ~ (g ~ m^-2 ~ yr^-1)),  # Y-axis label
+       color = "Treatment") +  # Legend title
+  theme_bw() +  # White background with black lines
+  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 18),  # Rotate x-axis labels if needed
+        axis.title.x = element_text(vjust = -0.5, size = 18),  # Adjust x-axis title position
+        axis.title.y = element_text(vjust = 1.5, size = 18),  # Adjust y-axis title position
+        axis.text.y = element_text(size = 18),  # Adjust y-axis text size
+        legend.text = element_text(size = 16),  # Legend text size
+        legend.title = element_text(size = 18));production_diff_plot
+
+# Save the plot
+ggsave("output/figures/kelp.fish.production.png", production_diff_plot, width = 20, height = 10, units = "in", dpi = 300)
+
+# Calculate the average of mean_production_kelp
+average_production_kelp <- mean(merged_data$mean_production_kelp)
+
+# Print the average
+average_production_kelp # 4.359285 g/m2/yr
+```
+
+# Plot all two panels together using patchwork
+```{r}
+# Combine the plots with patchwork
+prod_combined_plot <- (production_plot | production_diff_plot) +
+  plot_annotation(tag_levels = 'A') + 
+  theme(plot.tag = element_text(size = 20, face = "bold"));prod_combined_plot
+
+# Save the combined plot
+ggsave("output/figures/figure_1.png", combined_plot, width = 12, height = 8, units = "in", dpi = 300)
+
+```
 
 # Calculate Habitat Productivity Index (HPI), habitat productivity index (HPI), a proposed measure of habitat suitability and productive capacity, can be calculated as the product of average fish biomass and the species P/B coefficient (HPI = B × P/B) (Randall and Minns 2000)
+```{r}
 # Calculate HPI
@@
 
-# Define the color palette
-#my_palette <- brewer.pal(n = 6, name = "Paired")  # Adjust 'n' as per the number of years in your data
-
 # Test for normality and perform appropriate statistical test
@@
 alternative hypothesis: true location shift is not equal to 0
-
 # Create box plot x axis as type (control and treatment) and y is HPI for kelp 

### 2024-07-16 scripts/manuscript_figures.Rmd
annotated production values

@@
 production_plot <- ggplot(pb_summary, aes(x = YEAR, y = mean_production, color = TREATMENT, group = TREATMENT)) +
-  geom_point(size = 2) +         # Add points
+  geom_point(size = 3) +         # Add points
   geom_line() +          # Add lines
@@
         legend.title = element_text(size = 18)) +
-  scale_y_continuous(limits = c(0, 8), expand = c(0, 0));production_plot # Additional attempt to change legend title  # Use viridis colorblind-friendly palette
+  scale_y_continuous(limits = c(0, 8), expand = c(0, 0));production_plot # Additional attempt to change legend title 
+#blue color is #00BFC4
 
@@
 production_diff_plot <- ggplot(merged_data, aes(x = YEAR, y = production_difference, group = TREATMENT_kelp)) +
-  geom_point(size = 2) +  # Points with light blue outline
-  geom_line() +  # Lines with light blue color
+  geom_point(size = 3) +  # Points in skyblue color
+  geom_line() +  # Lines in skyblue color
   geom_errorbar(aes(ymin = production_difference - se_difference, ymax = production_difference + se_difference), 
@@
 
+# Update production_diff_plot with annotations
+production_diff_plot_annot <- production_diff_plot +
+  geom_text(aes(label = sprintf("%.2f", production_difference), y = production_difference),
+            color = "black", size = 4, vjust = 2.42, hjust = 1.06, check_overlap = TRUE);production_diff_plot_annot
+
 # Save the plot
@@
 # Save the combined plot
-ggsave("output/figures/figure_1.png", combined_plot, width = 12, height = 8, units = "in", dpi = 300)
+ggsave("output/figures/fish_production.png", prod_combined_plot, width = 12, height = 8, units = "in", dpi = 300)
 

### 2024-08-01 scripts/manuscript_figures.Rmd
updated figure 5 plots and stats

--- 2024-07-16 scripts/manuscript_figures.Rmd
@@
 
-#combine rugosity and kelp biomass
-master.dat2 <- left_join(mean_rugosity, morph.filt)
-
 # combine for final dataframe
-master.dat <- left_join(master.dat1, master.dat2)
+master.dat <- left_join(master.dat1, morph.filt)
 
@@
 #select necessary columns
-all.params.data <- master.dat %>% dplyr::select("YEAR", "SITE", "COVER", "Fish.Biomass..g.m2.", "mean_rugosity", "average_kelp_biomass")
+all.params.data <- master.dat %>% dplyr::select("YEAR", "SITE", "COVER", "Fish.Biomass..g.m2.", "average_kelp_biomass")
 
@@
 names(all.params.data)[4] <- "Fish Biomass (g/cm2)"
-names(all.params.data)[5] <- "Rugosity"
-names(all.params.data)[6] <- "Kelp Biomass (g/cm2)"
+names(all.params.data)[5] <- "Kelp Biomass (g/cm2)"
 
 #reorder columns to show biomass by eachother
-all.params.data <- all.params.data[, c(1,2, 4, 6, 5, 3)]
+all.params.data <- all.params.data[, c(1,2, 4, 5, 3)]
 
@@
 # Select the relevant columns
-selected_columns <- all.params.data[, c("Fish Biomass (g/cm2)", "Kelp Biomass (g/cm2)", "Rugosity", "Kelp Cover (%)")]
+selected_columns <- all.params.data[, c("Fish Biomass (g/cm2)", "Kelp Biomass (g/cm2)", "Kelp Cover (%)")]
 
@@
 shapiro.test(all.params.data$`Fish Biomass (g/cm2)`) 
-shapiro.test(all.params.data$Rugosity)
 shapiro.test(all.params.data$`Kelp Biomass (g/cm2)`)
@@
 
-# Multiple regression between fish biomass (g/m2), kelp biomass (g/m2), kelp cover, rugosity, temperature in ggplot 
+# Multiple regression between fish biomass (g/m2), kelp biomass (g/m2), kelp cover,  temperature in ggplot 
 # combine all dataframes
@@
 # Rename columns to remove spaces and special characters
-names(all.params.data) <- c("YEAR", "SITE", "Fish_Biomass", "Rugosity", "Kelp_Cover", "Kelp_Biomass")
+names(all.params.data) <- c("YEAR", "SITE", "Fish_Biomass", "Kelp_Cover", "Kelp_Biomass")
 
@@
 # Check for zero or negative values
-zero_or_negative <- sapply(all.params.data[, c("Fish_Biomass", "Rugosity", "Kelp_Biomass", "Kelp_Cover")], function(x) sum(x <= 0))
+zero_or_negative <- sapply(all.params.data[, c("Fish_Biomass", "Kelp_Biomass", "Kelp_Cover")], function(x) sum(x <= 0))
 zero_or_negative
@@
 all.params.data <- all.params.data[all.params.data$Fish_Biomass > 0 & 
-                                   all.params.data$Rugosity > 0 & 
                                    all.params.data$Kelp_Biomass > 0 & 
@@
 
-# Fit a multiple regression model
-fit1 <- lm(log(Fish_Biomass + 1.1) ~ log(Rugosity + 1.1) * log(Kelp_Biomass + 1.1) * log(Kelp_Cover + 1.1), data = all.params.data)
-summary(fit1)
-
-# Diagnostic plots for the regression model
-par(mfrow=c(2,2))
-plot(fit1)
-
-# Produce added multivariable plots
-av_plots <- avPlots(fit1)
-
-#The x-axis displays a single predictor variable and the y-axis displays the response variable.
-#The blue line shows the association between the predictor variable and the response variable, while holding the value of all other predictor variables constant.
-#The points that are labelled in each plot represent the 2 observations with the largest residuals and the 2 observations with the largest partial leverage.
-
-#view results of model
-summary_text1 <- capture.output(summary(fit1))
-writeLines(summary_text1, "output/model.result.tables/rugosity.included/summary_original.model.txt")
-
-```
-
-# Try another type of multiple regession plot
-```{r}
-#fit multiple linear regression model WITH SITE AND YEAR
-fit2 <- lm(Fish_Biomass ~ SITE*YEAR, data = all.params.data)
-summary(fit2)
-
-#view results of model
-summary_text2 <- capture.output(summary(fit2))
-writeLines(summary_text2, "output/model.result.tables/rugosity.included/summary_model_site_year_fixed_fishbiomass.txt")
-
-# Fit the mixed-effects model
-fit3 <- lmer(log(Fish_Biomass) ~ log(Rugosity) * log(Kelp_Biomass) * log(Kelp_Cover) + 
-              (1 | SITE) + (1 | YEAR), data = all.params.data)
-summary(fit3)
-
-#view results of model
-summary_text3 <- capture.output(summary(fit3))
-writeLines(summary_text3, "output/model.result.tables/rugosity.included/summary_model_site_year_random.txt")
-
-# site and year no significant interaction or significant main effects
-
-# run with just site as random effect
-# Fit the mixed-effects model
-fit4 <- lmer(log(Fish_Biomass + 1.1) ~ log(Rugosity + 1.1) * log(Kelp_Biomass + 1.1) * log(Kelp_Cover + 1.1) + 
-              (1 | SITE), data = all.params.data)
-summary(fit4)
-
-#view results of model
-summary_text4 <- capture.output(summary(fit4))
-writeLines(summary_text4, "output/model.result.tables/rugosity.included/summary_model_site_random.txt")
-
-# site and year no significant main effects
-
-# run with just year as random effect
-# Fit the mixed-effects model
-fit5 <- lmer(log(Fish_Biomass + 1.1) ~ log(Rugosity + 1.1) * log(Kelp_Biomass + 1.1) * log(Kelp_Cover + 1.1) + 
-              (1 | YEAR), data = all.params.data)
-summary(fit5)
-
-#view results of model
-summary_text5 <- capture.output(summary(fit5))
-writeLines(summary_text5, "output/model.result.tables/rugosity.included/summary_model_year_random.txt")
-
-#year is significantly different for kelp cover
-
-```
-
-# Run all above models without rugosity
+```
+
+# Run all models without rugosity
 ```{r}
@@
   theme_bw() +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 18),  # Rotate x-axis labels if needed
-        axis.title.x = element_text(vjust = -0.5, size = 18),  # Adjust x-axis title position
+  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 18),
+        axis.title.x = element_text(vjust = -0.5, size = 18),
         axis.title.y = element_text(vjust = 1.5, size = 18),
@@
         legend.text = element_text(size = 16),
-        legend.title = element_text(size = 18)) +
-  scale_y_continuous(limits = c(0, 8), expand = c(0, 0));production_plot # Additional attempt to change legend title  # Use viridis colorblind-friendly palette
+        legend.title = element_text(size = 18),
+        legend.position = "top") +
+  scale_y_continuous(limits = c(0, 8), expand = c(0, 0));production_plot # Additional attempt to change legend title 
+#blue color is #00BFC4
 
 # Save the plot
-ggsave("output/figures/production.habitat.png", production_plot_A, width = 20, height = 10, units = "in", dpi = 300)
+ggsave("output/figures/production.habitat.png", production_plot, width = 20, height = 10, units = "in", dpi = 300)
 
@@
 # Calculate the differences
-merged_data$production_difference <- merged_data$mean_production_kelp - merged_data$mean_production_control
+merged_data$production_difference <- merged_data$mean_production_kelp + merged_data$mean_production_control
 merged_data$se_difference <- sqrt(merged_data$se_production_kelp^2 + merged_data$se_production_control^2)
@@
        color = "Treatment") +  # Legend title
-  theme_bw() +  # White background with black lines
-  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 18),  # Rotate x-axis labels if needed
-        axis.title.x = element_text(vjust = -0.5, size = 18),  # Adjust x-axis title position
-        axis.title.y = element_text(vjust = 1.5, size = 18),  # Adjust y-axis title position
-        axis.text.y = element_text(size = 18),  # Adjust y-axis text size
-        legend.text = element_text(size = 16),  # Legend text size
-        legend.title = element_text(size = 18));production_diff_plot
+  theme_bw() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 18),
+        axis.title.x = element_text(vjust = -0.5, size = 18),
+        axis.title.y = element_text(vjust = 1.5, size = 18),
+        axis.text.y = element_text(size = 18),
+        legend.position = "none");production_diff_plot
+
+# Update production_diff_plot with annotations
+production_diff_plot_annot <- production_diff_plot +
+  geom_text(aes(label = sprintf("%.2f", production_difference), y = production_difference),
+            color = "black", size = 4, vjust = 2.8, hjust = 1.06, check_overlap = TRUE);production_diff_plot_annot
 
 # Save the plot
-ggsave("output/figures/kelp.fish.production.png", production_diff_plot, width = 20, height = 10, units = "in", dpi = 300)
+ggsave("output/figures/kelp.fish.production.png", production_diff_plot_annot, width = 20, height = 10, units = "in", dpi = 300)
 
@@
 average_production_kelp # 4.359285 g/m2/yr
+
+# Check normality and run appropriate statistical test:
+
+# Separate data into kelp and control
+kelp_data <- calc.prod %>% filter(TREATMENT == "kelp") %>% pull(production)
+control_data <- calc.prod %>% filter(TREATMENT == "control") %>% pull(production)
+
+# Shapiro-Wilk normality test
+shapiro_kelp <- shapiro.test(kelp_data)
+shapiro_control <- shapiro.test(control_data)
+
+# Print results
+print(shapiro_kelp)
+print(shapiro_control)
+
+# Data is not normally distributed so run Mann-Whitney U Test (Wilcoxon Rank-Sum Test)
+# Perform Mann-Whitney U Test
+wilcox_test_result <- wilcox.test(kelp_data, control_data)
+
+# Print results
+print(wilcox_test_result)
+
+# Calculate mean and standard error for each treatment
+kelp.prod.dat <- calc.prod %>%
+  group_by(TREATMENT) %>%
+  summarize(mean_production = mean(production, na.rm = TRUE),
+            se_production = sqrt(sum(production^2)) / n())
+
+# Create the plot with mean adn st error
+# Create the base plot with point ranges
+kelp.treatment <- ggplot(kelp.prod.dat, aes(x = TREATMENT, y = mean_production, fill = TREATMENT)) +
+  geom_pointrange(aes(ymin = mean_production - se_production, ymax = mean_production + se_production), 
+                  size = 1.2, color = "black", shape = 21, fill = "white") +
+  labs(x = "Treatment",
+       y = expression("Fish Production" ~ (g ~ m^-2 ~ yr^-1))) +
+  theme_bw() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 18),
+        axis.title.x = element_text(vjust = -0.5, size = 18),
+        axis.title.y = element_text(vjust = 1.5, size = 18),
+        axis.text.y = element_text(size = 18),
+        legend.position = "none"); kelp.treatment
+
+# Calculate the maximum value for y-axis
+max_y <- max(kelp.prod.dat$mean_production + kelp.prod.dat$se_production)
+
+# Add the Wilcoxon test results as an annotation
+kelp.treatment <- kelp.treatment + annotate("text", x = 0.5, y = max_y, 
+                        label = "Wilcoxon test\np < 0.0001",
+                        hjust = 0, vjust = 1, size = 5, color = "black");kelp.treatment
+
+# Add significance asterisks
+kelp.treatment <- kelp.treatment + geom_text(aes(x = 1.5, y = max_y - 0.2 , label = "***"), size = 6, color = "black"); kelp.treatment
+
 ```
@@
 ```{r}
+
 # Combine the plots with patchwork
-prod_combined_plot <- (production_plot | production_diff_plot) +
+prod_combined_plot <- (production_plot | kelp.treatment) +
   plot_annotation(tag_levels = 'A') + 
-  theme(plot.tag = element_text(size = 20, face = "bold"));prod_combined_plot
+  theme(plot.tag = element_text(size = 18, face = "bold"));prod_combined_plot
 
 # Save the combined plot
-ggsave("output/figures/figure_1.png", combined_plot, width = 12, height = 8, units = "in", dpi = 300)
+ggsave("output/figures/fish_production.png", prod_combined_plot, width = 20, height = 10, units = "in", dpi = 300)
 

### 2024-08-01 scripts/manuscript_figures.Rmd
updated tables

@@
 
+# Perform ANOVA on the mixed model
+anova_results <- Anova(lmm_model, type = 3)
+print(anova_results)
+
+# Extract coefficients and their standard errors
+model_summary <- broom.mixed::tidy(lmm_model)
+
 # Create a data frame with coefficients, standard errors, and p-values
@@
 
-# Create flextable object
-flex <- flextable(df)
+# Display the coefficients table
+print(df)
 
@@
   plot_annotation(tag_levels = 'A') + 
-  theme(plot.tag = element_text(size = 18, face = "bold"));prod_combined_plot
+   theme(plot.tag = element_text(face = "bold", size = 18));prod_combined_plot
 
@@
 Kruskal-Wallis chi-squared = 104.02, df = 23, p-value = 2.83e-12
+
+# Make publishable table for fish production and life history information
+```{r}
+# read in data
+life.hist <- read.csv("data/fish.life.history.data.csv")
+
+# Create a gt table
+fish_table <- gt(life.hist) %>%
+  tab_header(
+    title = "Fish Species Life History Traits"
+  ) %>%
+  fmt_number(
+    columns = vars(ALPHA, BETA, Lmat, Wmat, P.B.ratio, K.growth.coefficient),
+    decimals = 2
+  ) %>%
+  tab_style(
+    style = list(
+      cell_text(weight = "bold")
+    ),
+    locations = cells_title(groups = "title")
+  ) %>%
+  tab_style(
+    style = list(
+      cell_text(weight = "bold")
+    ),
+    locations = cells_column_labels(everything())
+  )
+
+# Save as HTML
+gtsave(fish_table, file = "output/figures/life_history_table.html")
+
+# Save as PDF (requires the webshot2 package)
+# webshot::install_phantomjs() # Run this once if you haven't installed PhantomJS
+gtsave(fish_table, file = "output/figures/life_history_table.pdf", expand = 2)
+
+```

### 2024-08-07 scripts/manuscript_figures.Rmd
cleaned up manuscript script and output folders

@@
 
-nMDS or PCA on benthic community structure
+Invertebrate community structure calculations
 ```{r}
@@
 
-# Calculate Bray-Curtis dissimilarity
-m_com <- vegdist(numeric_data, method = "bray")
-
-# Perform nMDS
-nmds <- metaMDS(m_com, distance = "bray")
-
-# Extract scores
-scores <- as.data.frame(scores(nmds))
-
-# Add row numbers as a unique identifier
-scores$ID <- row.names(scores)
-
-# Create a data frame with SITE and YEAR information
-site_year_info <- wide_data[rowSums(wide_data[, c("crustacean", "echinoderms", "molluscs")]) > 0, c("SITE", "YEAR")]
-site_year_info$ID <- row.names(site_year_info)
-
-# Merge SITE and YEAR information into scores
-scores <- merge(scores, site_year_info, by = "ID")
-
-# Define custom shape values for more than 8 categories
-custom_shape_values <- c(0, 1, 2, 3, 4, 5, 6, 23)
-
-# Use these custom shapes in your plot
-nmds_plot <- ggplot(scores, aes(x = NMDS1, y = NMDS2, color = SITE, shape = as.factor(YEAR))) +
-  geom_point(size = 8) +
-  #stat_ellipse(aes(group = YEAR), level = 0.95, linetype = 2, size = 1, alpha = 0.5) + # Add ellipses
-  labs(x = "NMDS1", y = "NMDS2", color = "Site", shape = "Year") +
-  scale_color_manual(values = c("Fort Wetherill" = "grey", "King's Beach" = "black")) +
-  scale_shape_manual(values = custom_shape_values) +  # Specify the custom shapes
-  theme_bw() +
-  theme(
-    axis.title.x = element_text(size = 18),
-    axis.title.y = element_text(size = 18),
-    axis.text.x = element_text(size = 18),
-    axis.text.y = element_text(size = 18),
-    legend.text = element_text(size = 16),
-    legend.title = element_text(size = 18)
-  );nmds_plot
-
-# Save the plot to a file
-ggsave("output/figures/nmds.plot.png", plot = nmds_plot, width = 15, height = 10)
-
-```
-
-PERMANOVA and SIMPER analysis
+```
+
+Figure and calculations for functional group change in relative abundance over time using invert count data
 ```{r}
@@
 
-# Create environmental data frame
-environmental_data <- wide_data[non_zero_rows, c("TRANSECT", "YEAR", "SITE")]
-
-# Convert YEAR to factor
-environmental_data$YEAR <- as.factor(environmental_data$YEAR)
-
-# Perform PERMANOVA
-permanova_result <- adonis2(community_data ~ SITE + YEAR, data = environmental_data, permutations = 999, method = "bray")
-
-# View the results
-print(permanova_result)
-
-# Create dataframe from PERMANOVA results
-permanova_df <- data.frame(
-  Factor = c("SITE", "YEAR", "Residual", "Total"),
-  Df = c(1, 7, 44, 52),
-  SumOfSqs = c(0.5649, 5.0492, 8.7778, 14.3919),
-  R2 = c(0.03925, 0.35084, 0.60991, 1.00000),
-  F = c(2.8317, 3.6157, NA, NA),
-  Pr_F = c(0.050, 0.001, NA, NA)  # Changed column name to Pr_F
-)
-
-# Create flextable object
-flex_perm <- flextable(permanova_df)
-
-# Customize the flextable
-flex_perm <- flex_perm %>%
-  set_header_labels(
-    Factor = "Factor",
-    Df = "Df",
-    SumOfSqs = "Sum of Squares",
-    R2 = "R2",
-    F = "F",
-    Pr_F = "Pr(>F)"  # Changed to match the new column name
-  ) %>%
-  colformat_double(digits = 4) %>%
-  bold(i = ~ Pr_F <= 0.05, j = "Pr_F") %>%  # Bold values if Pr_F is 0.05 or lower
-  add_header_lines(values = c(
-    "Permutation test for adonis under reduced model",
-    "Number of permutations: 999"
-  )) %>%
-  autofit() %>%
-  theme_box()
-
-# Save flextable as PNG
-png_file <- "output/model.result.tables/permanova.png"
-save_as_image(flex_perm, path = png_file)
-
-# Perform SIMPER analysis
-simper_result <- simper(community_data, group = environmental_data$SITE, permutations = 999)
-
-# View summary of results
-summary(simper_result)
-
-# Create dataframe from SIMPER results
-simp <- data.frame(
-  Species = rownames(summary(simper_result)$`Fort Wetherill_King's Beach`),
-  Average = summary(simper_result)$`Fort Wetherill_King's Beach`$average,
-  Ratio = summary(simper_result)$`Fort Wetherill_King's Beach`$ratio,
-  Fort_Wetherill = summary(simper_result)$`Fort Wetherill_King's Beach`$ava,
-  Kings_Beach = summary(simper_result)$`Fort Wetherill_King's Beach`$avb,
-  Cumulative_Sum = summary(simper_result)$`Fort Wetherill_King's Beach`$cumsum,
-  p_value = summary(simper_result)$`Fort Wetherill_King's Beach`$p
-)
-
-# Create flextable object
-flex_simp <- flextable(simp)
-
-# Customize the flextable
-flex_simp <- flex_simp %>%
-  set_header_labels(
-    Species = "Species",
-    Average = "Avg. Contribution",
-    Ratio = "Ratio",
-    Fort_Wetherill = "Fort Wetherill Abundance",
-    Kings_Beach = "King's Beach Abundance",
-    Cumulative_Sum = "Cumulative Sum",
-    p_value = "p-value"
-  ) %>%
-  autofit() %>%
-  theme_box()
-
-# Save flextable as PNG
-png_file <- "output/model.result.tables/simper_results.png"
-save_as_image(flex_simp, path = png_file)
-
-# Extract the relevant information from the SIMPER results
-simper_data <- data.frame(
-  Species = rownames(summary(simper_result)$`Fort Wetherill_King's Beach`),
-  Contribution = summary(simper_result)$`Fort Wetherill_King's Beach`$average,
-  Fort_Wetherill = summary(simper_result)$`Fort Wetherill_King's Beach`$ava,
-  Kings_Beach = summary(simper_result)$`Fort Wetherill_King's Beach`$avb
-)
-
-# Calculate the percentage contribution
-simper_data$Percentage <- simper_data$Contribution / sum(simper_data$Contribution) * 100
-
-# Reshape the data for plotting
-simper_long <- tidyr::pivot_longer(simper_data, 
-                                   cols = c(Fort_Wetherill, Kings_Beach),
-                                   names_to = "Site", 
-                                   values_to = "Abundance")
-
-# Create the plot
-simper_plot <- ggplot(simper_long, aes(x = Species, y = Abundance, fill = Site)) +
-  geom_bar(stat = "identity", position = "dodge") +
-  geom_text(aes(label = sprintf("%.1f%%", Percentage[match(Species, simper_data$Species)])),
-            position = position_dodge(width = 0.9),
-            vjust = -0.5, 
-            size = 3) +
-  labs(title = "Species Contributions to Differences Between Sites",
-       subtitle = "Based on SIMPER analysis",
-       y = "Average Abundance",
-       x = "Species") +
-  scale_fill_manual(values = c("Fort_Wetherill" = "grey", "Kings_Beach" = "black")) +
-  theme_minimal() +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1),
-        legend.title = element_blank());simper_plot
-
-# Save the plot to a file
-ggsave("output/figures/simper.plot.png", plot = simper_plot, width = 15, height = 10)
-
-```
-
-PERMANOVA results:
-adonis2(formula = community_data ~ SITE + YEAR, data = environmental_data, permutations = 999, method = "bray")
-         Df SumOfSqs      R2      F Pr(>F)    
-SITE      1   0.5649 0.03925 2.8317  0.043 *  
-YEAR      7   5.0492 0.35084 3.6157  0.001 ***
-Residual 44   8.7778 0.60991                  
-Total    52  14.3919 1.00000
-
-SIMPER results:
-ontrast: Fort Wetherill_King's Beach
-
-            average      sd   ratio     ava     avb cumsum     p
-molluscs    0.58750 0.29784 1.97250 5.57600 5.84300  0.863 0.576
-echinoderms 0.07140 0.12811 0.55760 0.21400 0.34300  0.968 0.547
-crustacean  0.02160 0.05018 0.43030 0.21900 0.09300  1.000 0.526
-Permutation: free
-Number of permutations: 999
-
-# Figure for functional group change in abundance over time using invert count data
-```{r}
+# Calculate total abundance 
+total_abundance <- invert.count %>%
+  group_by(YEAR) %>%
+  summarise(total_abundance_g_per_m2 = sum(abundance))
+
+# merge total abundance with original data
+invert.abun <- merge(invert.count, total_abundance, by = "YEAR")
+
+# calculate relative abundance
+invert.abun <- invert.abun %>%
+  mutate(relative_abundance = (abundance / total_abundance_g_per_m2) * 100)
+
 # Check for normality of residuals
 # Log transform the abundance data
-invert.count$log_abundance <- log(invert.count$abundance + 1.1)
+invert.abun$log_rel_abundance <- log(invert.abun$abundance + 1.1)
 
 # Check for normality of transformed data
-anova_model <- aov(log_abundance ~ type * YEAR, data = invert.count)
+anova_model <- aov(log_rel_abundance ~ type * YEAR, data = invert.abun)
 residuals <- residuals(anova_model)
@@
 # Perform Kruskal-Wallis test for interaction type and YEAR
-interaction_data <- invert.count %>%
+interaction_data <- invert.abun %>%
   mutate(interaction = interaction(type, YEAR))
 
-kruskal_test_interaction <- kruskal.test(abundance ~ interaction, data = interaction_data)
+kruskal_test_interaction <- kruskal.test(log_rel_abundance ~ interaction, data = interaction_data)
 print(kruskal_test_interaction)
@@
 save_as_image(flex, png_file)
-
-# Assuming invert.count is your dataframe and abundance_g_per_m2 is already calculated
-total_abundance <- invert.count %>%
-  group_by(YEAR) %>%
-  summarise(total_abundance_g_per_m2 = sum(abundance))
-
-# merge total abundance with original data
-invert.abun <- merge(invert.count, total_abundance, by = "YEAR")
-
-# calculate relative abundance
-invert.abun <- invert.abun %>%
-  mutate(relative_abundance = (abundance / total_abundance_g_per_m2) * 100)
 
@@
 
-data:  abundance by interaction
+data:  log_rel_abundance by interaction
 Kruskal-Wallis chi-squared = 104.02, df = 23, p-value = 2.83e-12
